---
title: "Analysis of CRC Spatial Transcriptomics data from 10X: Pathway Activity using Progeny"
author: 'Alberto Valdeolivas Urbelz (PS-Biomics&Path)'
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    code_folding: hide
    self_contained: true
    number_sections: yes
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: spacelab
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
params:
  data_directory : "/projects/site/pred/SpatialOmics/Experiment_CRC_AllSamples/"
  analysis_name : "analysis_alberto/"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

The present script takes the the Spatial Omics data generated with 10X 
technology on a set of CRC samples and explores the pathway activity per spot
using Progeny. 

```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(vctrs)
library(patchwork, lib.loc = "/apps/rocs/2020.08/cascadelake/software/R/4.1.2-foss-2020a/lib64/R/library")
library(Seurat)
# library(tidyverse)
library(dplyr)
library(ggplot2)
library(tidyr)
library(stringr)
library(readr)
library(purrr)
library(progeny)
source(file = "WrapperFunction/SeuratWrappers.R")

data_directory <-  params$data_directory
analysis_name <- params$analysis_name

input_files <- "IntermediaryFiles/"
input_names <- "SeuratList_Clusters_Res05.rds"
output_names <- "SeuratList_Clusters_Res05_Progeny.rds"

files_to_read <- paste0(data_directory, analysis_name,input_files,input_names)
results_directory <- paste0(data_directory, analysis_name, input_files)
# setwd(paste0(data_directory, analysis_name))
```

We read the Seurat objects files after Quality Control

```{r, warning=FALSE}
seurat_objects <- readRDS(files_to_read)
```

# Computing Pathway activity with Progeny

We compute Pathway activity using Progeny

```{r, message=FALSE, warning=FALSE}
seurat_objects <- get.progeny.scores(seurat_objects)
# pathway_plots <- get.pathway.plots(seurat_objects, pt.size.factor = 2.25)
```

# Results {.tabset}

We display the spatial plots showing pathway activity per spot and violing plots
showing pathway activity per cluster.

```{r, results = 'asis', message=FALSE, dpi=80, warning=FALSE}
for ( sample in names(seurat_objects)){
  
  cat( "## Sample : ", sample , "\n\n" )
  
  cat( "Showing sample : ", sample , "\n\n")
  
  pathways <- 
    unique(rownames(GetAssayData(seurat_objects[[sample]], assay = "progeny")))
  DefaultAssay(seurat_objects[[sample]]) <- "progeny"
  
  for (j in seq(from = 1,to= length(pathways), by=2)){
    current_pathways <- pathways[c(j,j+1)]
    print(SpatialFeaturePlot(seurat_objects[[sample]], 
      features = current_pathways, ncol = 2, alpha = 1, pt.size.factor = 2))  
    cat( "\n\n")
    print(VlnPlot(seurat_objects[[sample]], 
                  features = current_pathways, ncol=2))
    cat( "\n\n")
  }
}
```

# Conclusions

# Saving the results

```{r, eval=FALSE}
saveRDS(seurat_objects, file = paste0(results_directory, output_names))
```

# Session Info Details

```{r, echo=FALSE, eval=TRUE}
sessionInfo()
```
