---
title: "Analysis of CRC Spatial Transcriptomics data from 10X: CMS2 Inter-patient heterogeneity"
author: 'Alberto Valdeolivas Urbelz (PS-Biomics&Path)'
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    code_folding: hide
    self_contained: true
    number_sections: yes
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: spacelab
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
params:
  data_directory : "/pstore/data/biomics/_pre_portfolio/_platform_evaluation/SpatialTranscriptomics/Experiment_CRC_AllSamples/"
  analysis_name : "analysis_alberto/"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

The present script takes Spatial Omics data generated with 10X technology
of a set of colorectal cancer samples and explores the inter-patient 
hetereogeneity in those samples that have been classified as CMS2. To do so, 
we will perform pseudo-bulk of the CMS2 regions. 

```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(vctrs)
library(patchwork, lib.loc = "/apps/rocs/2020.08/cascadelake/software/R/4.1.2-foss-2020a/lib64/R/library")
library(Seurat)
library(dplyr)
library(tibble)
library(readr)
library(stringr)
library(ggplot2)
library(cowplot)
library(msigdbr)
library(clusterProfiler)
library(RColorBrewer)

source(file = "WrapperFunction/SeuratWrappers.R")
source(file = "WrapperFunction/PseudoBulkUtils.R")

data_directory <-  params$data_directory
analysis_name <- params$analysis_name

input_files <- "IntermediaryFiles/"
input_names <- "SeuratList_Clusters_Res05_Dorothea.rds"


files_to_read <- paste0(data_directory, analysis_name,input_files,input_names)
results_directory <- paste0(data_directory, analysis_name, input_files)
```

we read the results of the deconvolution containing the total
number of cells per spot. From that, we compute the cell type proportions per 
spot. 


## Integration

```{r warning=FALSE, message=FALSE}
seurat_objects_all <- readRDS(files_to_read)

samples_to_consider <- c("SN048_A121573_Rep1", "SN048_A121573_Rep2",
  "SN123_A595688_Rep1","SN124_A595688_Rep2", "SN123_A938797_Rep1_X", 
  "SN124_A938797_Rep2", "SN84_A120838_Rep1", "SN84_A120838_Rep2") 

seurat_objects <- seurat_objects_all[samples_to_consider]

Seurat.merge <- merge(seurat_objects[[1]], seurat_objects[-1])

DefaultAssay(Seurat.merge) <- "dorothea"
Seurat.merge <- ScaleData(Seurat.merge)
Seurat.merge <- RunPCA(Seurat.merge, verbose = FALSE, features = rownames(Seurat.merge))
Seurat.merge <- FindNeighbors(Seurat.merge, dims = 1:10, verbose = FALSE)
Seurat.merge <- FindClusters(Seurat.merge, verbose = FALSE, resolution = 0.5)
Seurat.merge <- RunUMAP(Seurat.merge, dims = 1:10)
```


```{r, warning=FALSE, message=FALSE, dpi=100}
palette <- RColorBrewer::brewer.pal(n=length(levels(Seurat.merge)),"Paired")

df_patient_ID_transform <- data.frame(
  patient = c("A120838","A121573","A416371","A551763","A595688","A798015",
              "A938797"), 
  patient_ID = c("S4_Col_Sig","S5_Rec","S3_Col_R","S1_Cec","S2_Col_R",
                 "S7_Rec/Sig","S6_Rec"))

Seurat.merge@meta.data$patient_ID <- 
  str_match(Seurat.merge@meta.data$orig.ident, "_\\s*(.*?)\\s*_")[,2] %>%
  as_tibble_col(column_name = "patient") %>% 
  dplyr::left_join(df_patient_ID_transform) %>%
  dplyr::pull(patient_ID)

DimPlot(Seurat.merge, reduction = "umap", # shape.by = "dorothea_snn_res.0.5", 
        pt.size = 0.75, cols = palette, split.by = "patient_ID") + 
  theme(axis.ticks = element_blank(), axis.text = element_blank(), 
        legend.title = element_text(size=14, face="bold")) +
  guides(col =guide_legend(title="Cluster ID", override.aes = list(size = 3.5))) 
```

```{r, warning=FALSE, message=FALSE}
umap_clustering <- 
  DimPlot(Seurat.merge, reduction = "umap", # shape.by = "dorothea_snn_res.0.5", 
        pt.size = 0.75, cols = palette) + 
  theme(axis.ticks = element_blank(), axis.text = element_blank(), 
        legend.title = element_blank(), legend.justification = "center", 
        plot.title = element_text(hjust = 0.5)) +
  guides(col =guide_legend(title="Cluster ID", override.aes = list(size = 3.5))) + 
  labs(title = "Cluster ID") 

DefaultAssay(object = Seurat.merge) <- "dorothea"

umap_MYC <- 
  FeaturePlot(Seurat.merge, reduction = "umap", features = "MYC", 
        pt.size = 1.25) +  
  theme(axis.ticks = element_blank(), axis.text = element_blank(), 
        legend.text = element_text(size=14), 
        legend.position = "right", legend.justification = "center") +
  labs(title = "MYC TF Activity")  + 
  scale_color_gradient2(low = "lightblue", mid ="white", high = "#67000D")

umap_E2F4 <- 
  FeaturePlot(Seurat.merge, reduction = "umap", features = "E2F4", 
        pt.size = 1.25) + 
  theme(axis.ticks = element_blank(), axis.text = element_blank(), 
        legend.text = element_text(size=14), 
        legend.position = "right", legend.justification = "center") +
  labs(title = "E2F4 TF Activity")   + 
  scale_color_gradient2(low = "lightblue", mid ="white", high = "#67000D")

umap_JUN <- 
  FeaturePlot(Seurat.merge, reduction = "umap", features = "JUN", 
        pt.size = 1.25) +  #, cols =c("blue", "#67000D")) + 
  theme(axis.ticks = element_blank(), axis.text = element_blank(), 
        legend.text = element_text(size=14), 
        legend.position = "right", legend.justification = "center") +
  labs(title = "JUN TF Activity")  + 
  scale_color_gradient2(low = "lightblue", mid ="white", high ="#67000D")

umap_ETS1 <- 
  FeaturePlot(Seurat.merge, reduction = "umap", features = "ETS1", 
        pt.size = 1.25) + 
  theme(axis.ticks = element_blank(), axis.text = element_blank(), 
        legend.text = element_text(size=14), 
        legend.position = "right", legend.justification = "center") +
  labs(title = "ETS1 TF Activity")  + 
  scale_color_gradient2(low = "lightblue", mid ="white", high = "#67000D")
```

```{r, warning=FALSE, message=FALSE, dpi=300, fig.height=6, fig.width=10}
umap_clustering + umap_MYC + umap_ETS1
```


## Clustering Results {.tabset}

```{r, warning=FALSE, message=FALSE, results='asis'} 
a <- 1
for (current_sample in samples_to_consider){
  
  cat( "### Sample :", current_sample ,"\n\n" )  
  
  seurat_subset <- 
    subset(x = Seurat.merge, subset = orig.ident == current_sample)
  
  print(SpatialDimPlot(seurat_subset, images =  Images(seurat_subset)[a], 
    pt.size.factor = 1.75, image.alpha = 0.25, cols = palette, stroke = 0))# 
  
  
  print(SpatialDimPlot(seurat_subset, 
    cells.highlight = CellsByIdentities(object = seurat_subset, 
    idents = levels(Idents(seurat_subset))), 
                 facet.highlight = TRUE, ncol = 3, 
    images = Images(seurat_subset)[a],
    cols.highlight=c("blue", "red"), image.alpha = 0.25))
  
  
  a <- a + 1 
  
  cat( "\n\n" ) 
  
}
```

### Cluster Plot Publication

```{r, warning=FALSE, message=FALSE}
current_sample <- "SN048_A121573_Rep1"
a <- 1

seurat_subset <- 
    subset(x = Seurat.merge, subset = orig.ident == current_sample)


spatial_plot_A121573_Rep1 <- 
  SpatialDimPlot(seurat_subset, images =  Images(seurat_subset)[a], 
    pt.size.factor = 1.75, image.alpha = 0.25, cols = palette, stroke = 0) + 
  NoLegend() + theme_void() + 
  theme(legend.position ="bottom",  
        legend.text = element_text(size=12, face="bold", family="Arial"), 
        legend.title = element_text(size=14, face="bold", family="Arial")) + 
  guides(fill=guide_legend(title="Cluster ID", override.aes = list(size = 5)))


### 

current_sample <- "SN048_A121573_Rep2"
a <- 2

seurat_subset <- 
    subset(x = Seurat.merge, subset = orig.ident == current_sample)


spatial_plot_A121573_Rep2 <- 
  SpatialDimPlot(seurat_subset, images =  Images(seurat_subset)[a], 
    pt.size.factor = 1.75, image.alpha = 0.25, cols = palette, stroke = 0) + NoLegend() 

### 

current_sample <- "SN123_A595688_Rep1"
a <- 3

seurat_subset <- 
    subset(x = Seurat.merge, subset = orig.ident == current_sample)


spatial_plot_A595688_Rep1 <- 
  SpatialDimPlot(seurat_subset, images =  Images(seurat_subset)[a], 
    pt.size.factor = 2, image.alpha = 0.25, cols = palette, stroke = 0) + NoLegend() 

### 

current_sample <- "SN124_A595688_Rep2"
a <- 4

seurat_subset <- 
    subset(x = Seurat.merge, subset = orig.ident == current_sample)


spatial_plot_A595688_Rep2 <- 
  SpatialDimPlot(seurat_subset, images =  Images(seurat_subset)[a], 
    pt.size.factor = 2.25, image.alpha = 0.25, cols = palette, stroke = 0) + NoLegend() 

### 

current_sample <- "SN123_A938797_Rep1_X"
a <- 5

seurat_subset <- 
    subset(x = Seurat.merge, subset = orig.ident == current_sample)


spatial_plot_A938797_Rep1 <- 
  SpatialDimPlot(seurat_subset, images =  Images(seurat_subset)[a], 
    pt.size.factor = 1.75, image.alpha = 0.25, cols = palette, stroke = 0) + NoLegend() 

### 

current_sample <- "SN124_A938797_Rep2"
a <- 6

seurat_subset <- 
    subset(x = Seurat.merge, subset = orig.ident == current_sample)


spatial_plot_A938797_Rep2 <- 
  SpatialDimPlot(seurat_subset, images =  Images(seurat_subset)[a], 
    pt.size.factor = 1.75, image.alpha = 0.25, cols = palette, stroke = 0) + NoLegend() 

### 

current_sample <- "SN84_A120838_Rep1"
a <- 7

seurat_subset <- 
    subset(x = Seurat.merge, subset = orig.ident == current_sample)


spatial_plot_A120838_Rep1 <- 
  SpatialDimPlot(seurat_subset, images =  Images(seurat_subset)[a], 
    pt.size.factor = 2.75, image.alpha = 0.25, cols = palette, stroke = 0) + NoLegend() 

### 

current_sample <- "SN84_A120838_Rep2"
a <- 8

seurat_subset <- 
    subset(x = Seurat.merge, subset = orig.ident == current_sample)


spatial_plot_A120838_Rep2 <- 
  SpatialDimPlot(seurat_subset, images =  Images(seurat_subset)[a], 
    pt.size.factor = 2.25, image.alpha = 0.25, cols = palette, stroke = 0) + NoLegend() 
```

```{r, warning=FALSE, message=FALSE}

clusters_legend <- get_legend(spatial_plot_A121573_Rep1)
spatial_plot_A121573_Rep1_noLegend <- spatial_plot_A121573_Rep1 + NoLegend()

spatial_plot_A595688_Rep1 +  plot_spacer() + spatial_plot_A595688_Rep2 + 
  spatial_plot_A120838_Rep1 + plot_spacer() + spatial_plot_A120838_Rep2 + 
  spatial_plot_A121573_Rep1_noLegend + plot_spacer() + spatial_plot_A121573_Rep2 +
  spatial_plot_A938797_Rep1 + plot_spacer() + spatial_plot_A938797_Rep2 + 
  plot_spacer() +  ggpubr::as_ggplot(clusters_legend) + plot_spacer() + 
  plot_layout(widths = c(5, -2.5, 5), heights = c(2,2,2,2,0.25))
```




## Cluster TF markers {.tabset}

```{r, warning=FALSE, message=FALSE}
TFmarkers <- 
  FindAllMarkers(Seurat.merge, only.pos = TRUE, min.pct = 0.25, 
    logfc.threshold = 1, verbose = FALSE, test.use="roc")

visualization_TFs <- TFmarkers %>% 
  dplyr::group_by(cluster) %>% 
  dplyr::slice_max(avg_log2FC, n = 10) %>% 
  dplyr::filter(avg_log2FC>1.25) %>% 
  dplyr::pull(gene)
```



```{r, warning=FALSE, message=FALSE}
DoHeatmap(Seurat.merge, features = visualization_TFs, angle = 0, 
          size=4, hjust=0.5,group.colors = palette) + NoLegend() + 
  scale_y_discrete(labels=c("MYC"=expression(bold("MYC")), "JUN"=expression(bold("JUN")),
      "E2F4"=expression(bold("E2F4")), "ETS1"=expression(bold("ETS1")), parse=TRUE))
```


```{r, warning=FALSE, message=FALSE, results='asis'}
for (current_tf in visualization_TFs){
  
  cat( "### TF :", current_tf ,"\n\n" )
  
  print(VlnPlot(Seurat.merge, features = current_tf))
  cat( "\n\n")
    
} 
```

## Enrichment Results {.tabset}

```{r, warning=FALSE, message=FALSE}
my_categories <- c("H")
my_sub_categories <- 
   c("CP:KEGG", "CP:REACTOME", "CP:BIOCARTA", "GO:BP", "GO:CC","GO:MF")
all_gene_sets <- msigdbr(species = "Homo sapiens")

gene_sets_selected <- 
  dplyr::filter(all_gene_sets, 
    (gs_subcat %in% my_sub_categories)  |  (gs_cat %in% my_categories))

msigdbr_t2g <- gene_sets_selected %>% 
   dplyr::select(gs_name, gene_symbol) %>% 
  as.data.frame()

```

```{r, warning=FALSE, message=FALSE, results='asis'}
nr_results <- 12
threshold <- 0.01
all_genes <- rownames(Seurat.merge)

for (current_cluster in unique(Idents(Seurat.merge)) ){

  selected_genes <- TFmarkers %>% 
    dplyr::filter(cluster == current_cluster) %>% 
    dplyr::pull(gene)
  
  enrichment_results <- enricher(selected_genes,  universe = all_genes, 
   pvalueCutoff = 0.01,  pAdjustMethod = "fdr", TERM2GENE = msigdbr_t2g)
  
  Toplot_df <- enrichment_results@result %>% 
        dplyr::top_n(nr_results, desc(pvalue )) %>% 
        # ungroup() %>% 
        mutate(logPvalue = -log(pvalue)) %>%
        mutate(ID_toplot = substr(ID, 1,50))
      
  cat( "### Cluster :", current_cluster ,"\n\n" )
  
  plot_enrichment <- ggplot(Toplot_df, aes(logPvalue, 
                                               reorder(ID_toplot, logPvalue))) +
        geom_col(aes(fill=logPvalue)) + # facet_wrap(~ source) + 
        scale_fill_gradient(low = "#FFA07A", high = "#800000", 
                            na.value = "whitesmoke") +   
        theme_minimal() + 
        theme(legend.position = "bottom", 
              legend.text = element_text(size = 8, hjust = 1, angle = 90),
              axis.text.x = element_text(angle = 0, hjust = 0, size = 8, face="bold"), 
              axis.text.y = element_text(angle = 0, hjust = 1, size = 8, face="bold")) + 
        # ylab(paste0("Pathways and Ontologies")) +
        ylab(paste0("")) + 
        xlab(paste0("-Log (p-Value)")) + 
        labs(fill = "-Log (p-Value)") + 
        scale_x_continuous(position = "bottom") + 
        # facet_grid(vars(source), scales = "free", space = "free") + 
        geom_vline(xintercept = -log(threshold), linetype = "dashed", 
                   colour = "steelblue", size = 1) + 
        ggtitle(paste0("sample: ", current_sample))
  
  print(plot_enrichment)
  
  cat( "\n\n")
    
}  
```


```{r, warning=FALSE,message=FALSE, eval=TRUE}
filename_SeuratObject <- 
  paste0(data_directory, analysis_name, input_files, 
      "MergeClustering/", "SeuratObject_dorothea_res05.rds")

saveRDS(list(TFmarkers, Seurat.merge), file = filename_SeuratObject)


Seurat.merge@meta.data %>% 
  tibble::rownames_to_column(var="spot_id") %>% 
  plyr::mutate(spot_id = str_remove(spot_id, "_[1-9]")) %>% 
  dplyr::mutate(spot_id=paste0("Count_",orig.ident, "_",spot_id)) %>%
  dplyr::mutate(spot_id=str_replace(spot_id, "SN123_A938797_Rep1_X","SN123_A938797_Rep1")) %>%
  dplyr::mutate(orig.ident = str_replace(orig.ident, "SN123_A938797_Rep1_X", "SN123_A938797_Rep1")) %>%
  write_delim(file = paste0(data_directory, analysis_name, input_files, 
      "MergeClustering/", "MetadataClustersDorothea.tsv"), delim = "\t", 
      col_names = TRUE)
```


# Session Info Details

```{r, echo=FALSE, eval=TRUE}
sessionInfo()
```


