---
title: "Analysis of CRC Spatial Transcriptomics data from 10X: Misty"
author: 'Alberto Valdeolivas Urbelz (PS-Biomics&Path)'
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    code_folding: hide
    self_contained: true
    number_sections: yes
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: spacelab
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
params:
  data_directory : "/pstore/data/biomics/_pre_portfolio/_platform_evaluation/SpatialTranscriptomics/Experiment_CRC_AllSamples/"
  analysis_name : "analysis_alberto/"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

```{r, warning=FALSE, message=FALSE}
library(dplyr, lib.loc = "/pstore/apps/bioinfo/R/4.0.1-foss")
library(patchwork, lib.loc = "/pstore/home/valdeola/R/x86_64-pc-linux-gnu-library/4.0.1-foss")
library(Seurat, lib.loc = "/pstore/apps/bioinfo/R/4.0.1-foss")
library(tibble, lib.loc = "/pstore/apps/bioinfo/R/4.0.1-foss")
library(readr, lib.loc = "/pstore/apps/bioinfo/R/4.0.1-foss")
library(stringr, lib.loc = "/pstore/apps/bioinfo/R/4.0.1-foss")
library(ggplot2, lib.loc = "/pstore/apps/bioinfo/R/4.0.1-foss")
library(cowplot, lib.loc = "/pstore/apps/bioinfo/R/4.0.1-foss")
library(msigdbr, lib.loc = "/pstore/apps/bioinfo/R/4.0.1-foss")
library(clusterProfiler, lib.loc = "/pstore/apps/bioinfo/R/4.0.1-foss")
library(RColorBrewer, lib.loc = "/pstore/apps/bioinfo/R/4.0.1-foss")
library(mistyR, lib.loc = "/pstore/apps/bioinfo/R/4.0.1-foss")
library(purrr, lib.loc = "/pstore/apps/bioinfo/R/4.0.1-foss")
library(kableExtra, lib.loc = "/pstore/apps/bioinfo/R/4.0.1-foss")


data_directory <-  params$data_directory
analysis_name <- params$analysis_name

input_files <- "IntermediaryFiles/"
input_names <- "MergeClustering/SeuratObject_dorothea_res05.rds"
input_names_2 <- "SeuratList_Clusters_Res05_Dorothea.rds"

files_to_read <- paste0(data_directory, analysis_name,input_files,input_names)
files_to_read_2 <- paste0(data_directory, analysis_name,input_files,input_names_2)
results_directory <- paste0(data_directory, analysis_name, input_files)
```

```{r, warning=FALSE, message=FALSE}
## Misty required functions
run_misty_seurat <- function(visium.slide,
                             # Seurat object with spatial transcriptomics data.
                             view.assays,
                             # Named list of assays for each view.
                             view.features,
                             # Named list of features/markers to use.
                             # Use all by default.
                             view.types,
                             # Named list of the type of view to construct
                             # from the assay.
                             view.params,
                             # Named list with parameters (NULL or value)
                             # for each view.
                             spot.ids = NULL,
                             # spot IDs to use. Use all by default.
                             out.alias = "results"
                             # folder name for output
) {
  # Extracting geometry
  geometry <- GetTissueCoordinates(visium.slide,
                                   cols = c("row", "col"), scale = NULL
  )
  # Extracting data
  view.data <- map(view.assays,
                   extract_seurat_data,
                   geometry = geometry,
                   visium.slide = visium.slide
  )
  # Constructing and running a workflow
  build_misty_pipeline(
    view.data = view.data,
    view.features = view.features,
    view.types = view.types,
    view.params = view.params,
    geometry = geometry,
    spot.ids = spot.ids,
    out.alias = out.alias
  )
}




# Extracts data from an specific assay from a Seurat object
# and aligns the IDs to the geometry
extract_seurat_data <- function(visium.slide,
                                assay,
                                geometry) {
  data <- GetAssayData(visium.slide, assay = assay) %>% as.matrix() %>%
    t() %>%
    as_tibble(rownames = NA)
  return(data %>% slice(match(rownames(.), rownames(geometry))))
}
# Filters data to contain only features of interest
filter_data_features <- function(data,
                                 features) {
  if (is.null(features)) features <- colnames(data)
  return(data %>% rownames_to_column() %>%
           dplyr::select(rowname, all_of(features)) %>% rename_with(make.names) %>%
           column_to_rownames())
}


## View creation



# Builds views depending on the paramaters defined
create_default_views <- function(data,
                                 view.type,
                                 view.param,
                                 view.name,
                                 spot.ids,
                                 geometry) {
  view.data.init <- create_initial_view(data)
  if (!(view.type %in% c("intra", "para", "juxta"))) {
    view.type <- "intra"
  }
  if (view.type == "intra") {
    data.red <- view.data.tmp$data %>%
      rownames_to_column() %>%
      filter(rowname %in% spot.ids) %>%
      dplyr::select(-rowname)
  } else if (view.type == "para") {
    view.data.tmp <- view.data.init %>%
      add_paraview(geometry, l = view.param)
    data.ix <- paste0("paraview.", view.param)
    data.red <- view.data.tmp[[data.ix]]$data %>%
      mutate(rowname = rownames(data)) %>%
      filter(rowname %in% spot.ids) %>%
      dplyr::select(-rowname)
  } else if (view.type == "juxta") {
    view.data.tmp <- view.data.init %>%
      add_juxtaview(
        positions = geometry,
        neighbor.thr = view.param
      )
    data.ix <- paste0("juxtaview.", view.param)
    data.red <- view.data.tmp[[data.ix]]$data %>%
      mutate(rowname = rownames(data)) %>%
      filter(rowname %in% spot.ids) %>%
      dplyr::select(-rowname)
  }
  if (is.null(view.param) == TRUE) {
    misty.view <- create_view(
      paste0(view.name),
      data.red
    )
  } else {
    misty.view <- create_view(
      paste0(view.name, "_", view.param),
      data.red
    )
  }
  clear_cache(view.data.init$misty.uniqueid)
  return(misty.view)
}


## Building a mistyR pipeline and running the model


# Builds automatic MISTy workflow and runs it
build_misty_pipeline <- function(view.data,
                                 view.features,
                                 view.types,
                                 view.params,
                                 geometry,
                                 spot.ids = NULL,
                                 out.alias = "default") {
  # Adding all spots ids in case they are not defined
  if (is.null(spot.ids)) {
    spot.ids <- rownames(view.data[[1]])
  }
  # First filter the features from the data
  view.data.filt <- map2(view.data, view.features, filter_data_features)
  # Create initial view
  views.main <- create_initial_view(view.data.filt[[1]] %>%
                                      rownames_to_column() %>%
                                      filter(rowname %in% spot.ids) %>%
                                      dplyr::select(-rowname))
  # Create other views
  view.names <- names(view.data.filt)
  all.views <- pmap(list(
    view.data.filt[-1],
    view.types[-1],
    view.params[-1],
    view.names[-1]
  ),
  create_default_views,
  spot.ids = spot.ids,
  geometry = geometry
  )
  pline.views <- add_views(
    views.main,
    unlist(all.views, recursive = FALSE)
  )
  # Run MISTy
  run_misty(pline.views, out.alias)
}
```


## Reading Seurat objects and TF differentially activated 


```{r warning=FALSE, message=FALSE}
seurat_objects_markers <- readRDS(files_to_read)
Seurat_merge_object <- seurat_objects_markers[[2]]
```

We select the TFs whose activity we are aiming at predicting. 

```{r, warning=FALSE, message=FALSE}
TF_to_predict <- seurat_objects_markers[[1]] %>% 
  dplyr::filter(cluster== "1") %>% 
  dplyr::filter(myAUC >= 0.75) %>%
  dplyr::pull(gene) %>% unique()
```

We select genes that are annoated as ligands in OmniPath and expressed in our
regions in every sample. We select the same ligands as used in LIANA for 
consistency with the next section. 

```{r, warning=FALSE, message=FALSE}
omni_resources <- readRDS(system.file(package = "liana", 
  "omni_resources.rds"))
ligands_omni <- unique(omni_resources$OmniPath$source_genesymbol)
```

Which of those ligands are expressed in the tumor region in all of our slides
in more than 15% of the spots? 

```{r, warning=FALSE, message=FALSE}
samples_to_consider <- c("SN048_A121573_Rep1", "SN048_A121573_Rep2",
  "SN123_A595688_Rep1","SN124_A595688_Rep2", "SN123_A938797_Rep1_X", 
  "SN124_A938797_Rep2", "SN84_A120838_Rep1", "SN84_A120838_Rep2") 

coverage_threshold <- 0.1

list_slide_ligands <- list()

cluster_id_tumor_micro <- c("0", "1")
cluster_id_others <- 
  levels(Idents(Seurat_merge_object))[!(levels(Idents(Seurat_merge_object)) %in% cluster_id_tumor_micro)]

DefaultAssay(Seurat_merge_object) = "SCT"
Genes_diff_Expressed <- 
  FindMarkers(Seurat_merge_object, 
            ident.1	= WhichCells(Seurat_merge_object, idents=cluster_id_tumor_micro), 
            ident.2	 = WhichCells(Seurat_merge_object, idents= cluster_id_others),
            logfc.threshold = 0.5, only.pos = TRUE, 
            recorrect_umi=FALSE) %>% rownames()

for (current_sample in samples_to_consider){
  
  current_seurat_obj <- subset(x = Seurat_merge_object, idents = c("0", "1"), 
    subset = orig.ident == current_sample)
  DefaultAssay(current_seurat_obj) = "SCT"
  
  slide_ligands <- ligands_omni[ligands_omni %in% rownames(current_seurat_obj)] 
  
  gex <- as.matrix(current_seurat_obj@assays$SCT@data)[slide_ligands,]
  lig_coverage <- rowSums(gex>0)/ncol(gex)
  slide_ligands <- names(lig_coverage[lig_coverage>=coverage_threshold])
  
  list_slide_ligands[[current_sample]] <- slide_ligands
  
}

ligands_predictors <- 
  names(which(table(unlist(list_slide_ligands)) == 8) )

ligands_predictors_DE <- 
  ligands_predictors[ligands_predictors %in% Genes_diff_Expressed]
```

## Running Misty {.tabset}

We define the views. 

```{r, warning=FALSE, message=FALSE}
# Define assay for each view
view.assays <- list(
  "main" = "dorothea",
  "juxta.ligands" = "SCT"# ,
  # "para.ligands" = "SCT"
)

# Define features for each view
view.features <- list(
  "main" = TF_to_predict,
  "juxta.ligands" = ligands_predictors_DE # , 
  # "para.ligands" = slide_ligands
)

# Define spatial context for each view
view.types <- list(
  "main" = "intra",
  "juxta.ligands" = "juxta"  # ,
#   "para.ligands" = "para"
)

# Define additional parameters (l in the case of paraview)
view.params <- list(
  "main" = NULL,
  "juxta.ligands" = 2 # ,
  # "para.ligands" = 10
)
```


```{r, warning=FALSE, message=FALSE, results='asis'}
seurat_list_objects <- readRDS(files_to_read_2)

for (current_sample in samples_to_consider){
  
  cat( "### Sample :", current_sample ,"\n\n" )  
  
  current_seurat_obj <- seurat_list_objects[[current_sample]]
  DefaultAssay(current_seurat_obj) <- "SCT"
  
  misty.out <- 
    paste0(data_directory, analysis_name, 
         "IntermediaryFiles/Misty_Results/results_TF_Ligands_DorotheaClusters/", 
         current_sample)
  
  misty.results <- run_misty_seurat(
    visium.slide = current_seurat_obj,
    view.assays = view.assays,
    view.features = view.features,
    view.types = view.types,
    view.params = view.params,
    spot.ids = NULL, # Using the whole slide
    out.alias = misty.out
    ) %>%
  collect_results()

  plot_gainR2 <- misty.results %>%
    plot_improvement_stats("gain.R2")
  print(plot_gainR2)
  
  cat( "\n\n" ) 
  
  plot_gainRMSE <- misty.results %>%
    plot_improvement_stats("gain.RMSE") 
  print(plot_gainRMSE)
  
  cat( "\n\n" ) 
  
  print(misty.results$improvements %>%
    filter(measure == "p.R2") %>%
    arrange(value) %>% 
    dplyr::select(-sample) %>% 
    kbl() %>% kable_styling())
  
  cat( "\n\n" ) 
  
  plot_contributions <- misty.results %>% plot_view_contributions()
  print(plot_contributions)
  
  cat( "\n\n" ) 
  
  plot_intra <- 
    misty.results %>% plot_interaction_heatmap(view = "intra") 
  
  cat( "\n\n" ) 
  print(plot_intra)
  
  cat( "\n\n" ) 
  plot_juxta <- 
    misty.results %>% plot_interaction_heatmap(view = "juxta.ligands_2") 
  
  print(plot_juxta)
  
  cat( "\n\n" ) 
  
}
```

# Session Info Details

```{r, echo=FALSE, eval=TRUE}
sessionInfo()
```




