---
title: "Analysis of CRC Spatial Transcriptomics data from 10X: Quality Control: Global Metrics"
author: 'Alberto Valdeolivas Urbelz (PS-Biomics&Path)'
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    code_folding: hide
    self_contained: true
    number_sections: yes
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: spacelab
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
params:
  data_directory : "/projects/site/pred/SpatialOmics/Experiment_CRC_AllSamples/"
  analysis_name : "analysis_alberto/"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Plot Arrangement of Cell2Location results

# Getting Ready

We first load the libraries and set the paths to the raw data. 

```{r, message=FALSE, warning=FALSE}
library(harmony)
library(patchwork, lib.loc = "/apps/rocs/2020.08/cascadelake/software/R/4.1.2-foss-2020a/lib64/R/library")
library(Seurat)
library(readr)
library(stringr)
library(ggplot2)
library(ggpubr)
library(msigdbr)
library(RColorBrewer)
library(progeny)
library(dorothea)

data_directory <-  params$data_directory
analysis_name <- params$analysis_name

input_folder <- "IntermediaryFiles/"
input_names_1 <- "SeuratList_Clusters_Res05.rds"
input_names_2 <- "Patho_Annotations/"

files_to_read_1 <- paste0(data_directory, analysis_name,input_folder,input_names_1)
files_to_read_2 <- paste0(data_directory, analysis_name,input_folder,input_names_2)

df_patient_ID_transform <- data.frame(
  patient = c("A120838","A121573","A416371","A551763","A595688","A798015",
              "A938797"), 
  patient_ID = c("S4_Col_Sig","S5_Rec","S3_Col_R","S1_Cec","S2_Col_R",
                 "S7_Rec/Sig","S6_Rec"))

samples_to_consider <- c("SN048_A121573_Rep1", "SN048_A121573_Rep2",
  "SN123_A595688_Rep1","SN124_A595688_Rep2", "SN123_A938797_Rep1_X", 
  "SN124_A938797_Rep2", "SN84_A120838_Rep1", "SN84_A120838_Rep2")
```

We include pathologist annotations into Seurat objects

```{r, warning=FALSE, message=FALSE}
seurat_objects <- readRDS(files_to_read_1)
seurat_objects_CMS2 <- seurat_objects[samples_to_consider]

## Attention: in one sample annotated as "tumor&stroma IC med to high"
## instead of "tumor&stroma IC med to high"
# tumor_annotations <- 
#  c("tumor", "tumor&stroma_IC med to high", "tumor&stroma_IC low", 
#    "tumor&stroma IC med to high")
tumor_annotations <- 
  c("tumor")

seurat_objects_CMS2 <- 
  lapply(seurat_objects_CMS2, function(x){
    current_sample <- unique(x@meta.data$orig.ident)
    
    patho_anno_current_sample <- 
      read_csv(file = paste0(files_to_read_2, "Pathologist_Annotations_", 
        current_sample, ".csv"))
    
    colnames(patho_anno_current_sample) <- c("Barcode", "Pathologist_Annotations")
    
    x@meta.data <- x@meta.data %>%
      tibble::rownames_to_column("spot_id") %>%
      dplyr::left_join(patho_anno_current_sample, by = c("spot_id" = "Barcode")) %>%
      dplyr::mutate(Pathologist_Annotations = str_replace(Pathologist_Annotations, 
        "tumor&stroma IC med to high", "tumor&stroma_IC med to high")) %>% 
      tibble::column_to_rownames("spot_id") 
      
    
    Idents(x) <- 'Pathologist_Annotations'
    x <- subset(x, idents = intersect(levels(Idents(x)), tumor_annotations))
    
    return(x)
})
```

```{r, warning=FALSE, message=FALSE}
SpatialPlot(seurat_objects_CMS2[[1]], image.alpha = 0.6, stroke = 1, crop = TRUE,
  pt.size.factor = 1.75, cols = "#FC9272") + NoLegend() + 
  theme(plot.margin=grid::unit(c(0,0,0,0), "mm"))
```

```{r, warning=FALSE, message=FALSE}
SpatialPlot(seurat_objects_CMS2[[2]], image.alpha = 0.6, stroke = 1, crop = TRUE,
  pt.size.factor = 1.75, cols = "#FC9272") + NoLegend() + 
  theme(plot.margin=grid::unit(c(0,0,0,0), "mm"))
```

```{r, warning=FALSE, message=FALSE}
SpatialPlot(seurat_objects_CMS2[[3]], image.alpha = 0.6, stroke = 1, crop = TRUE,
  pt.size.factor = 2.5, cols = "#FC9272") + NoLegend() + 
  theme(plot.margin=grid::unit(c(0,0,0,0), "mm"))
```

```{r, warning=FALSE, message=FALSE}
SpatialPlot(seurat_objects_CMS2[[4]], image.alpha = 0.6, stroke = 1, crop = TRUE,
  pt.size.factor = 2.5, cols = "#FC9272") + NoLegend() + 
  theme(plot.margin=grid::unit(c(0,0,0,0), "mm"))
```

```{r, warning=FALSE, message=FALSE}
SpatialPlot(seurat_objects_CMS2[[5]], image.alpha = 0.6, stroke = 1, crop = TRUE,
  pt.size.factor = 2.25, cols = "#FC9272") + NoLegend() + 
  theme(plot.margin=grid::unit(c(0,0,0,0), "mm"))
```

```{r, warning=FALSE, message=FALSE}
SpatialPlot(seurat_objects_CMS2[[6]], image.alpha = 0.6, stroke = 1, crop = TRUE,
  pt.size.factor = 2.25, cols = "#FC9272") + NoLegend() + 
  theme(plot.margin=grid::unit(c(0,0,0,0), "mm"))
```

```{r, warning=FALSE, message=FALSE}
SpatialPlot(seurat_objects_CMS2[[7]], image.alpha = 0.6, stroke = 1, crop = TRUE,
  pt.size.factor = 3.25, cols = "#FC9272") + NoLegend() + 
  theme(plot.margin=grid::unit(c(0,0,0,0), "mm"))
```

```{r, warning=FALSE, message=FALSE}
SpatialPlot(seurat_objects_CMS2[[8]], image.alpha = 0.6, stroke = 1, crop = TRUE,
  pt.size.factor = 2.25, cols = "#FC9272") + NoLegend() + 
  theme(plot.margin=grid::unit(c(0,0,0,0), "mm"))
```


```{r, warning=FALSE, message=FALSE}
merge_seurat_objects_CMS2 <- 
  merge(x = seurat_objects_CMS2[[1]], y= seurat_objects_CMS2[-1])


merge_seurat_objects_CMS2_2 <-  
  CreateSeuratObject(counts = GetAssayData(merge_seurat_objects_CMS2, slot = "counts"), 
    min.cells = 5) %>%
    SCTransform(verbose = FALSE) %>%
    RunPCA(pc.genes = .@var.genes, npcs = 30, verbose = FALSE)

merge_seurat_objects_CMS2_2@meta.data$orig.ident <- 
  merge_seurat_objects_CMS2@meta.data$orig.ident  

merge_seurat_objects_CMS2_2@meta.data$Pathologist_Annotations <- 
  merge_seurat_objects_CMS2@meta.data$Pathologist_Annotations  

rm(seurat_objects_CMS2)
rm(seurat_objects)
rm(merge_seurat_objects_CMS2)

merge_seurat_objects_CMS2_2@meta.data$slide <- 
  merge_seurat_objects_CMS2_2@meta.data$orig.ident %>%
  str_extract("[^_]+")

merge_seurat_objects_CMS2_2@meta.data$patient <- 
  str_match(merge_seurat_objects_CMS2_2@meta.data$orig.ident, "_\\s*(.*?)\\s*_")[,2]

merge_seurat_objects_CMS2_2@meta.data$replicate <- 
  merge_seurat_objects_CMS2_2@meta.data$orig.ident %>%
  str_remove("[^A]+") %>% 
  str_replace("A938797_Rep1_X", "A938797_Rep1")

merge_seurat_objects_CMS2_2@meta.data$rep <- 
  merge_seurat_objects_CMS2_2@meta.data$replicate %>%
  str_remove("[^R]+")  

merge_seurat_objects_CMS2_2@meta.data$patient_ID <- 
  dplyr::inner_join(merge_seurat_objects_CMS2_2@meta.data,df_patient_ID_transform) %>% 
  dplyr::pull(patient_ID)

merge_seurat_objects_CMS2_2@meta.data <- merge_seurat_objects_CMS2_2@meta.data  %>% 
  dplyr::mutate(patient_ID_rep = paste0(.$patient_ID,str_remove(replicate,"[^_Rep]+")))
```

# Results

## Before Batch correction

### PCs before batch correction

```{r, warning=FALSE, message=FALSE, dpi=300}
# palette <- RColorBrewer::brewer.pal(n=14, "Paired")
DimPlot(object = merge_seurat_objects_CMS2_2,  reduction = "pca", 
  pt.size = .25, group.by = "orig.ident")
```


```{r, warning=FALSE, message=FALSE, dpi=300}
VlnPlot(object = merge_seurat_objects_CMS2_2, features = "PC_1", 
              group.by = "orig.ident", pt.size = .1)
```

```{r, warning=FALSE, message=FALSE, dpi=300}
# palette <- RColorBrewer::brewer.pal(n=14, "Paired")
DimPlot(object = merge_seurat_objects_CMS2_2,  reduction = "pca", 
  pt.size = .25, group.by = "slide")
```

```{r, warning=FALSE, message=FALSE, dpi=300}
VlnPlot(object = merge_seurat_objects_CMS2_2, features = "PC_1", 
              group.by = "patient_ID", pt.size = .1)
```

```{r, warning=FALSE, message=FALSE, dpi=300}
# palette <- RColorBrewer::brewer.pal(n=14, "Paired")
DimPlot(object = merge_seurat_objects_CMS2_2,  reduction = "pca", 
  pt.size = .25, group.by = "patient_ID")
```

```{r, warning=FALSE, message=FALSE, dpi=300}
VlnPlot(object = merge_seurat_objects_CMS2_2, features = "PC_1", 
              group.by = "slide", pt.size = .1)
```

### UMAPs before batch correction

```{r, warning=FALSE, message=FALSE, dpi=300}
merge_seurat_objects_CMS2_2 <- merge_seurat_objects_CMS2_2 %>%  RunUMAP(dims = 1:30)
DimPlot(merge_seurat_objects_CMS2_2, reduction = "umap", group.by = "orig.ident", 
        pt.size = .1) 
```

```{r, warning=FALSE, message=FALSE, dpi=300}
DimPlot(merge_seurat_objects_CMS2_2, reduction = "umap", group.by = "slide", 
        pt.size = .1) 
```

```{r, warning=FALSE, message=FALSE, dpi=300}
DimPlot(merge_seurat_objects_CMS2_2, reduction = "umap", group.by = "patient_ID", 
        pt.size = .1) 
```

## Batch correction with Harmony

We correct batch effects per capture area

```{r,warning=FALSE, message=FALSE}
merge_seurat_objects_CMS2_2 <- merge_seurat_objects_CMS2_2 %>% 
    RunHarmony(c("orig.ident"), plot_convergence = TRUE, assay.use = "SCT",
               max.iter.harmony = 20)
```

### PCs after batch correction

```{r, warning=FALSE, message=FALSE, dpi=300}
# palette <- RColorBrewer::brewer.pal(n=14, "Paired")
DimPlot(object = merge_seurat_objects_CMS2_2,  reduction = "harmony", 
  pt.size = .25, group.by = "orig.ident")
```

```{r, warning=FALSE, message=FALSE, dpi=300}
VlnPlot(object = merge_seurat_objects_CMS2_2, features = "harmony_1", 
              group.by = "orig.ident", pt.size = .1)
```

```{r, warning=FALSE, message=FALSE, dpi=300}
# palette <- RColorBrewer::brewer.pal(n=14, "Paired")
DimPlot(object = merge_seurat_objects_CMS2_2,  reduction = "harmony", 
  pt.size = .25, group.by = "slide")
```


```{r, warning=FALSE, message=FALSE, dpi=300}
VlnPlot(object = merge_seurat_objects_CMS2_2, features = "harmony_1", 
              group.by = "patient_ID", pt.size = .1)
```

```{r, warning=FALSE, message=FALSE, dpi=300}
# palette <- RColorBrewer::brewer.pal(n=14, "Paired")
DimPlot(object = merge_seurat_objects_CMS2_2,  reduction = "harmony", 
  pt.size = .25, group.by = "patient_ID")
```


```{r, warning=FALSE, message=FALSE, dpi=300}
VlnPlot(object = merge_seurat_objects_CMS2_2, features = "harmony_1", 
              group.by = "slide", pt.size = .1)
```

### UMAPs before batch correction


```{r, warning=FALSE, message=FALSE}
merge_seurat_objects_CMS2_2 <- merge_seurat_objects_CMS2_2 %>% 
    RunUMAP(reduction = "harmony", dims = 1:30) 
```

```{r, warning=FALSE, message=FALSE, dpi=300}
DimPlot(merge_seurat_objects_CMS2_2, reduction = "umap", group.by = "orig.ident", 
        pt.size = .25) 
```

```{r, warning=FALSE, message=FALSE, dpi=300}
DimPlot(merge_seurat_objects_CMS2_2, reduction = "umap", group.by = "slide", 
        pt.size = .1) 
```

```{r, warning=FALSE, message=FALSE, dpi=300, fig.height=8, fig.width=10}
palette <- RColorBrewer::brewer.pal(n=4,"Dark2")
DimPlot(merge_seurat_objects_CMS2_2, reduction = "umap", group.by = "patient_ID", 
        pt.size = 1.5, cols = palette, split.by = "rep") + 
  theme(axis.ticks = element_blank(), axis.text = element_blank(), 
        legend.title = element_blank(), legend.text = element_text(size=14), 
        legend.position = "bottom", legend.justification = "center") + 
  scale_shape_manual(values=seq(0,6)) + 
  # scale_size_manual(values=rep(10,7)) + 
  labs(title = NULL) + 
  guides(col =guide_legend(order =1, override.aes = list(size = 7), nrow = 1))
```

## UMAPS including Pathway, gene and TF information


We compute pathway and TF activity in the merged object

```{r, warning=FALSE, message=FALSE}
merge_seurat_objects_CMS2_2 <- 
  progeny::progeny(merge_seurat_objects_CMS2_2,
    scale=FALSE, organism="Human", top=500, perm=1,
    return_assay = TRUE, assay_name = "SCT")  

confidence_levels <- c("A","B","C")

dorothea_regulon <- 
        get(data("dorothea_hs", package = "dorothea"))  

regulon_filtered <- dorothea_regulon %>%
      dplyr::filter(confidence %in% confidence_levels)

merge_seurat_objects_CMS2_2 <- 
          run_viper(merge_seurat_objects_CMS2_2, regulon_filtered, 
          assay_key = "SCT", 
                  options = list(
                    method = "scale", minsize = 4, 
                    eset.filter = FALSE, cores = 1, verbose = FALSE))
```


```{r, warning=FALSE, message=FALSE}
umap_p1 <- 
  DimPlot(merge_seurat_objects_CMS2_2, reduction = "umap", group.by = "patient_ID", 
        pt.size = 1.5, cols = palette) + 
  theme(axis.ticks = element_blank(), axis.text = element_blank(), 
        legend.title = element_blank(), legend.text = element_text(size=11), 
        legend.position = "right", legend.justification = "center") + 
  scale_shape_manual(values=seq(0,6)) + 
  # scale_size_manual(values=rep(10,7)) + 
  labs(title = "Tumor Spots") + 
  guides(col =guide_legend(order =1, override.aes = list(size = 4)))

DefaultAssay(object = merge_seurat_objects_CMS2_2) <- "SCT"


umap_p2 <- 
  FeaturePlot(merge_seurat_objects_CMS2_2, reduction = "umap", features = "NUPR1", 
        pt.size = 1.5, cols =c("lightgrey", "#67000D")) + 
  theme(axis.ticks = element_blank(), axis.text = element_blank(), 
        legend.text = element_text(size=14), 
        legend.position = "right", legend.justification = "center") +
  labs(title = "NUPR1 Expression")  

DefaultAssay(object = merge_seurat_objects_CMS2_2) <- "progeny"

## Between EGFR and MAPK decide. 
umap_p3 <- 
  FeaturePlot(merge_seurat_objects_CMS2_2, reduction = "umap", features = "EGFR", 
        pt.size = 1.5, cols =c("lightgrey", "#67000D")) + 
  theme(axis.ticks = element_blank(), axis.text = element_blank(), 
        legend.text = element_text(size=14), 
        legend.position = "right", legend.justification = "center") +
  labs(title = "EGFR Pathway Activity")  + 
  scale_color_gradient2(low = "darkblue", mid ="white", high = "#67000D")

DefaultAssay(object = merge_seurat_objects_CMS2_2) <- "dorothea"



umap_p4 <- 
  FeaturePlot(merge_seurat_objects_CMS2_2, reduction = "umap", features = "FOXM1", 
        pt.size = 1.5, cols =c("blue", "#67000D")) + 
  theme(axis.ticks = element_blank(), axis.text = element_blank(), 
        legend.text = element_text(size=14), 
        legend.position = "right", legend.justification = "center") +
  labs(title = "FOXM1 TF Activity")  + 
  scale_color_gradient2(low = "darkblue", mid ="white", high = "#67000D")
```

```{r, warning=FALSE, message=FALSE, , dpi=300, fig.height=12, fig.width=12}
umap_p1 + umap_p2 + umap_p3 + umap_p4 + plot_layout(ncol = 2)
```


# Conclusion 


# Session Info Details

```{r, echo=FALSE, eval=TRUE}
sessionInfo()
```


