---
title: "Analysis of CRC Spatial Transcriptomics data from 10X: Proportion Plots"
author: 'Alberto Valdeolivas Urbelz (PS-Biomics&Path)'
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    code_folding: hide
    self_contained: true
    number_sections: yes
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: spacelab
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
params:
  data_directory : "/projects/site/pred/SpatialOmics/Experiment_CRC_AllSamples/"
  analysis_name : "analysis_alberto/"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(vctrs)
library(patchwork, lib.loc = "/apps/rocs/2020.08/cascadelake/software/R/4.1.2-foss-2020a/lib64/R/library")
library(Seurat)
# library(SeuratObject)
library(readr)
library(stringr)
library(purrr)
library(cowplot)
library(ggpubr)
library(kableExtra)


data_directory <-  params$data_directory
analysis_name <- params$analysis_name

input_folder <- "IntermediaryFiles/"
input_names_1 <- "SeuratList_Clusters_Res05.rds"
input_names_2 <- "Patho_Annotations/"

files_to_read_1 <- paste0(data_directory, analysis_name,input_folder,input_names_1)
files_to_read_2 <- paste0(data_directory, analysis_name,input_folder,input_names_2)

df_patient_ID_transform <- data.frame(
  patient = c("A120838","A121573","A416371","A551763","A595688","A798015",
              "A938797"), 
  patient_ID = c("S4_Col_Sig","S5_Rec","S3_Col_R","S1_Cec","S2_Col_R",
                 "S7_Rec/Sig","S6_Rec"))

samples_to_consider <- c("SN048_A121573_Rep1", "SN048_A121573_Rep2",
  "SN123_A595688_Rep1","SN124_A595688_Rep2", "SN123_A938797_Rep1_X", 
  "SN124_A938797_Rep2", "SN84_A120838_Rep1", "SN84_A120838_Rep2")
```

We include pathologist annotations into Seurat objects

```{r, warning=FALSE, message=FALSE}
seurat_objects <- readRDS(files_to_read_1)
seurat_objects_CMS2 <- seurat_objects[samples_to_consider]

## Attention: in one sample annotated as "tumor&stroma IC med to high"
## instead of "tumor&stroma IC med to high"
tumor_annotations_all <- 
  c("tumor", "tumor&stroma_IC med to high", "tumor&stroma_IC low", 
    "tumor&stroma IC med to high")
tumor_annotations <- 
  c("tumor")

seurat_objects_CMS2_subset <- 
  lapply(seurat_objects_CMS2, function(x){
    current_sample <- unique(x@meta.data$orig.ident)
    
    patho_anno_current_sample <- 
      read_csv(file = paste0(files_to_read_2, "Pathologist_Annotations_", 
        current_sample, ".csv"))
    
    colnames(patho_anno_current_sample) <- c("Barcode", "Pathologist_Annotations")
    
    x@meta.data <- x@meta.data %>%
      tibble::rownames_to_column("spot_id") %>%
      dplyr::left_join(patho_anno_current_sample, by = c("spot_id" = "Barcode")) %>%
      dplyr::mutate(Pathologist_Annotations = str_replace(Pathologist_Annotations, 
        "tumor&stroma IC med to high", "tumor&stroma_IC med to high")) %>% 
      tibble::column_to_rownames("spot_id") 
      
    
    Idents(x) <- 'Pathologist_Annotations'
    x <- subset(x, idents = intersect(levels(Idents(x)), tumor_annotations))
    
    return(x)
})


# We generate a file to export for the microbiome part.
patho_anno_current_sample_all <- data.frame()
for (current_sample in samples_to_consider){
  
    patho_anno_current_sample <- 
      read_csv(file = paste0(files_to_read_2, "Pathologist_Annotations_", 
        current_sample, ".csv"))
    
    colnames(patho_anno_current_sample) <- c("Barcode", "Pathologist_Annotations")
    
    patho_anno_current_sample$sample <- current_sample
    
    patho_anno_current_sample_all <- 
      rbind(patho_anno_current_sample_all, patho_anno_current_sample)
}

patho_anno_current_sample_all_toexport <- 
  patho_anno_current_sample_all %>% 
  dplyr::mutate(patient = str_match(sample, "_\\s*(.*?)\\s*_")[,2]) %>% 
  dplyr::mutate(sample = str_replace(sample, "SN123_A938797_Rep1_X", "SN123_A938797_Rep1")) %>% 
  dplyr::mutate(replicate = str_remove(sample, "[^A]+")) %>%
  dplyr::mutate(rep = str_remove(replicate, "[^R]+")) %>% 
  dplyr::left_join(df_patient_ID_transform) %>% 
  dplyr::mutate(patient_ID = paste0(patient_ID, "_", rep))

saveRDS(patho_anno_current_sample_all_toexport, 
  paste0(data_directory, analysis_name, input_folder, "Microbiome/tumor_spots.rds"))
```

We make the frequency/proportion plots with the annotations. 


```{r, warning=FALSE, message=FALSE}
seurat_objects_CMS2_merge <- 
  merge(x= seurat_objects_CMS2_subset[[1]],y=seurat_objects_CMS2_subset[-1])

df_plot_Anno <- 
  seurat_objects_CMS2_merge@meta.data %>% 
  dplyr::select(orig.ident, Pathologist_Annotations) %>% 
  dplyr::group_by(orig.ident) %>% 
  dplyr::count(Pathologist_Annotations) %>% 
  dplyr::mutate(patient = str_match(orig.ident, "_\\s*(.*?)\\s*_")[,2]) %>% 
  dplyr::mutate(replicate = str_remove(orig.ident, "[^A]+")) %>%
  dplyr::mutate(replicate = str_replace(replicate, "A938797_Rep1_X", "A938797_Rep1")) %>% 
  dplyr::mutate(rep = str_remove(replicate, "[^R]+")) %>% 
  dplyr::left_join(df_patient_ID_transform) %>% 
  dplyr::mutate(patient_ID = paste0(patient_ID, "_", rep))


df_plot_Anno$patient_ID <- 
  factor(df_plot_Anno$patient_ID , 
         levels = sort(unique(df_plot_Anno$patient_ID), decreasing = TRUE))

plot_annotations <- 
  ggplot(df_plot_Anno,aes(x = patient_ID , y= n , fill = Pathologist_Annotations)) +   
  geom_bar(stat="identity", position = "fill", width = 0.75, col ="Black") + 
  coord_flip() + 
  theme_minimal() + 
  scale_fill_brewer(palette = "Set1") + 
  theme( # remove the vertical grid lines
    panel.grid.major = element_blank(), 
    axis.title = element_blank(), 
    axis.text.x = element_text(size=12, face = "bold", family="Arial", angle = 90, vjust = -0.1),
    axis.text.y = element_text(size=12, face = "bold", family="Arial"),
    plot.title =  element_text(face="bold", size=14, hjust = 0.5)) + 
  labs(title ="CMS2 Annotations \n\ Proportions", x ="Sample") + 
  theme(legend.position ="bottom",  
        axis.title = element_blank(), 
        legend.text = element_text(size=12, face="bold", family="Arial"), 
        legend.title = element_blank()) + 
  scale_y_continuous(position = "left", breaks = c(0,0.5,1)) + 
  guides(fill=guide_legend(label.position ="left", ncol = 2 , reverse = TRUE))


df_plot_Anno_totals_CMS2 <- 
  df_plot_Anno %>% dplyr::group_by(patient_ID) %>% 
  dplyr::summarise(total_spots = sum(n ))

plot_totals_CMS2 <- df_plot_Anno_totals_CMS2 %>% 
  ggplot(aes(x=total_spots, y=patient_ID      )) + geom_point(size=3, col="black") +
  scale_x_continuous(trans = "log2", position = "top") + 
  theme(axis.title.y = element_blank(), # axis.text.y = element_blank(), 
        axis.ticks = element_blank(), panel.border = element_blank(),
        panel.grid.minor = element_blank(), panel.grid.major.y =  element_blank(),
        panel.grid.major.x = element_line(size=1, linetype=2, colour="Darkred"), 
        axis.title.x = element_text(size=14, face = "bold", family="Arial"), 
        axis.text.x = element_text(size=12, face = "bold", family="Arial")) + 
  labs(x="Spots Annotated \n\ as Tumor") 
```

Then, we read cell2location results, to explore the cell type content/abundance 
of this tumor annotated spots

```{r, warning=FALSE, message=FALSE}
filename_results_C2L_Korean <- paste0(data_directory, analysis_name, "Cell2Location/results/LocationModelLinearDependentWMultiExperiment_14experiments_36clusters_20654locations_4188genes/W_cell_density_q05.csv")
  

results_C2L_Korean <- read_csv(filename_results_C2L_Korean) %>%
  dplyr::mutate(spot_id = str_remove(.$spot_id, pattern = "Count_")) %>%
  dplyr::mutate(spot_id = str_replace(.$spot_id, pattern = "SN123_A938797_Rep1_" ,
                                      replacement = "SN123_A938797_Rep1_X_")) 

colnames(results_C2L_Korean) <-
  str_replace(colnames(results_C2L_Korean), 
              pattern = "q05_spot_factors", replacement = "")

results_C2L_Korean <- results_C2L_Korean %>% 
  dplyr::mutate(sample = str_extract(.$spot_id, pattern =".*Rep[1-2]")) %>%
  dplyr::mutate(spot_id = str_remove(.$spot_id, pattern =".*Rep[1-2]_")) %>%
  dplyr::mutate(spot_id = str_remove(spot_id, pattern = "^X_")) %>%
  tidyr::pivot_longer(!c(spot_id,sample), names_to = "Cell_subtype", values_to = "cell_density_q05") 
```

We then match the spots annotated as tumors with the results of cell2location

```{r, warning=FALSE, message=FALSE}
spots_tumor_annotations <- seurat_objects_CMS2_merge@meta.data %>% 
  dplyr::select(orig.ident, Pathologist_Annotations) %>% 
  tibble::rownames_to_column(var = "spot_id") %>%
  dplyr::mutate(spot_id = str_remove(spot_id, "_[0-9]")) %>% 
  dplyr::mutate(orig.ident = str_replace(orig.ident, "SN123_A938797_Rep1_X", "SN123_A938797_Rep1"))


df_C2L_PathoAnno_CMS2_content <- results_C2L_Korean %>%
  dplyr::inner_join(spots_tumor_annotations, by = c("spot_id" = "spot_id", "sample" = "orig.ident")) %>%
  # dplyr::filter(Cell_subtype != "CMS2") %>% 
  dplyr::mutate(Cell_subtype = ifelse(Cell_subtype == "CMS2", "CMS2", "Other")) %>% 
  dplyr::select(sample, cell_density_q05, Cell_subtype) %>%
  dplyr::group_by(sample, Cell_subtype) %>%
  dplyr::summarise_all(dplyr::funs(sum)) %>% 
  dplyr::mutate(patient = str_match(sample , "_\\s*(.*?)\\s*_")[,2]) %>% 
  dplyr::mutate(replicate = str_remove(sample , "[^A]+")) %>%
  dplyr::mutate(rep = str_remove(replicate, "[^R]+")) %>% 
  dplyr::left_join(df_patient_ID_transform) %>% 
  dplyr::mutate(patient_ID = paste0(patient_ID, "_", rep)) 


# %>% 
#  dplyr::group_by(patient_ID) %>% 
# dplyr::slice_max(order_by=cell_density_q05, n=nr_celltypes_to_include) %>% 
# dplyr::ungroup()


df_C2L_PathoAnno_CMS2_content$patient_ID  <- 
  factor(df_C2L_PathoAnno_CMS2_content$patient_ID , 
         levels = sort(unique(df_C2L_PathoAnno_CMS2_content$patient_ID), decreasing = TRUE))

plot_inner_frequency_CMS2_Others <- 
  ggplot(df_C2L_PathoAnno_CMS2_content,aes(x = patient_ID  , y= cell_density_q05, fill = Cell_subtype)) + 
  geom_bar(stat="identity", position = "fill", width = 0.75, col ="Black") + 
  coord_flip() + 
  theme_minimal() + 
  scale_fill_brewer(palette = "Set1") + 
  theme( # remove the vertical grid lines
    panel.grid.major = element_blank(), 
    axis.title = element_blank(), 
    axis.text.x = element_text(size=12, face = "bold", family="Arial", angle = 90, vjust = -0.1),
    axis.text.y = element_text(size=12, face = "bold", family="Arial"),
    plot.title =  element_text(face="bold", size=14, hjust = 0.5)) + 
  labs(title ="CMS2 Proportions \n\ in Tumors ", x ="Sample") + 
  theme(legend.position ="bottom",  
        axis.title = element_blank(), 
        legend.text = element_text(size=12, face="bold", family="Arial"), 
        legend.title = element_blank()) + 
  scale_y_continuous(position = "left", breaks = c(0,0.5,1)) + 
  guides(fill=guide_legend(label.position ="left", ncol = 2))
```

```{r, warning=FALSE, message=FALSE, dpi=300, fig.height=6, fig.width=6}
plot_inner_frequency_CMS2_Others
```

```{r, warning=FALSE, message=FALSE}
df_C2L_PathoAnno_CMS2_content %>% 
  dplyr::group_by(patient_ID) %>% 
  dplyr::mutate(proportions = cell_density_q05/sum(cell_density_q05)) %>% 
  dplyr::select(patient_ID, Cell_subtype, proportions) %>% 
  dplyr::arrange(desc(patient_ID), desc(proportions)) %>% 
  kbl() %>% kable_styling()
```


## Check the Neighborhood of the tumor{.tabset}

I check the neighborhood of all the spots annotated as tumor. 

```{r, warning=FALSE, message=FALSE}
treshold_dist <- 2 
Spots_CMS2_neighborhood <- 
  lapply(seurat_objects_CMS2, function(x){
    
    current_sample <- unique(x@meta.data$orig.ident)
    
    geometry <- GetTissueCoordinates(x, cols = c("row", "col"), scale = NULL)
    
    dist_matrix <- dist(geometry, method = "euclidean") %>% as.matrix()
    
    patho_anno_current_sample <- 
      read_csv(file = paste0(files_to_read_2, "Pathologist_Annotations_", 
        current_sample, ".csv"))
    
    colnames(patho_anno_current_sample) <- c("Barcode", "Pathologist_Annotations")
    
    all_spots <- rownames(geometry)
    
    tumor_spots <- 
      dplyr::filter(patho_anno_current_sample, Pathologist_Annotations  %in% tumor_annotations) %>% 
      dplyr::pull(Barcode)  %>% intersect(all_spots)
    
    
    neighbors_spots_list <-  tumor_spots %>% 
      map(function(y) {
        names(which(dist_matrix[all_spots, y] <= treshold_dist))
      })
    
    neighbors_spots <- setdiff(unique(unlist(neighbors_spots_list)),tumor_spots )
    
    return(neighbors_spots)
})
```

```{r, warning=FALSE, message=FALSE, results='asis'}
for (current_sample in names(seurat_objects_CMS2)){
  
  cat("### Current Sample: ", current_sample, "\n\n")
  
  patho_anno_current_sample <- 
      read_csv(file = paste0(files_to_read_2, "Pathologist_Annotations_", 
        current_sample, ".csv"))
  
  colnames(patho_anno_current_sample) <- c("Barcode", "Pathologist_Annotations")

  
  tumor_spots <- 
      dplyr::filter(patho_anno_current_sample, Pathologist_Annotations  == "tumor") %>% 
      dplyr::pull(Barcode)
  
  print(SpatialPlot(seurat_objects_CMS2[[current_sample]], 
              image.alpha = 0.5, crop = TRUE, 
        cells.highlight = list(tumor_spots,Spots_CMS2_neighborhood[[current_sample]]), 
        pt.size.factor = 2.25,  stroke = 1, 
        facet.highlight = TRUE) + 
          NoLegend()) 
  
  
  cat("\n\n")
}
```



## Enrichment and depletion in the tumor neighborhood

```{r, warning=FALSE, message=FALSE}
neigh_spots_df <- data.frame(
  sample = rep(names(Spots_CMS2_neighborhood), 
               lapply(Spots_CMS2_neighborhood, length)), 
  spot_id = unlist(Spots_CMS2_neighborhood)) %>% 
  dplyr::mutate(sample = str_replace(sample, "A938797_Rep1_X", "A938797_Rep1"))

# We save the data for microbiome analysis
neigh_spots_df_toexport <- 
  neigh_spots_df %>% 
  dplyr::mutate(patient = str_match(sample, "_\\s*(.*?)\\s*_")[,2]) %>% 
  dplyr::mutate(sample = str_replace(sample, "SN123_A938797_Rep1_X", "SN123_A938797_Rep1")) %>% 
  dplyr::mutate(replicate = str_remove(sample, "[^A]+")) %>%
  dplyr::mutate(rep = str_remove(replicate, "[^R]+")) %>% 
  dplyr::left_join(df_patient_ID_transform) %>% 
  dplyr::mutate(patient_ID = paste0(patient_ID, "_", rep))

saveRDS(neigh_spots_df_toexport, 
  paste0(data_directory, analysis_name, input_folder, "Microbiome/TME_spots.rds"))
```


```{r, warning=FALSE, message=FALSE, eval=FALSE}
samples_to_consider <- c("SN048_A121573_Rep1", "SN048_A121573_Rep2",
  "SN123_A595688_Rep1","SN124_A595688_Rep2", "SN123_A938797_Rep1", 
  "SN124_A938797_Rep2", "SN84_A120838_Rep1", "SN84_A120838_Rep2")

df_C2L_Proportions_Allspots <- results_C2L_Korean %>%
  # dplyr::filter(Cell_subtype != "CMS2") %>% 
  dplyr::filter(Cell_subtype != "Unknown") %>%
  dplyr::filter(sample %in% samples_to_consider) %>% 
  dplyr::select(spot_id, sample, cell_density_q05, Cell_subtype) %>%
  dplyr::mutate(spot_sample_id = paste0(sample, "_", spot_id)) %>% 
  dplyr::group_by(spot_sample_id) %>% 
  dplyr::mutate(Proportions_perSpot = cell_density_q05/sum(cell_density_q05)) %>%
  dplyr::ungroup()
  
avg_proportion_TME <- results_C2L_Korean %>%
  dplyr::inner_join(neigh_spots_df, by = c("spot_id" = "spot_id", "sample" = "sample")) %>% 
  # dplyr::filter(Cell_subtype != "CMS2") %>%
  dplyr::filter(Cell_subtype != "Unknown") %>%
  dplyr::select(spot_id, sample, cell_density_q05, Cell_subtype) %>%
  dplyr::mutate(spot_sample_id = paste0(sample, "_", spot_id)) %>% 
  dplyr::group_by(spot_sample_id) %>% 
  dplyr::mutate(Proportions_perSpot = cell_density_q05/sum(cell_density_q05)) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(sample, Cell_subtype) %>% 
  dplyr::mutate(avg_proportion_cell_TME = mean(Proportions_perSpot)) %>%
  dplyr::distinct(sample, Cell_subtype, avg_proportion_cell_TME)


n_random <- 10000
df_all <- data.frame()
for (current_sample in samples_to_consider){
  
  df_sample <- df_C2L_Proportions_Allspots %>% 
      dplyr::filter(sample == current_sample)
  
  df_sample_final <- data.frame()
  
  for (i in seq(n_random)){
    
     df_sample_random <- df_sample %>% 
        dplyr::mutate(rand_proportions = base::sample(Proportions_perSpot)) %>%
        dplyr::inner_join(neigh_spots_df, by = c("spot_id" = "spot_id", "sample" = "sample")) %>%
        dplyr::group_by(Cell_subtype) %>% 
        dplyr::mutate(avg_proportion_random = mean(rand_proportions)) %>%
        dplyr::distinct(sample, Cell_subtype, avg_proportion_random)
  
     df_sample_random_real <- df_sample_random %>% 
       dplyr::inner_join(avg_proportion_TME) %>%
       dplyr::mutate(difference_real_random = avg_proportion_cell_TME - avg_proportion_random) %>%
       dplyr::select(sample, Cell_subtype, difference_real_random)
     
     df_sample_final <- rbind(df_sample_final, df_sample_random_real)
     
  }  
  
  df_to_add <-df_sample_final %>% dplyr::group_by(Cell_subtype) %>% 
    dplyr::mutate(NES = mean(difference_real_random)/sd(difference_real_random)) %>% 
    dplyr::distinct(sample, Cell_subtype, NES)
  
  df_all <- rbind(df_all, df_to_add)
  
}
saveRDS(df_all, 
  paste0(data_directory, analysis_name, input_folder, "df_all_random.rds"))
```

```{r, warning=FALSE, message=FALSE}
df_all <- 
  readRDS(paste0(data_directory, analysis_name, input_folder, "df_all_random.rds"))

## Which cell types are we going to plot: 
# cells_to_visualize <- 
#   final_df_avg %>% 
#  dplyr::arrange(desc(abs(NES))) %>% 
#  dplyr::distinct(Cell_subtype) %>% 
#   dplyr::pull(Cell_subtype) %>% head(n=10)

final_df_avg_toplot<- df_all %>% 
  # dplyr::filter(Cell_subtype %in% cells_to_visualize) %>% 
  dplyr::mutate(patient = str_match(sample, "_\\s*(.*?)\\s*_")[,2]) %>% 
  dplyr::mutate(replicate = str_remove(sample, "[^A]+")) %>%
  dplyr::mutate(rep = str_remove(replicate, "[^R]+")) %>% 
  dplyr::left_join(df_patient_ID_transform) %>% 
  dplyr::mutate(patient_ID = paste0(patient_ID, "_", rep)) %>% 
  dplyr::mutate(Direction = ifelse(NES >= 0, "Enrichment", "Depletion"))


final_df_avg_toplot$patient_ID  <- 
  factor(final_df_avg_toplot$patient_ID , 
         levels = sort(unique(final_df_avg_toplot$patient_ID), decreasing = TRUE))

  
enrichment_depletion_plot <- 
  ggplot(final_df_avg_toplot, aes(x = patient_ID, 
    y = Cell_subtype, colour = Direction, size = abs(NES))) + 
    geom_point() + 
    scale_size_continuous(range = c(1.5, 9.5), breaks = c(5,35), labels = c("Small","Large")) + theme_bw() + 
    theme(axis.title = element_blank(), 
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          axis.text.x = element_text(size=12, face = "bold", family="Arial", angle = 90, hjust = 1),
          axis.text.y = element_text(size=12, face = "bold", family="Arial"),
          plot.title =  element_text(face="bold", size=14, hjust = 0.5) , 
          legend.position ="right", legend.text = element_text(size=10, face="bold", family="Arial"),
          legend.title =  element_text(size=12, face="bold", family="Arial")) +
          scale_color_manual(values = c( "#377EB8", "#E41A1C")) + 
          coord_flip() + 
          guides(col = guide_legend(order = 1, override.aes = list(size = 5)),  
          size =guide_legend(title = "Size Effect", order =2, override.aes = list(size = c(3.5,7.5)))) +
          labs(title ="Enrichment \n\ Depletion") 
```


```{r, warning=FALSE, message=FALSE, dpi=300, fig.height=8, fig.width=10}
enrichment_depletion_plot
```


## Which are the main other tumor populations in the inner tumor? 


```{r, warning=FALSE, message=FALSE, eval=TRUE}
df_C2L_proportions <- results_C2L_Korean %>%
  dplyr::inner_join(spots_tumor_annotations, by = c("spot_id" = "spot_id", "sample" = "orig.ident")) %>%
  dplyr::filter(Cell_subtype != "CMS2") %>% 
  dplyr::filter(Cell_subtype != "Unknown") %>% 
  dplyr::select(sample, cell_density_q05, Cell_subtype, spot_id) %>% 
  dplyr::mutate(spot_sample_id = paste0(sample, "_", spot_id)) %>% 
  dplyr::group_by(spot_sample_id) %>% 
  dplyr::mutate(Proportions_perSpot = cell_density_q05/sum(cell_density_q05)) %>%
  dplyr::ungroup() %>% 
  dplyr::mutate(patient = str_match(sample , "_\\s*(.*?)\\s*_")[,2])

df_all_kruskal_results <- data.frame()

for (current_celltype in unique(df_C2L_proportions$Cell_subtype)){
  current_df <- 
    df_C2L_proportions %>% dplyr::filter(Cell_subtype == current_celltype)
  current_kruskal_results <- 
    kruskal.test(Proportions_perSpot ~ patient, data = current_df)
  # boxplot(Proportions_perSpot ~ patient, data = current_df)
  current_kruskal_df <- 
    data.frame( cell_type = current_celltype,
      statistic = current_kruskal_results$statistic, 
      p.value = current_kruskal_results$p.value)                                

  df_all_kruskal_results <- rbind(df_all_kruskal_results, current_kruskal_df)
}

df_C2L_proportions_patient_topCells <- df_C2L_proportions %>% 
  dplyr::group_by(patient, Cell_subtype) %>% 
  dplyr::mutate(avg_patient_celltype = mean(cell_density_q05)) %>% 
  dplyr::select(patient, Cell_subtype, avg_patient_celltype) %>% 
  dplyr::distinct() %>% dplyr::group_by(patient) %>% 
  dplyr::slice_max(avg_patient_celltype, n = 12) %>% 
  dplyr::pull(Cell_subtype) %>% unique()
  
kruskal_results_top_cells <- 
  df_all_kruskal_results %>% dplyr::slice_max(statistic, n=20) %>% 
  dplyr::pull(cell_type)

cell_to_visualize <- intersect(df_C2L_proportions_patient_topCells,kruskal_results_top_cells)

df_C2L_PathoAnno_inner_content <- results_C2L_Korean %>%
  dplyr::inner_join(spots_tumor_annotations, by = c("spot_id" = "spot_id", "sample" = "orig.ident")) %>%
  dplyr::filter(Cell_subtype != "CMS2") %>% 
  dplyr::filter(Cell_subtype != "Unkown") %>% 
  dplyr::select(sample, cell_density_q05, Cell_subtype) %>%
  dplyr::group_by(sample, Cell_subtype) %>%
  dplyr::summarise_all(dplyr::funs(sum)) %>% 
  dplyr::mutate(patient = str_match(sample , "_\\s*(.*?)\\s*_")[,2]) %>% 
  dplyr::mutate(replicate = str_remove(sample , "[^A]+")) %>%
  dplyr::mutate(rep = str_remove(replicate, "[^R]+")) %>% 
  dplyr::left_join(df_patient_ID_transform) %>% 
  dplyr::mutate(patient_ID = paste0(patient_ID, "_", rep))   %>%
  dplyr::filter(Cell_subtype %in% cell_to_visualize)


df_C2L_PathoAnno_inner_content$patient_ID  <- 
  factor(df_C2L_PathoAnno_inner_content$patient_ID , 
         levels = sort(unique(df_C2L_PathoAnno_inner_content$patient_ID), decreasing = TRUE))

plot_inner_frequency <- 
  ggplot(df_C2L_PathoAnno_inner_content,aes(x = patient_ID  , y= cell_density_q05, fill = Cell_subtype)) + 
  geom_bar(stat="identity", position = "fill", width = 0.75, col ="Black") + 
  coord_flip() + 
  theme_minimal() + 
  scale_fill_brewer(palette = "Set1") + 
  theme( # remove the vertical grid lines
    panel.grid.major = element_blank(), 
    axis.title = element_blank(), 
    axis.text.x = element_text(size=12, face = "bold", family="Arial", angle = 90, vjust = -0.1),
    axis.text.y = element_text(size=12, face = "bold", family="Arial"),
    plot.title =  element_text(face="bold", size=14, hjust = 0.5)) + 
  labs(title ="Other Cell Types \n\ Proportions in Tumors", x ="Sample") + 
  theme(legend.position ="bottom",  
        axis.title = element_blank(), 
        legend.text = element_text(size=12, face="bold", family="Arial"), 
        legend.title = element_blank()) + 
  scale_y_continuous(position = "left", breaks = c(0,0.5,1)) + 
  guides(fill=guide_legend(label.position ="left", ncol = 2))
```


```{r, warning=FALSE, message=FALSE, dpi=300, fig.height=6, fig.width=6, eval=TRUE}
plot_inner_frequency 
```


```{r, warning=FALSE, message=FALSE, dpi=300, fig.height=6, fig.width=8, eval=TRUE}
plot_inner_frequency_CMS2_Others + plot_inner_frequency + 
  theme(axis.text.y = element_blank())
```


## Neighborhood proportions


```{r, warning=FALSE, message=FALSE, eval=TRUE}
# cells_to_consider <- cells_to_visualize
df_C2L_PathoAnno_neigh_content_proportions <- results_C2L_Korean %>%
  dplyr::inner_join(neigh_spots_df, by = c("spot_id" = "spot_id", "sample" = "sample")) %>%
  dplyr::filter(Cell_subtype != "CMS2") %>% 
  dplyr::filter(Cell_subtype != "Unknown") %>% 
  dplyr::select(sample, cell_density_q05, Cell_subtype, spot_id) %>% 
  dplyr::mutate(spot_sample_id = paste0(sample, "_", spot_id)) %>% 
  dplyr::group_by(spot_sample_id) %>% 
  dplyr::mutate(Proportions_perSpot = cell_density_q05/sum(cell_density_q05)) %>%
  dplyr::ungroup() %>% 
  dplyr::mutate(patient = str_match(sample , "_\\s*(.*?)\\s*_")[,2]) %>%
  dplyr::inner_join(df_patient_ID_transform, by ="patient")
  

df_all_kruskal_results <- data.frame()

for (current_celltype in unique(df_C2L_proportions$Cell_subtype)){
  current_df <- 
    df_C2L_PathoAnno_neigh_content_proportions %>% 
    dplyr::filter(Cell_subtype == current_celltype)
  current_kruskal_results <- 
    kruskal.test(Proportions_perSpot ~ patient_ID, data = current_df)
  # boxplot(Proportions_perSpot ~ patient, data = current_df)
  current_kruskal_df <- 
    data.frame( cell_type = current_celltype,
      statistic = current_kruskal_results$statistic, 
      p.value = current_kruskal_results$p.value)   
  
  
    current_plot <- 
      ggplot(current_df, aes(x=patient_ID, y=Proportions_perSpot, fill=patient_ID)) + 
      geom_boxplot()+
      theme_bw() + 
      scale_fill_brewer(palette = "Set1") + 
      theme(panel.grid.major = element_blank(), 
        axis.title = element_blank(), 
        axis.text.x = element_text(size=12, face = "bold", family="Arial", angle = 90, vjust = -0.1),
        axis.text.y = element_text(size=12, face = "bold", family="Arial"),
        plot.title =  element_text(face="bold", size=14, hjust = 0.5)) + 
        labs(title=current_celltype,x="", y = "Proportions per Spot") +
        theme(legend.position ="bottom",  
        axis.title = element_blank(), 
        legend.text = element_text(size=12, face="bold", family="Arial"), 
        legend.title = element_blank())
    
    print(current_plot)
  
  
  if (current_celltype  == "Myofibroblasts"){
    kruskal_myofibroblasts <- 
      ggplot(current_df, aes(x=patient_ID, y=Proportions_perSpot, fill=patient_ID)) + 
      geom_boxplot()+
      theme_bw() + 
      scale_fill_brewer(palette = "Set1") + 
      theme(panel.grid.major = element_blank(), 
        axis.title.x = element_blank(), 
        axis.title.y = element_text(size=14, face = "bold", family="Arial"),
        axis.text.x = element_text(size=12, face = "bold", family="Arial", angle = 90, vjust = -0.1),
        axis.text.y = element_text(size=12, face = "bold", family="Arial"),
        plot.title =  element_text(face="bold", size=14, hjust = 0.5)) + 
        labs(title="Myofibroblasts",x="", y = "Proportions per Spot", fill = "Patient") +
        theme(legend.position ="bottom",  
        legend.text = element_text(size=12, face="bold", family="Arial"), 
        legend.title = element_text(size=14, face="bold", family="Arial")) + 
        geom_text(x=2.5, y=0.65, label= paste0("Pvalue = ", signif(current_kruskal_df$p.value,digits=3)), 
                  size = 4.5, family = "Arial",  colour = "darkblue") + 
      guides(fill = guide_legend(nrow = 2))
    
  }
  
    
  if (current_celltype  == "Mature Enterocytes type 2"){
    kruskal_enterocytes2 <- 
      ggplot(current_df, aes(x=patient_ID, y=Proportions_perSpot, fill=patient_ID)) + 
      geom_boxplot()+
      theme_bw() + 
      scale_fill_brewer(palette = "Set1") + 
      theme(panel.grid.major = element_blank(), 
        axis.title.x = element_blank(), 
        axis.title.y = element_text(size=14, face = "bold", family="Arial"),
        axis.text.x = element_text(size=12, face = "bold", family="Arial", angle = 90, vjust = -0.1),
        axis.text.y = element_text(size=12, face = "bold", family="Arial"),
        plot.title =  element_text(face="bold", size=14, hjust = 0.5)) + 
        labs(title="Mature Enterocytes type 2",x="", y = "Proportions per Spot", fill = "Patient") +
        theme(legend.position ="bottom",  
        legend.text = element_text(size=12, face="bold", family="Arial"), 
        legend.title = element_text(size=14, face="bold", family="Arial")) + 
        geom_text(x=2.5, y=0.65, label= paste0("Pvalue = ", signif(current_kruskal_df$p.value,digits=3)), 
                  size = 4.5, family = "Arial",  colour = "darkblue")
    
  }  

  if (current_celltype  == "SPP1+"){
    kruskal_SPP1 <- 
      ggplot(current_df, aes(x=patient_ID, y=Proportions_perSpot, fill=patient_ID)) + 
      geom_boxplot()+
      theme_bw() + 
      scale_fill_brewer(palette = "Set1") + 
      theme(panel.grid.major = element_blank(), 
        axis.title.x = element_blank(), 
        axis.title.y = element_text(size=14, face = "bold", family="Arial"),
        axis.text.x = element_text(size=12, face = "bold", family="Arial", angle = 90, vjust = -0.1),
        axis.text.y = element_text(size=12, face = "bold", family="Arial"),
        plot.title =  element_text(face="bold", size=14, hjust = 0.5)) + 
        labs(title="SPP1+",x="", y = "Proportions per Spot", fill = "Patient") +
        theme(legend.position ="bottom",  
        legend.text = element_text(size=12, face="bold", family="Arial"), 
        legend.title = element_text(size=14, face="bold", family="Arial")) + 
        scale_y_continuous(limits = c(0, 0.25)) + 
        geom_text(x=2.5, y=0.25, label= paste0("Pvalue = ", signif(current_kruskal_df$p.value,digits=3)), 
                  size = 4.5, family = "Arial",  colour = "darkblue")
    
  }
  

  df_all_kruskal_results <- rbind(df_all_kruskal_results, current_kruskal_df)
}

df_C2L_proportions_patient_neigh_topCells <- 
  df_C2L_PathoAnno_neigh_content_proportions %>% 
  dplyr::group_by(patient, Cell_subtype) %>% 
  dplyr::mutate(avg_patient_celltype = mean(cell_density_q05)) %>% 
  dplyr::select(patient, Cell_subtype, avg_patient_celltype) %>% 
  dplyr::distinct() %>% dplyr::group_by(patient) %>% 
  dplyr::slice_max(avg_patient_celltype, n = 12) %>% 
  dplyr::pull(Cell_subtype) %>% unique()
  
kruskal_results_top_cells <- 
  df_all_kruskal_results %>% dplyr::slice_max(statistic, n=20) %>% 
  dplyr::pull(cell_type)

# cell_to_visualize <- 
#  intersect(df_C2L_proportions_patient_neigh_topCells,kruskal_results_top_cells)


df_C2L_PathoAnno_neigh_content <- results_C2L_Korean %>%
  dplyr::inner_join(neigh_spots_df, by = c("spot_id" = "spot_id", "sample" = "sample")) %>%
  dplyr::filter(Cell_subtype != "CMS2") %>% 
  dplyr::select(sample, cell_density_q05, Cell_subtype) %>%
  dplyr::group_by(sample, Cell_subtype) %>%
  dplyr::summarise_all(dplyr::funs(sum)) %>% 
  dplyr::mutate(patient = str_match(sample , "_\\s*(.*?)\\s*_")[,2]) %>% 
  dplyr::mutate(replicate = str_remove(sample , "[^A]+")) %>%
  dplyr::mutate(rep = str_remove(replicate, "[^R]+")) %>% 
  dplyr::left_join(df_patient_ID_transform) %>% 
  dplyr::mutate(patient_ID = paste0(patient_ID, "_", rep))   %>%
  dplyr::filter(Cell_subtype %in% cell_to_visualize)



df_C2L_PathoAnno_neigh_content$patient_ID  <- 
  factor(df_C2L_PathoAnno_neigh_content$patient_ID , 
         levels = sort(unique(df_C2L_PathoAnno_neigh_content$patient_ID), decreasing = TRUE))

plot_neigh_frequency <- 
  ggplot(df_C2L_PathoAnno_neigh_content,aes(x = patient_ID  , y= cell_density_q05, fill = Cell_subtype)) + 
  geom_bar(stat="identity", position = "fill", width = 0.75, col ="Black") + 
  coord_flip() + 
  theme_minimal() + 
  scale_fill_brewer(palette = "Set1") + 
  theme( # remove the vertical grid lines
    panel.grid.major = element_blank(), 
    axis.title = element_blank(), 
    axis.text.x = element_text(size=12, face = "bold", family="Arial", angle = 90, vjust = -0.1),
    axis.text.y = element_text(size=12, face = "bold", family="Arial"),
    plot.title =  element_text(face="bold", size=14, hjust = 0.5)) + 
  labs(title ="Cell Type Proportions \n\ in the tumor Neighboorhod", x ="Sample") + 
  theme(legend.position ="bottom",  
        axis.title = element_blank(), 
        legend.text = element_text(size=12, face="bold", family="Arial"), 
        legend.title = element_blank()) + 
  scale_y_continuous(position = "left", breaks = c(0,0.5,1)) + 
  guides(fill=guide_legend(label.position ="left", ncol = 2))


df_plot_Anno_totals_neigh <- results_C2L_Korean %>%
  dplyr::inner_join(neigh_spots_df, by = c("spot_id" = "spot_id", "sample" = "sample")) %>% 
  dplyr::mutate(patient = str_match(sample , "_\\s*(.*?)\\s*_")[,2]) %>% 
  dplyr::mutate(replicate = str_remove(sample , "[^A]+")) %>%
  dplyr::mutate(rep = str_remove(replicate, "[^R]+")) %>% 
  dplyr::left_join(df_patient_ID_transform) %>% 
  dplyr::mutate(patient_ID = paste0(patient_ID, "_", rep)) %>% 
  dplyr::count(patient_ID,Cell_subtype) %>% 
  dplyr::ungroup() %>% dplyr::distinct(patient_ID   , n)


df_plot_Anno_totals_neigh$patient_ID  <- 
  factor(df_plot_Anno_totals_neigh$patient_ID , 
         levels = sort(unique(df_plot_Anno_totals_neigh$patient_ID), decreasing = TRUE))

plot_totals_CMS2 <- df_plot_Anno_totals_neigh %>% 
  ggplot(aes(x=n, y=patient_ID      )) + geom_point(size=3, col="black") +
  scale_x_continuous(trans = "log2", position = "bottom", breaks = c(128,256,512)) + 
  theme(axis.title= element_blank(), axis.text.y = element_blank(), 
        axis.ticks = element_blank(), panel.border = element_blank(),
        panel.grid.minor = element_blank(), panel.grid.major.y =  element_blank(),
        panel.grid.major.x = element_line(size=1, linetype=2, colour="Darkred"), 
        plot.title = element_text(size=14, face = "bold", family="Arial", hjust = 0.5), 
        axis.text.x = element_text(size=12, face = "bold", family="Arial", angle = 90, hjust = 1, vjust = 1)) + 
  labs(title="# Spots in the \n\ Tumor Neighboorhod") 
```


```{r, warning=FALSE, message=FALSE, dpi=300, fig.height=6, fig.width=6, eval=TRUE}
frequency_legend <- get_legend(plot_neigh_frequency)
plot_neigh_frequency_nolegend <- plot_neigh_frequency + theme(legend.position = "none") 

plot_neigh_frequency_nolegend + plot_spacer() + plot_totals_CMS2 +
  ggpubr::as_ggplot(frequency_legend) + plot_layout(widths = c(5, -0.75, 2), heights = c(4.5,1))
```

```{r, warning=FALSE, message=FALSE, dpi=300, fig.height=6, fig.width=8}
patient_legend <- get_legend(kruskal_myofibroblasts)

kruskal_myofibroblasts_nolegend <- kruskal_myofibroblasts + theme(legend.position = "none") + 
  labs(tag = "a")
kruskal_enterocytes2_nolegend <- kruskal_enterocytes2 + theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +  
  labs(tag = "b")
kruskal_SPP1_nolegend <- kruskal_SPP1 + theme(legend.position = "none") + 
  theme(axis.title.y = element_blank()) +  
  labs(tag = "c")

kruskal_myofibroblasts_nolegend + plot_spacer() + kruskal_enterocytes2_nolegend + 
  plot_spacer() + kruskal_SPP1_nolegend + 
  plot_spacer() + plot_spacer() +  ggpubr::as_ggplot(patient_legend) + 
  plot_spacer() + plot_spacer() + 
  plot_layout(widths = c(3, 0.15, 3,0.15,3), heights = c(6,1))
```


# Session Info Details

```{r, echo=FALSE, eval=TRUE}
sessionInfo()
```