---
title: "Analysis of CRC Spatial Transcriptomics data from 10X: Pathway Activity using Progeny"
author: 'Alberto Valdeolivas Urbelz (PS-Biomics&Path)'
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    code_folding: hide
    self_contained: true
    number_sections: yes
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: spacelab
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
params:
  data_directory : "/projects/site/pred/SpatialOmics/Experiment_CRC_AllSamples/"
  analysis_name : "analysis_alberto/"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

The present script takes the the Spatial Omics data generated with 10X 
technology on a set of CRC samples and their associated liver sections with 
metastases and computes pathway activity using Progeny. The datasets are extracted
from the publication: **Spatiotemporal Immune Landscape of Colorectal Cancer Liver Metastasis at Single-Cell Level** [DOI: 10.1158/2159-8290.CD-21-0316](https://cancerdiscovery.aacrjournals.org/content/12/1/134.long)

```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(patchwork, lib.loc = "/apps/rocs/2020.08/cascadelake/software/R/4.1.2-foss-2020a/lib64/R/library")
library(Seurat)
# library(tidyverse)
library(tidyr)
library(stringr)
library(readr)
library(purrr)
library(progeny)
source(file = "../WrapperFunction/SeuratWrappers.R")

data_directory <-  params$data_directory
analysis_name <- params$analysis_name

input_files <- "ST_LiverMetastases_pub/IntermediaryFiles/"
input_names <- "SeuratList_Clusters_Res05.rds"
output_names <- "SeuratList_Clusters_Res05_Progeny.rds"

files_to_read <- paste0(data_directory, analysis_name,input_files,input_names)
results_directory <- paste0(data_directory, analysis_name, input_files)
# setwd(paste0(data_directory, analysis_name))
```

We read the Seurat objects files after Quality Control

```{r, warning=FALSE}
seurat_objects <- readRDS(files_to_read)
```

# Computing Pathway activity with Progeny

We compute Pathway activity using Progeny

```{r, message=FALSE, warning=FALSE}
seurat_objects <- get.progeny.scores(seurat_objects)
# pathway_plots <- get.pathway.plots(seurat_objects, pt.size.factor = 2.25)
```

# Results {.tabset}

We display the spatial plots showing pathway activity per spot and violing plots
showing pathway activity per cluster.

```{r, results = 'asis', message=FALSE, dpi=80, warning=FALSE, eval=FALSE}
for ( sample in names(seurat_objects)){
  
  cat( "## Sample : ", sample , "\n\n" )
  
  cat( "Showing sample : ", sample , "\n\n")
  
  pathways <- 
    unique(rownames(GetAssayData(seurat_objects[[sample]], assay = "progeny")))
  DefaultAssay(seurat_objects[[sample]]) <- "progeny"
  
  for (j in seq(from = 1,to= length(pathways), by=2)){
    current_pathways <- pathways[c(j,j+1)]
    print(SpatialFeaturePlot(seurat_objects[[sample]], 
      features = current_pathways, ncol = 2, alpha = 1, pt.size.factor = 2))  
    cat( "\n\n")
    print(VlnPlot(seurat_objects[[sample]], 
                  features = current_pathways, ncol=2))
    cat( "\n\n")
  }
}
```

# Publication plots

```{r, warning=FALSE, message=FALSE}
current_sample <- 'ST-colon1'
current_obj <- seurat_objects[[current_sample]]
DefaultAssay(current_obj) <- "progeny"



current_pathway <- c("WNT")
breaks_vec <- 
    round(seq(min(GetAssayData(current_obj)[current_pathway,]), 
              max(GetAssayData(current_obj)[current_pathway,]), length.out= 4),1)


ST_colon1_WNT <- 
    Seurat::SpatialFeaturePlot(
      object = current_obj,
      features = c(current_pathway), alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("WNT Pathway activity") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(angle = 90, face="bold", size=10),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 
```


```{r, warning=FALSE, message=FALSE}
ST_colon1_WNT
```

```{r, warning=FALSE, message=FALSE}
current_pathway <- c("MAPK")
breaks_vec <- 
    round(seq(min(GetAssayData(current_obj)[current_pathway,]), 
              max(GetAssayData(current_obj)[current_pathway,]), length.out= 4),1)


ST_colon1_MAPK <- 
    Seurat::SpatialFeaturePlot(
      object = current_obj,
      features = c(current_pathway), alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("MAPK Pathway activity") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(angle = 90, face="bold", size=10),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 

```

```{r, warning=FALSE, message=FALSE}
ST_colon1_MAPK
```


```{r, warning=FALSE, message=FALSE}
current_sample <- 'ST-liver1'
current_obj <- seurat_objects[[current_sample]]
DefaultAssay(current_obj) <- "progeny"


current_pathway <- c("WNT")
breaks_vec <- 
    round(seq(min(GetAssayData(current_obj)[current_pathway,]), 
              max(GetAssayData(current_obj)[current_pathway,]), length.out= 4),1)


ST_liver1_WNT <- 
    Seurat::SpatialFeaturePlot(
      object = current_obj,
      features = c(current_pathway), alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("WNT Pathway activity") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(angle = 90, face="bold", size=10),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 
```


```{r, warning=FALSE, message=FALSE}
ST_liver1_WNT
```

```{r, warning=FALSE, message=FALSE}
current_pathway <- c("MAPK")
breaks_vec <- 
    round(seq(min(GetAssayData(current_obj)[current_pathway,]), 
              max(GetAssayData(current_obj)[current_pathway,]), length.out= 4),1)


ST_liver1_MAPK <- 
    Seurat::SpatialFeaturePlot(
      object = current_obj,
      features = c(current_pathway), alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("MAPK Pathway activity") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(angle = 90, face="bold", size=10),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 

```

```{r, warning=FALSE, message=FALSE}
ST_liver1_MAPK
```


```{r, warning=FALSE, message=FALSE}
current_sample <- 'ST-liver2'
current_obj <- seurat_objects[[current_sample]]
DefaultAssay(current_obj) <- "progeny"


current_pathway <- c("VEGF")
breaks_vec <- 
    round(seq(min(GetAssayData(current_obj)[current_pathway,]), 
              max(GetAssayData(current_obj)[current_pathway,]), length.out= 4),1)


ST_liver2_VEGF <- 
    Seurat::SpatialFeaturePlot(
      object = current_obj,
      features = c(current_pathway), alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("VEGF Pathway activity") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(angle = 90, face="bold", size=10),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 
```


```{r, warning=FALSE, message=FALSE}
ST_liver2_VEGF
```




# Conclusions

# Saving the results

```{r, eval=FALSE}
saveRDS(seurat_objects, file = paste0(results_directory, output_names))
```

# Session Info Details

```{r, echo=FALSE, eval=TRUE}
sessionInfo()
```
