---
title: "Analysis of CRC Spatial Transcriptomics data from 10X: Misty aggregated results"
author: 'Alberto Valdeolivas Urbelz (PS-Biomics&Path)'
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    code_folding: hide
    self_contained: true
    number_sections: yes
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: spacelab
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
params:
  data_directory : "/projects/site/pred/SpatialOmics/Experiment_CRC_AllSamples/"
  analysis_name : "analysis_alberto/"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(vctrs)
library(patchwork,lib.loc = "/apps/rocs/2020.08/cascadelake/software/R/4.1.2-foss-2020a/lib64/R/library")
library(Seurat)
library(dplyr)
library(tibble)
library(readr)
library(stringr)
library(cowplot)
library(RColorBrewer)
library(mistyR)
library(purrr)
library(kableExtra)
library(pheatmap)


data_directory <- 
 params$data_directory
analysis_name <- 
  params$analysis_name


result_folders <- "ST_LiverMetastases_pub/IntermediaryFiles/results_TF_Ligands_DorotheaClusters"
input_files <- "ST_LiverMetastases_pub/IntermediaryFiles/"
input_names <- "SeuratList_Clusters_Res05_Dorothea.rds"

samples_to_consider <- c() 

results_misty_folders <- paste0(data_directory, analysis_name, result_folders)
files_to_read_seurat <- paste0(data_directory, analysis_name,input_files,input_names)
```

# Results

We check the results for different combinations of our data. 

```{r, warning=FALSE, message=FALSE}
misty_folders_all <- as.list(list.dirs(results_misty_folders, recursive = FALSE))
```

## Large CRC tumors in the colon 

```{r, warning=FALSE, message=FALSE}

misty_folders_selected <- misty_folders_all[c(1,2,3)]

misty_results <- collect_results(misty_folders_selected)
```

```{r, warning=FALSE, message=FALSE, dpi=300}
plot.data <- 
  misty_results$importances.aggregated %>%
  dplyr::filter(view == "juxta.ligands_2") %>% 
  dplyr::select(Predictor, Target, Importance ) %>% 
  tidyr::pivot_wider(names_from = Predictor,values_from = Importance ) %>% 
  tibble::column_to_rownames(var = 'Target') %>% 
  as.matrix() %>% t()

palette <-c(rev(brewer.pal(n = 5, name ="RdBu")), "#000000")
pheatmap(plot.data, treeheight_row=0,treeheight_col=0,  
         color =palette, breaks = c(min(plot.data),-1,-0.5,0,0.5,1,max(plot.data)))
```


## Non-treated CRC Tumors in the colon. 


```{r, warning=FALSE, message=FALSE}

misty_folders_selected <- misty_folders_all[c(1,2)]

misty_results <- collect_results(misty_folders_selected)
```

```{r, warning=FALSE, message=FALSE, dpi=300}
plot.data <- 
  misty_results$importances.aggregated %>%
  dplyr::filter(view == "juxta.ligands_2") %>% 
  dplyr::select(Predictor, Target, Importance ) %>% 
  tidyr::pivot_wider(names_from = Predictor,values_from = Importance ) %>% 
  tibble::column_to_rownames(var = 'Target') %>% 
  as.matrix() %>% t()

palette <-c(rev(brewer.pal(n = 5, name ="RdBu")), "#000000")
pheatmap(plot.data, treeheight_row=0,treeheight_col=0,  
         color =palette, breaks = c(min(plot.data),-1,-0.5,0,0.5,1,max(plot.data)))
```

## CRC Tumors treated or untreated with large tumor areas. 


```{r, warning=FALSE, message=FALSE}
misty_folders_selected <- misty_folders_all[c(1,3)]

misty_results <- collect_results(misty_folders_selected)
```

```{r, warning=FALSE, message=FALSE, dpi=300}
plot.data <- 
  misty_results$importances.aggregated %>%
  dplyr::filter(view == "juxta.ligands_2") %>% 
  dplyr::select(Predictor, Target, Importance ) %>% 
  tidyr::pivot_wider(names_from = Predictor,values_from = Importance ) %>% 
  tibble::column_to_rownames(var = 'Target') %>% 
  as.matrix() %>% t()

palette <-c(rev(brewer.pal(n = 5, name ="RdBu")), "#000000")
pheatmap(plot.data, treeheight_row=0,treeheight_col=0,  
         color =palette, breaks = c(min(plot.data),-1,-0.5,0,0.5,1,max(plot.data)))
```

## All CRC metastatic tumors in the liver 

```{r, warning=FALSE, message=FALSE}

misty_folders_selected <- misty_folders_all[c(4,5,6,7)]

misty_results <- collect_results(misty_folders_selected)

```
  
```{r, warning=FALSE, message=FALSE, dpi=300}
plot.data <- 
  misty_results$importances.aggregated %>%
  dplyr::filter(view == "juxta.ligands_2") %>% 
  dplyr::select(Predictor, Target, Importance ) %>% 
  tidyr::pivot_wider(names_from = Predictor,values_from = Importance ) %>% 
  tibble::column_to_rownames(var = 'Target') %>% 
  as.matrix() %>% t()

palette <-c(rev(brewer.pal(n = 5, name ="RdBu")), "#000000")
pheatmap(plot.data, treeheight_row=0,treeheight_col=0,  
         color =palette, breaks = c(min(plot.data),-1,-0.5,0,0.5,1,max(plot.data)))
```



##  CRC metastatic tumors in the liver non treated patients

```{r, warning=FALSE, message=FALSE}
misty_folders_selected <- misty_folders_all[c(4,5)]

misty_results <- collect_results(misty_folders_selected)

```

```{r, warning=FALSE, message=FALSE, dpi=300}
plot.data <- 
  misty_results$importances.aggregated %>%
  dplyr::filter(view == "juxta.ligands_2") %>% 
  dplyr::select(Predictor, Target, Importance ) %>% 
  tidyr::pivot_wider(names_from = Predictor,values_from = Importance ) %>% 
  tibble::column_to_rownames(var = 'Target') %>% 
  as.matrix() %>% t()

palette <-c(rev(brewer.pal(n = 5, name ="RdBu")), "#000000")
pheatmap(plot.data, treeheight_row=0,treeheight_col=0,  
         color =palette, breaks = c(min(plot.data),-1,-0.5,0,0.5,1,max(plot.data)))
```

##  CRC metastatic tumors in the liver treated patients

```{r, warning=FALSE, message=FALSE}
misty_folders_selected <- misty_folders_all[c(6,7)]

misty_results <- collect_results(misty_folders_selected)

```

```{r, warning=FALSE, message=FALSE, dpi=300}
plot.data <- 
  misty_results$importances.aggregated %>%
  dplyr::filter(view == "juxta.ligands_2") %>% 
  dplyr::select(Predictor, Target, Importance ) %>% 
  tidyr::pivot_wider(names_from = Predictor,values_from = Importance ) %>% 
  tibble::column_to_rownames(var = 'Target') %>% 
  as.matrix() %>% t()

palette <-c(rev(brewer.pal(n = 5, name ="RdBu")), "#000000")
pheatmap(plot.data, treeheight_row=0,treeheight_col=0,  
         color =palette, breaks = c(min(plot.data),-1,-0.5,0,0.5,1,max(plot.data)))
```

##  CRC metastatic tumors in the liver large tumors

```{r, warning=FALSE, message=FALSE}
misty_folders_selected <- misty_folders_all[c(4,5,7)]

misty_results <- collect_results(misty_folders_selected)

```

```{r, warning=FALSE, message=FALSE, dpi=300}
plot.data <- 
  misty_results$importances.aggregated %>%
  dplyr::filter(view == "juxta.ligands_2") %>% 
  dplyr::select(Predictor, Target, Importance ) %>% 
  tidyr::pivot_wider(names_from = Predictor,values_from = Importance ) %>% 
  tibble::column_to_rownames(var = 'Target') %>% 
  as.matrix() %>% t()

palette <-c(rev(brewer.pal(n = 5, name ="RdBu")), "#000000")
pheatmap(plot.data, treeheight_row=0,treeheight_col=0,  
         color =palette, breaks = c(min(plot.data),-1,-0.5,0,0.5,1,max(plot.data)))
```

## Publication plot

```{r, warning=FALSE, message=FALSE}
seurat_objects <- readRDS(files_to_read_seurat)
```

### ST-colon 1

```{r, warning=FALSE, message=FALSE}
current_sample <- "ST-colon1"
current_object <- seurat_objects[[current_sample]]
DefaultAssay(current_object) <- "SCT"
current_gene <- c("RNF43")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["RNF43",]), 
              max(GetAssayData(current_object)["RNF43",]), length.out= 4),1)

ST_colon1_RNF43 <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = current_gene, alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("RNF43 Expression") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL ) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(NULL),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 



DefaultAssay(current_object) <- "dorothea"
current_gene <- c("TEAD1")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["TEAD1",]), 
              max(GetAssayData(current_object)["TEAD1",]), length.out= 4),1)


ST_colon1_TEAD1 <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = c(current_gene), alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("TEAD1 TF activity") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(angle = 90, face="bold", size=10),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 

current_gene <- c("JUN")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["JUN",]), 
              max(GetAssayData(current_object)["JUN",]), length.out= 4),1)


ST_colon1_JUN <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = c(current_gene), alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("JUN TF activity") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(angle = 90, face="bold", size=10),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 
```

```{r, warning=FALSE, message=FALSE}
ST_colon1_RNF43
```

```{r, warning=FALSE, message=FALSE}
ST_colon1_TEAD1
```

```{r, warning=FALSE, message=FALSE}
ST_colon1_JUN
```


```{r, warning=FALSE, message=FALSE}
DefaultAssay(current_object) <- "SCT"
current_gene <- c("CXCL14")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["CXCL14",]), 
              max(GetAssayData(current_object)["CXCL14",]), length.out= 4),1)

ST_colon1_CXCL14 <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = current_gene, alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("CXCL14 Expression") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL ) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(NULL),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 



DefaultAssay(current_object) <- "dorothea"
current_gene <- c("MAF")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["MAF",]), 
              max(GetAssayData(current_object)["MAF",]), length.out= 4),1)


ST_colon1_MAF <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = c(current_gene), alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("MAF TF activity") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(angle = 90, face="bold", size=10),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 
```


```{r, warning=FALSE, message=FALSE}
ST_colon1_CXCL14
```

```{r, warning=FALSE, message=FALSE}
ST_colon1_MAF
```

```{r, warning=FALSE, message=FALSE}
DefaultAssay(current_object) <- "SCT"
current_gene <- c("DCN")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["DCN",]), 
              max(GetAssayData(current_object)["DCN",]), length.out= 4),1)

ST_colon1_DCN <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = current_gene, alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("DCN Expression") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL ) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(NULL),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 



DefaultAssay(current_object) <- "dorothea"
current_gene <- c("ETV4")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["ETV4",]), 
              max(GetAssayData(current_object)["ETV4",]), length.out= 4),1)


ST_colon1_ETV4 <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = c(current_gene), alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("ETV4 TF activity") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(angle = 90, face="bold", size=10),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 
```

```{r}
ST_colon1_DCN
```


```{r}
ST_colon1_ETV4
```



### ST-colon 2

```{r, warning=FALSE, message=FALSE}
current_sample <- "ST-colon2"
current_object <- seurat_objects[[current_sample]]
DefaultAssay(current_object) <- "SCT"
current_gene <- c("RNF43")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["RNF43",]), 
              max(GetAssayData(current_object)["RNF43",]), length.out= 4),1)

ST_colon2_RNF43 <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = current_gene, alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("RNF43 Expression") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL ) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(NULL),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 



DefaultAssay(current_object) <- "dorothea"
current_gene <- c("TEAD1")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["TEAD1",]), 
              max(GetAssayData(current_object)["TEAD1",]), length.out= 4),1)


ST_colon2_TEAD1 <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = c(current_gene), alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("TEAD1 TF activity") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(angle = 90, face="bold", size=10),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 

current_gene <- c("JUN")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["JUN",]), 
              max(GetAssayData(current_object)["JUN",]), length.out= 4),1)


ST_colon2_JUN <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = c(current_gene), alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("JUN TF activity") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(angle = 90, face="bold", size=10),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 
```

```{r, warning=FALSE, message=FALSE}
ST_colon2_RNF43
```

```{r, warning=FALSE, message=FALSE}
ST_colon2_TEAD1
```

```{r, warning=FALSE, message=FALSE}
ST_colon2_JUN
```


```{r, warning=FALSE, message=FALSE}
DefaultAssay(current_object) <- "SCT"
current_gene <- c("CXCL14")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["CXCL14",]), 
              max(GetAssayData(current_object)["CXCL14",]), length.out= 4),1)

ST_colon2_CXCL14 <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = current_gene, alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("CXCL14 Expression") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL ) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(NULL),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 



DefaultAssay(current_object) <- "dorothea"
current_gene <- c("MAF")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["MAF",]), 
              max(GetAssayData(current_object)["MAF",]), length.out= 4),1)


ST_colon2_MAF <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = c(current_gene), alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("MAF TF activity") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(angle = 90, face="bold", size=10),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 
```


```{r, warning=FALSE, message=FALSE}
ST_colon2_CXCL14
```

```{r, warning=FALSE, message=FALSE}
ST_colon2_MAF
```

```{r, warning=FALSE, message=FALSE}
DefaultAssay(current_object) <- "SCT"
current_gene <- c("DCN")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["DCN",]), 
              max(GetAssayData(current_object)["DCN",]), length.out= 4),1)

ST_colon2_DCN <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = current_gene, alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("DCN Expression") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL ) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(NULL),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 



DefaultAssay(current_object) <- "dorothea"
current_gene <- c("ETV4")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["ETV4",]), 
              max(GetAssayData(current_object)["ETV4",]), length.out= 4),1)


ST_colon2_ETV4 <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = c(current_gene), alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("ETV4 TF activity") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(angle = 90, face="bold", size=10),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 
```

```{r}
ST_colon2_DCN
```


```{r}
ST_colon2_ETV4
```



### ST-colon 4

```{r, warning=FALSE, message=FALSE}
current_sample <- "ST-colon4"
current_object <- seurat_objects[[current_sample]]
DefaultAssay(current_object) <- "SCT"
current_gene <- c("RNF43")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["RNF43",]), 
              max(GetAssayData(current_object)["RNF43",]), length.out= 4),1)

ST_colon4_RNF43 <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = current_gene, alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("RNF43 Expression") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL ) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(NULL),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 



DefaultAssay(current_object) <- "dorothea"
current_gene <- c("TEAD1")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["TEAD1",]), 
              max(GetAssayData(current_object)["TEAD1",]), length.out= 4),1)


ST_colon4_TEAD1 <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = c(current_gene), alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("TEAD1 TF activity") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(angle = 90, face="bold", size=10),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 

current_gene <- c("JUN")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["JUN",]), 
              max(GetAssayData(current_object)["JUN",]), length.out= 4),1)


ST_colon4_JUN <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = c(current_gene), alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("JUN TF activity") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(angle = 90, face="bold", size=10),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 
```

```{r, warning=FALSE, message=FALSE}
ST_colon4_RNF43
```

```{r, warning=FALSE, message=FALSE}
ST_colon4_TEAD1
```

```{r, warning=FALSE, message=FALSE}
ST_colon4_JUN
```


```{r, warning=FALSE, message=FALSE}
DefaultAssay(current_object) <- "SCT"
current_gene <- c("CXCL14")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["CXCL14",]), 
              max(GetAssayData(current_object)["CXCL14",]), length.out= 4),1)

ST_colon4_CXCL14 <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = current_gene, alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("CXCL14 Expression") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL ) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(NULL),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 



DefaultAssay(current_object) <- "dorothea"
current_gene <- c("MAF")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["MAF",]), 
              max(GetAssayData(current_object)["MAF",]), length.out= 4),1)


ST_colon4_MAF <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = c(current_gene), alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("MAF TF activity") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(angle = 90, face="bold", size=10),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 
```


```{r, warning=FALSE, message=FALSE}
ST_colon4_CXCL14
```

```{r, warning=FALSE, message=FALSE}
ST_colon4_MAF
```


```{r, warning=FALSE, message=FALSE}
DefaultAssay(current_object) <- "SCT"
current_gene <- c("DCN")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["DCN",]), 
              max(GetAssayData(current_object)["DCN",]), length.out= 4),1)

ST_colon4_DCN <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = current_gene, alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("DCN Expression") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL ) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(NULL),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 



DefaultAssay(current_object) <- "dorothea"
current_gene <- c("ETV4")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["ETV4",]), 
              max(GetAssayData(current_object)["ETV4",]), length.out= 4),1)


ST_colon4_ETV4 <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = c(current_gene), alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("ETV4 TF activity") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(angle = 90, face="bold", size=10),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 
```

```{r}
ST_colon4_DCN
```


```{r}
ST_colon4_ETV4
```


### ST-Liver 1

```{r, warning=FALSE, message=FALSE}
current_sample <- "ST-liver1"
current_object <- seurat_objects[[current_sample]]
DefaultAssay(current_object) <- "SCT"
current_gene <- c("DCN")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["DCN",]), 
              max(GetAssayData(current_object)["DCN",]), length.out= 4),1)

ST_liver1_DCN <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = current_gene, alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("DCN Expression") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL ) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(NULL),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 



DefaultAssay(current_object) <- "dorothea"
current_gene <- c("ETV4")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["ETV4",]), 
              max(GetAssayData(current_object)["ETV4",]), length.out= 4),1)


ST_liver1_ETV4 <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = c(current_gene), alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("ETV4 TF activity") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(angle = 90, face="bold", size=10),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 
```

```{r}
ST_liver1_DCN
```


```{r}
ST_liver1_ETV4
```

```{r, warning=FALSE, message=FALSE}
DefaultAssay(current_object) <- "SCT"
current_gene <- c("RNF43")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["RNF43",]), 
              max(GetAssayData(current_object)["RNF43",]), length.out= 4),1)

ST_liver1_RNF43 <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = current_gene, alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("RNF43 Expression") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL ) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(NULL),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 



DefaultAssay(current_object) <- "dorothea"
current_gene <- c("JUN")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["JUN",]), 
              max(GetAssayData(current_object)["JUN",]), length.out= 4),1)


ST_liver1_JUN <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = c(current_gene), alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("JUN TF activity") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(angle = 90, face="bold", size=10),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 
```

```{r}
ST_liver1_RNF43
```


```{r}
ST_liver1_JUN
```


### ST-Liver 2

```{r, warning=FALSE, message=FALSE}
current_sample <- "ST-liver2"
current_object <- seurat_objects[[current_sample]]
DefaultAssay(current_object) <- "SCT"
current_gene <- c("DCN")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["DCN",]), 
              max(GetAssayData(current_object)["DCN",]), length.out= 4),1)

ST_liver2_DCN <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = current_gene, alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("DCN Expression") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL ) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(NULL),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 



DefaultAssay(current_object) <- "dorothea"
current_gene <- c("ETV4")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["ETV4",]), 
              max(GetAssayData(current_object)["ETV4",]), length.out= 4),1)


ST_liver2_ETV4 <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = c(current_gene), alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("ETV4 TF activity") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(angle = 90, face="bold", size=10),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 
```

```{r}
ST_liver2_DCN
```


```{r}
ST_liver2_ETV4
```


```{r, warning=FALSE, message=FALSE}
DefaultAssay(current_object) <- "SCT"
current_gene <- c("RNF43")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["RNF43",]), 
              max(GetAssayData(current_object)["RNF43",]), length.out= 4),1)

ST_liver2_RNF43 <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = current_gene, alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("RNF43 Expression") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL ) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(NULL),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 



DefaultAssay(current_object) <- "dorothea"
current_gene <- c("JUN")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["JUN",]), 
              max(GetAssayData(current_object)["JUN",]), length.out= 4),1)


ST_liver2_JUN <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = c(current_gene), alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("JUN TF activity") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(angle = 90, face="bold", size=10),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 
```

```{r}
ST_liver2_RNF43
```


```{r}
ST_liver2_JUN
```


### ST-Liver 4

```{r, warning=FALSE, message=FALSE}
current_sample <- "ST-liver4"
current_object <- seurat_objects[[current_sample]]
DefaultAssay(current_object) <- "SCT"
current_gene <- c("DCN")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["DCN",]), 
              max(GetAssayData(current_object)["DCN",]), length.out= 4),1)

ST_liver4_DCN <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = current_gene, alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("DCN Expression") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL ) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(NULL),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 



DefaultAssay(current_object) <- "dorothea"
current_gene <- c("ETV4")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["ETV4",]), 
              max(GetAssayData(current_object)["ETV4",]), length.out= 4),1)


ST_liver4_ETV4 <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = c(current_gene), alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("ETV4 TF activity") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(angle = 90, face="bold", size=10),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 
```

```{r}
ST_liver4_DCN
```


```{r}
ST_liver4_ETV4
```


```{r, warning=FALSE, message=FALSE}
DefaultAssay(current_object) <- "SCT"
current_gene <- c("RNF43")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["RNF43",]), 
              max(GetAssayData(current_object)["RNF43",]), length.out= 4),1)

ST_liver4_RNF43 <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = current_gene, alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("RNF43 Expression") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL ) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(NULL),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 



DefaultAssay(current_object) <- "dorothea"
current_gene <- c("JUN")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)["JUN",]), 
              max(GetAssayData(current_object)["JUN",]), length.out= 4),1)


ST_liver4_JUN <- 
    Seurat::SpatialFeaturePlot(
      object = current_object,
      features = c(current_gene), alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 1.75, stroke = 1) + 
      ggtitle("JUN TF activity") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(angle = 90, face="bold", size=10),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5)) 
```

```{r}
ST_liver4_RNF43
```


```{r}
ST_liver4_JUN
```


# Session Info Details

```{r, echo=FALSE, eval=TRUE}
sessionInfo()
```