---
title: "Analysis of CRC Spatial Transcriptomics data from 10X:Correlation Analysis of TFs Activities and Cell Type abundances for CMS"
author: 'Alberto Valdeolivas Urbelz (PS-Biomics&Path)'
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    code_folding: hide
    self_contained: true
    number_sections: yes
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: spacelab
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
params:
  data_directory : "/projects/site/pred/SpatialOmics/Experiment_CRC_AllSamples/"
  analysis_name : "analysis_alberto/"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

The present script takes the the Spatial Omics data generated with 10X 
technology on a set of CRC samples. In particular, it explores the correlation 
between TF activities and CMS cell type proportion per spot.  


```{r, warning=FALSE, message=FALSE}
# library(dplyr, lib.loc = "/pstore/apps/bioinfo/R/4.0.1-foss")
library(CCA)
library(ggplot2)
library(patchwork, lib.loc = "/apps/rocs/2020.08/cascadelake/software/R/4.1.2-foss-2020a/lib64/R/library")
library(Seurat)
library(readr)
library(stringr)
library(tibble)
library(ggrepel)
library(RColorBrewer)
library(pheatmap)
source(file = "../WrapperFunction/SeuratWrappers.R")

data_directory <-  params$data_directory
analysis_name <- params$analysis_name

input_folder_seurat <- "ST_LiverMetastases_pub/IntermediaryFiles/"
input_names_seurat <- "SeuratList_Clusters_Res05_Dorothea.rds"
files_to_read_seurat <- 
  paste0(data_directory, analysis_name,input_folder_seurat,input_names_seurat)


input_folder_C2L <- 
  "Cell2Location/results/LocationModelLinearDependentWMultiExperiment_8experiments_36clusters_31296locations_4188genesLiverMetastasis/"
input_file_C2L <- "W_mRNA_count_q05.csv"
files_to_read_C2L <- 
  paste0(data_directory, analysis_name,input_folder_C2L,input_file_C2L)
```

We read the Seurat objects containing the TF activity. 

```{r, warning=FALSE, message=FALSE}
seurat_objects <- readRDS(files_to_read_seurat)
```


We Load Cell2Location results cotaning the abundance of the different cell types
per spot. 

```{r, warning=FALSE, message=FALSE}
results_STLiver_Korean <- read_csv(files_to_read_C2L)


sum_UMI_spot <- rowSums(results_STLiver_Korean[,c(2: (ncol(results_STLiver_Korean)-1))])
results_STLiver_Korean[,-1] <- results_STLiver_Korean[,-1]/sum_UMI_spot

colnames(results_STLiver_Korean) <- str_replace(colnames(results_STLiver_Korean), 
    pattern = "q05_nUMI_factors", replacement = "")


## We are only going to consider the CMS signatures

results_STLiver_Korean <- 
  results_STLiver_Korean[,colnames(results_STLiver_Korean) %in% 
                c("spot_id","CMS1", "CMS2", "CMS3", "CMS4")]
```

# Results taking into account correlation between the spots of colon untreated samples

We display the different correlation plots when this is globally computed per 
all the samples. 

```{r, warning=FALSE, message=FALSE}
seurat_objects_CRC_Unt <- 
  seurat_objects[names(seurat_objects) %in% 
                   c("ST-colon1", "ST-colon2") == TRUE]
```


```{r, results = 'asis', message=FALSE, dpi=300, warning=FALSE}
list_df_TF <- lapply(seurat_objects_CRC_Unt, function(x){
  matrix <- GetAssayData(x)
  colnames(matrix) <- paste0(x$orig.ident,"_", colnames(matrix))
  return(matrix %>% t() %>% as.data.frame() %>% tibble::rownames_to_column(var = "spot_id"))
})

TF_df_allsamples <- do.call(rbind.data.frame, list_df_TF)


merge_CL2_TFs <- TF_df_allsamples %>%
  dplyr::inner_join(results_STLiver_Korean, by ="spot_id") %>% 
  dplyr::arrange(spot_id) 

X <- merge_CL2_TFs[,2:ncol(TF_df_allsamples)]
Y <- merge_CL2_TFs[,(ncol(TF_df_allsamples) +1):ncol(merge_CL2_TFs)]

correl <- matcor(X, Y)
canonical_correlation2 <- cc(X, Y)
  
cat( "\n\n")

cat( "## Cross TFs-Cell type abundances Correlation: ", "\n\n")
  
CrossCorrelation <- 
  correl$XYcor[1:(ncol(TF_df_allsamples)-1),ncol(TF_df_allsamples):(ncol(merge_CL2_TFs)-1)]


Higher_TFcorrelations <-   CrossCorrelation %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column(var ="TF") %>%
  tidyr::pivot_longer(!TF, values_drop_na = TRUE) %>% 
  dplyr::group_by(name) %>%
  # dplyr::mutate(value = round(value,2)) %>% 
  dplyr::slice_max(order_by = abs(value), n =10) %>% 
  dplyr::pull(TF) %>% unique()

CrossCorrelation_higherTFs <- 
  CrossCorrelation[rownames(CrossCorrelation) %in% Higher_TFcorrelations,]

MatrixCorrXY <-   CrossCorrelation_higherTFs %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column(var ="TF") %>%
  tidyr::pivot_longer(!TF, values_drop_na = TRUE) %>% 
  dplyr::mutate(value = round(value,2))


p_CrossCorrelation <- ggplot(data = MatrixCorrXY, 
                              aes(TF  ,name  , fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Pearson\nCorrelation") +
  theme_minimal()+  theme(legend.position="top")  +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 12, hjust = 1),
        axis.text.y = element_text(size = 12))+
  xlab("TFs") + ylab("Cell Types") + 
  coord_fixed() 


print(p_CrossCorrelation)
  
mypalette <- rev(brewer.pal(n = 9, name ="RdBu"))
  
par(mar=c(10,10,10,10))
    
pheatmap(t(CrossCorrelation[Higher_TFcorrelations,]),
         treeheight_row=0,treeheight_col=0, 
  color =colorRampPalette(mypalette)(100),cluster_rows = FALSE,
  fontsize = 12, cluster_cols = TRUE, breaks=seq(-1,1,length.out = 
  length(colorRampPalette(mypalette)(100)) + 1))


Main_tfs <- MatrixCorrXY %>% dplyr::pull(TF)

cat( "\n\n")
  
cat( "##  TF activities Correlation: ", "\n\n")
  
MatrixCorrX <-   correl$Xcor %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column(var ="TF") %>%
  tidyr::pivot_longer(!TF, values_drop_na = TRUE) %>% 
  dplyr::filter(TF %in% Main_tfs & name %in% Main_tfs)
  # dplyr::mutate(value = round(value,2))
  
p_cor_TFs <- ggplot(data = MatrixCorrX, aes(TF, name , fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Pearson\nCorrelation") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                 size = 10, hjust = 1),
        axis.text.y = element_text(size = 10))+
  xlab("TFs") + ylab("TFs") + 
  coord_fixed()

print(p_cor_TFs)
  
pheatmap(correl$Xcor, treeheight_row=0,treeheight_col=0, 
  color =colorRampPalette(mypalette)(100),cluster_rows = FALSE,
  fontsize = 12, cluster_cols = FALSE, breaks=seq(-1,1,length.out = 
  length(colorRampPalette(mypalette)(100)) + 1))


cat( "\n\n")
  
cat( "## Variables representation on the canonical variates: ", "\n\n")

idx <- which(colnames(X) %in% Higher_TFcorrelations)
X <- X[,idx]
canonical_correlation2 <- cc(X, Y)
  
  
d1 = 2 
d2 = 1
df2 = data.frame(comp1 = c(canonical_correlation2$scores$corr.X.xscores[, d1],
                           canonical_correlation2$scores$corr.Y.xscores[, d1]),
                comp2 = c(canonical_correlation2$scores$corr.X.xscores[, d2],
                           canonical_correlation2$scores$corr.Y.xscores[, d2]),
                color = c(rep("TFs", ncol(X)), 
                          (rep("Cell Types", ncol(Y)))))
  
p_Canonival_variates <-ggplot(df2, aes(comp1, comp2), colour = color) +
  geom_label_repel(aes(colour = color, label= rownames(df2)),
          vjust=0,size = 5, family = "sans", direction= "both",
          max.overlaps=12) +
  theme_bw() + 
  theme(legend.position="none") +
  theme(legend.title=element_blank())
  
print(p_Canonival_variates)
  
cat( "\n\n") 

```


# Results taking into account correlation between the spots of colon samples

We display the different correlation plots when this is globally computed per 
all the samples. 

```{r, warning=FALSE, message=FALSE}
seurat_objects_CRC <- 
  seurat_objects[names(seurat_objects) %in% 
                   c("ST-colon1", "ST-colon2", "ST-colon3", "ST-colon4") == TRUE]
```


```{r, results = 'asis', message=FALSE, dpi=300, warning=FALSE}
list_df_TF <- lapply(seurat_objects_CRC, function(x){
  matrix <- GetAssayData(x)
  colnames(matrix) <- paste0(x$orig.ident,"_", colnames(matrix))
  return(matrix %>% t() %>% as.data.frame() %>% tibble::rownames_to_column(var = "spot_id"))
})

TF_df_allsamples <- do.call(rbind.data.frame, list_df_TF)


merge_CL2_TFs <- TF_df_allsamples %>%
  dplyr::inner_join(results_STLiver_Korean, by ="spot_id") %>% 
  dplyr::arrange(spot_id) 

X <- merge_CL2_TFs[,2:ncol(TF_df_allsamples)]
Y <- merge_CL2_TFs[,(ncol(TF_df_allsamples) +1):ncol(merge_CL2_TFs)]

correl <- matcor(X, Y)
canonical_correlation2 <- cc(X, Y)
  
cat( "\n\n")

cat( "## Cross TFs-Cell type abundances Correlation: ", "\n\n")
  
CrossCorrelation <- 
  correl$XYcor[1:(ncol(TF_df_allsamples)-1),ncol(TF_df_allsamples):(ncol(merge_CL2_TFs)-1)]


Higher_TFcorrelations <-   CrossCorrelation %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column(var ="TF") %>%
  tidyr::pivot_longer(!TF, values_drop_na = TRUE) %>% 
  dplyr::group_by(name) %>%
  # dplyr::mutate(value = round(value,2)) %>% 
  dplyr::slice_max(order_by = abs(value), n =10) %>% 
  dplyr::pull(TF) %>% unique()

CrossCorrelation_higherTFs <- 
  CrossCorrelation[rownames(CrossCorrelation) %in% Higher_TFcorrelations,]

MatrixCorrXY <-   CrossCorrelation_higherTFs %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column(var ="TF") %>%
  tidyr::pivot_longer(!TF, values_drop_na = TRUE) %>% 
  dplyr::mutate(value = round(value,2))


p_CrossCorrelation <- ggplot(data = MatrixCorrXY, 
                              aes(TF  ,name  , fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Pearson\nCorrelation") +
  theme_minimal()+  theme(legend.position="top")  +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 12, hjust = 1),
        axis.text.y = element_text(size = 12))+
  xlab("TFs") + ylab("Cell Types") + 
  coord_fixed() 


print(p_CrossCorrelation)
  
mypalette <- rev(brewer.pal(n = 9, name ="RdBu"))
  
par(mar=c(10,10,10,10))
    
pheatmap(t(CrossCorrelation[Higher_TFcorrelations,]),
         treeheight_row=0,treeheight_col=0, 
  color =colorRampPalette(mypalette)(100),cluster_rows = FALSE,
  fontsize = 12, cluster_cols = TRUE, breaks=seq(-1,1,length.out = 
  length(colorRampPalette(mypalette)(100)) + 1))


Main_tfs <- MatrixCorrXY %>% dplyr::pull(TF)

cat( "\n\n")
  
cat( "##  TF activities Correlation: ", "\n\n")
  
MatrixCorrX <-   correl$Xcor %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column(var ="TF") %>%
  tidyr::pivot_longer(!TF, values_drop_na = TRUE) %>% 
  dplyr::filter(TF %in% Main_tfs & name %in% Main_tfs)
  # dplyr::mutate(value = round(value,2))
  
p_cor_TFs <- ggplot(data = MatrixCorrX, aes(TF, name , fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Pearson\nCorrelation") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                 size = 10, hjust = 1),
        axis.text.y = element_text(size = 10))+
  xlab("TFs") + ylab("TFs") + 
  coord_fixed()

print(p_cor_TFs)
  
pheatmap(correl$Xcor, treeheight_row=0,treeheight_col=0, 
  color =colorRampPalette(mypalette)(100),cluster_rows = FALSE,
  fontsize = 12, cluster_cols = FALSE, breaks=seq(-1,1,length.out = 
  length(colorRampPalette(mypalette)(100)) + 1))


cat( "\n\n")
  
cat( "## Variables representation on the canonical variates: ", "\n\n")

idx <- which(colnames(X) %in% Higher_TFcorrelations)
X <- X[,idx]
canonical_correlation2 <- cc(X, Y)
  
  
d1 = 2 
d2 = 1
df2 = data.frame(comp1 = c(canonical_correlation2$scores$corr.X.xscores[, d1],
                           canonical_correlation2$scores$corr.Y.xscores[, d1]),
                comp2 = c(canonical_correlation2$scores$corr.X.xscores[, d2],
                           canonical_correlation2$scores$corr.Y.xscores[, d2]),
                color = c(rep("TFs", ncol(X)), 
                          (rep("Cell Types", ncol(Y)))))
  
p_Canonival_variates <-ggplot(df2, aes(comp1, comp2), colour = color) +
  geom_label_repel(aes(colour = color, label= rownames(df2)),
          vjust=0,size = 5, family = "sans", direction= "both",
          max.overlaps=12) +
  theme_bw() + 
  theme(legend.position="none") +
  theme(legend.title=element_blank())
  
print(p_Canonival_variates)
  
cat( "\n\n") 

```


# Results taking into account correlation between all the samples samples

We display the different correlation plots when this is globally computed per 
all the samples. 

```{r, results = 'asis', message=FALSE, dpi=300, warning=FALSE}
list_df_TF <- lapply(seurat_objects, function(x){
  matrix <- GetAssayData(x)
  colnames(matrix) <- paste0(x$orig.ident,"_", colnames(matrix))
  return(matrix %>% t() %>% as.data.frame() %>% tibble::rownames_to_column(var = "spot_id"))
})

TF_df_allsamples <- do.call(rbind.data.frame, list_df_TF)


merge_CL2_TFs <- TF_df_allsamples %>%
  dplyr::inner_join(results_STLiver_Korean, by ="spot_id") %>% 
  dplyr::arrange(spot_id) 

X <- merge_CL2_TFs[,2:ncol(TF_df_allsamples)]
Y <- merge_CL2_TFs[,(ncol(TF_df_allsamples) +1):ncol(merge_CL2_TFs)]

correl <- matcor(X, Y)
canonical_correlation2 <- cc(X, Y)
  
cat( "\n\n")

cat( "## Cross TFs-Cell type abundances Correlation: ", "\n\n")
  
CrossCorrelation <- 
  correl$XYcor[1:(ncol(TF_df_allsamples)-1),ncol(TF_df_allsamples):(ncol(merge_CL2_TFs)-1)]


Higher_TFcorrelations <-   CrossCorrelation %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column(var ="TF") %>%
  tidyr::pivot_longer(!TF, values_drop_na = TRUE) %>% 
  dplyr::group_by(name) %>%
  # dplyr::mutate(value = round(value,2)) %>% 
  dplyr::slice_max(order_by = abs(value), n =10) %>% 
  dplyr::pull(TF) %>% unique()

CrossCorrelation_higherTFs <- 
  CrossCorrelation[rownames(CrossCorrelation) %in% Higher_TFcorrelations,]

MatrixCorrXY <-   CrossCorrelation_higherTFs %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column(var ="TF") %>%
  tidyr::pivot_longer(!TF, values_drop_na = TRUE) %>% 
  dplyr::mutate(value = round(value,2))


p_CrossCorrelation <- ggplot(data = MatrixCorrXY, 
                              aes(TF  ,name  , fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Pearson\nCorrelation") +
  theme_minimal()+  theme(legend.position="top")  +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 12, hjust = 1),
        axis.text.y = element_text(size = 12))+
  xlab("TFs") + ylab("Cell Types") + 
  coord_fixed() 


print(p_CrossCorrelation)
  
mypalette <- rev(brewer.pal(n = 9, name ="RdBu"))
  
par(mar=c(10,10,10,10))
    
pheatmap(t(CrossCorrelation[Higher_TFcorrelations,]),
         treeheight_row=0,treeheight_col=0, 
  color =colorRampPalette(mypalette)(100),cluster_rows = FALSE,
  fontsize = 12, cluster_cols = TRUE, breaks=seq(-1,1,length.out = 
  length(colorRampPalette(mypalette)(100)) + 1))

pheatmap(CrossCorrelation[Higher_TFcorrelations,],
         treeheight_row=0,treeheight_col=0, 
  color =colorRampPalette(mypalette)(100),cluster_rows = TRUE,
  fontsize = 12, cluster_cols = FALSE, breaks=seq(-1,1,length.out = 
  length(colorRampPalette(mypalette)(100)) + 1))


Main_tfs <- MatrixCorrXY %>% dplyr::pull(TF)

cat( "\n\n")
  
cat( "##  TF activities Correlation: ", "\n\n")
  
MatrixCorrX <-   correl$Xcor %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column(var ="TF") %>%
  tidyr::pivot_longer(!TF, values_drop_na = TRUE) %>% 
  dplyr::filter(TF %in% Main_tfs & name %in% Main_tfs)
  # dplyr::mutate(value = round(value,2))
  
p_cor_TFs <- ggplot(data = MatrixCorrX, aes(TF, name , fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Pearson\nCorrelation") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                 size = 10, hjust = 1),
        axis.text.y = element_text(size = 10))+
  xlab("TFs") + ylab("TFs") + 
  coord_fixed()

print(p_cor_TFs)
  
pheatmap(correl$Xcor, treeheight_row=0,treeheight_col=0, 
  color =colorRampPalette(mypalette)(100),cluster_rows = FALSE,
  fontsize = 12, cluster_cols = FALSE, breaks=seq(-1,1,length.out = 
  length(colorRampPalette(mypalette)(100)) + 1))

cat( "\n\n")
  
cat( "## Variables representation on the canonical variates: ", "\n\n")

idx <- which(colnames(X) %in% Higher_TFcorrelations)
X <- X[,idx]
canonical_correlation2 <- cc(X, Y)
  
  
d1 = 2 
d2 = 1
df2 = data.frame(comp1 = c(canonical_correlation2$scores$corr.X.xscores[, d1],
                           canonical_correlation2$scores$corr.Y.xscores[, d1]),
                comp2 = c(canonical_correlation2$scores$corr.X.xscores[, d2],
                           canonical_correlation2$scores$corr.Y.xscores[, d2]),
                color = c(rep("TFs", ncol(X)), 
                          (rep("Cell Types", ncol(Y)))))
  
p_Canonival_variates <-ggplot(df2, aes(comp1, comp2), colour = color) +
  geom_label_repel(aes(colour = color, label= rownames(df2)),
          vjust=0,size = 5, family = "sans", direction= "both",
          max.overlaps=12) +
  theme_bw() + 
  theme(legend.position="none") +
  theme(legend.title=element_blank())
  
print(p_Canonival_variates)
  
cat( "\n\n") 

```




# Conclusion 


# Session Info Details

```{r, echo=FALSE, eval=TRUE}
sessionInfo()
```