---
title: "Analysis of CRC Spatial Transcriptomics data from 10X: Pseudo-bulk CMS based classification"
author: 'Alberto Valdeolivas Urbelz (PS-Biomics&Path)'
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    code_folding: hide
    self_contained: true
    number_sections: yes
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: spacelab
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
params:
  data_directory : "/projects/site/pred/SpatialOmics/Experiment_CRC_AllSamples/"
  analysis_name : "analysis_alberto/"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

In this script, we generate pseudo-bulk of the external set of 10x VISIUM ST CRC samples extracted from the following publication:  

_Wu, Y. et al. Spatiotemporal Immune Landscape of Colorectal Cancer Liver Metastasis at Single-Cell Level. Cancer Discov. 12, 134–153 (2022)._

and we use https://github.com/Lothelab/CMScaller to assign them to the different CMS. 


```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(vctrs)
library(patchwork, lib.loc = "/apps/rocs/2020.08/cascadelake/software/R/4.1.2-foss-2020a/lib64/R/library")
library(Seurat)
library(scater)
library(dplyr)
library(readr)
library(stringr)
library(RColorBrewer)
library(cowplot)
library(Biobase)
library(CMScaller)


source(file = "../WrapperFunction/SeuratWrappers.R")
source(file = "../WrapperFunction/PseudoBulkUtils.R")


data_directory <-  params$data_directory
analysis_name <- params$analysis_name

input_folder <- "ST_LiverMetastases_pub/IntermediaryFiles/"
input_names_1 <- "SeuratList_Clusters_Res05.rds"

files_to_read_1 <- paste0(data_directory, analysis_name,input_folder,input_names_1)
```

# Results

I read the seurat object 

```{r, warning=FALSE, message=FALSE}
seurat_objects <- readRDS(files_to_read_1)
```


## Pseudo-bulk of the whole sample

```{r, warning=FALSE, message=FALSE}
pseudobulk_profiles_all <- lapply(seurat_objects, get_sample_pseudo)


gex_all <- do.call(cbind, lapply(pseudobulk_profiles_all, function(x){
  assay(x)}))

## We change the names for the plotting: 

colnames(gex_all) <- 
  c("ST-colon1_Unt", "ST-colon2_Unt", 
    "ST-colon3_Tre", "ST-colon4_Tre", 
    "ST-liver1_Unt", "ST-liver2_Unt",
    "ST-liver3_Tre", "ST-liver4_Tre")
```

We call CMScaller to see the results on the whole section of the samples. 

```{r, warning=FALSE, message=FALSE}
gex_all_entrez <- replaceGeneId(gex_all, id.in = "symbol", id.out = "entrez")
gex_all_entrez <- gex_all_entrez[!(str_detect(rownames(gex_all_entrez), "^NA.")),]

res_all <- CMScaller(gex_all_entrez, RNAseq=TRUE, doPlot = FALSE)

res_df_all <- res_all %>% as.data.frame() %>% tibble::rownames_to_column(var = "sample") %>% 
  dplyr::select(sample, prediction) 

res_df_all$sample <- 
  factor(res_df_all$sample , 
         levels = sort(unique(res_df_all$sample), decreasing = TRUE)) 


res_df_all %>% ggplot(aes(prediction, sample, color = prediction)) +
  geom_point(size = 8) + 
  scale_color_brewer(palette = "Paired") + 
  theme( # remove the vertical grid lines
    axis.title = element_text(size=14, face = "bold", family="Arial"), 
    axis.text.x = element_text(size=10, face = "bold", family="Arial", angle = 90),
    axis.text.y = element_text(size=10, face = "bold", family="Arial")) + 
  labs(x="Predicted CMS", y ="Sample") + 
  theme(legend.position ="none") 
```

```{r, warning=FALSE, message=FALSE}
cam <- CMSgsa(emat=gex_all_entrez, class=res_df_all$prediction,RNAseq=TRUE)
```

## Plots for Supplementary material 

```{r, warning=FALSE, message=FALSE, dpi=300}
res_df_all %>%  dplyr::mutate(type = "CMS") %>% 
  ggplot(aes(type, sample , fill= prediction)) + 
  geom_tile(size =1.25, color ="black", width = 0.95) + 
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        axis.ticks = element_blank(), panel.border = element_blank(), 
          panel.background = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          axis.text.x = element_blank(),
          axis.text.y = element_text(size=12, face = "bold", family="Arial"),
          legend.position ="right", legend.text = element_text(size=12, face="bold", family="Arial"),
          legend.title =  element_text(size=12, face="bold", family="Arial")) +
          scale_fill_brewer(palette = "Set1") + 
          guides(fill = guide_legend(title = "Predicted CMS", override.aes = list(size = 10))) 
```

# Session Info Details

```{r, echo=FALSE, eval=TRUE}
sessionInfo()
```