---
title: "Analysis of CRC Spatial Transcriptomics data from 10X: Proportion plots"
author: 'Alberto Valdeolivas Urbelz (PS-Biomics&Path)'
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    code_folding: hide
    self_contained: true
    number_sections: yes
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: spacelab
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
params:
  data_directory : "/projects/site/pred/SpatialOmics/Experiment_CRC_AllSamples/"
  analysis_name : "analysis_alberto/"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Porportion plots for ST CRC samples from the following publication: 

_Wu, Y. et al. Spatiotemporal Immune Landscape of Colorectal Cancer Liver Metastasis at Single-Cell Level. Cancer Discov. 12, 134–153 (2022)._


# Getting Ready

We first load the libraries and set the paths to the raw data. 

```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(patchwork, lib.loc = "/apps/rocs/2020.08/cascadelake/software/R/4.1.2-foss-2020a/lib64/R/library")
library(readr)
library(dplyr)
library(stringr)
library(purrr)
library(tidyr)
library(cowplot)
library(ggpubr)
library(kableExtra)

data_directory <-  params$data_directory
analysis_name <- params$analysis_name

input_folder <- "Cell2Location/results/LocationModelLinearDependentWMultiExperiment_8experiments_36clusters_31296locations_4188genesLiverMetastasis/"
input_names_1 <- "W_cell_density_q05.csv"


files_to_read_1 <- paste0(data_directory, analysis_name,input_folder,input_names_1)
```

```{r, warning=FALSE, message=FALSE}
results_STLiver_Korean <- read_csv(files_to_read_1) 

colnames(results_STLiver_Korean) <-
  str_replace(colnames(results_STLiver_Korean), 
              pattern = "q05_spot_factors", replacement = "")

results_STLiver_Korean <- results_STLiver_Korean %>% 
  dplyr::mutate(sample = str_extract(.$spot_id, pattern =".*_")) %>%
  dplyr::mutate(sample = str_remove(.$sample, pattern = "_")) %>% 
  dplyr::mutate(spot_id = str_remove(.$spot_id, pattern =".*_")) 
```


# Major Cell types

```{r warning=FALSE, message=FALSE}
## We define the different cell types: 
metadata_Lee_paper <- 
  read_tsv(file = paste0(data_directory, analysis_name, 
    "Cell2Location/inputs/scRNAseq-ref/raw/metadata.tsv"))

epithelial_cells <- metadata_Lee_paper %>% 
  dplyr::filter(Cell_type=="Epithelial cells")


tumor_epithelial_cells <- epithelial_cells %>%
  dplyr::filter(Cell_subtype %in% c("CMS1","CMS2","CMS3","CMS4")) %>%
  dplyr::select(Cell_type,Cell_subtype ) %>% distinct() 
tumor_epithelial_cells$Cell_type <- "Tumor Cells"

normal_epithelial_cells <- epithelial_cells %>%
  dplyr::filter(!(Cell_subtype %in% c("CMS1","CMS2","CMS3","CMS4")))  %>%
  dplyr::select(Cell_type,Cell_subtype ) %>% distinct() 

stromal_cells <- metadata_Lee_paper %>% 
  dplyr::filter(Cell_type=="Stromal cells") %>%
  dplyr::select(Cell_type,Cell_subtype ) %>% distinct()

myeloid_cells <- metadata_Lee_paper %>% 
  dplyr::filter(Cell_type=="Myeloids") %>%
  dplyr::select(Cell_type,Cell_subtype ) %>% distinct()

T_cells <- metadata_Lee_paper %>% 
  dplyr::filter(Cell_type=="T cells") %>%
  dplyr::select(Cell_type,Cell_subtype ) %>% distinct()

B_cells <- metadata_Lee_paper %>% 
  dplyr::filter(Cell_type=="B cells") %>%
  dplyr::select(Cell_type,Cell_subtype ) %>% distinct()

Mast_cells <- metadata_Lee_paper %>% 
  dplyr::filter(Cell_type=="Mast cells") %>%
  dplyr::select(Cell_type,Cell_subtype ) %>% distinct()

MainCellTypes_labels <- rbind(tumor_epithelial_cells, 
                              normal_epithelial_cells, 
                              stromal_cells, 
                              myeloid_cells,
                              T_cells, B_cells, Mast_cells) 

```

```{r, warning=FALSE, message=FALSE}
results_STLiver_Korean_mainCellLines <- results_STLiver_Korean %>%
  tidyr::pivot_longer(!c(spot_id,sample), names_to = "Cell_subtype", values_to = "cell_density_q05") %>%
  dplyr::left_join(MainCellTypes_labels, by="Cell_subtype") %>%
  dplyr::select(sample, cell_density_q05, Cell_type) %>%
  dplyr::group_by(sample, Cell_type) %>%
  summarise_all(funs(sum)) 



results_STLiver_Korean_mainCellLines <- 
  results_STLiver_Korean_mainCellLines %>% 
  dplyr::mutate(sample = case_when(sample == "ST-colon1" ~ "ST-colon1_Unt",
                                   sample =="ST-colon2" ~ "ST-colon2_Unt",
                                   sample =="ST-colon3" ~ "ST-colon3_Tre",
                                   sample =="ST-colon4" ~ "ST-colon4_Tre",
                                   sample =="ST-liver1" ~ "ST-liver1_Unt", 
                                   sample =="ST-liver2" ~ "ST-liver2_Unt", 
                                   sample =="ST-liver3" ~ "ST-liver3_Tre", 
                                   sample =="ST-liver4" ~ "ST-liver4_Tre")) 



  
  
results_STLiver_Korean_mainCellLines$sample <- 
  factor(results_STLiver_Korean_mainCellLines$sample, 
         levels = sort(unique(results_STLiver_Korean_mainCellLines$sample), decreasing = TRUE))


p1 <- ggplot(results_STLiver_Korean_mainCellLines,aes(x = sample , y= cell_density_q05, fill = Cell_type)) + 
  geom_bar(stat="identity", position = "fill", width = 0.75, col ="Black") + 
  coord_flip() + 
  theme_minimal() + 
  scale_fill_brewer(palette = "Set1") + 
theme( # remove the vertical grid lines
    panel.grid.major = element_blank(), 
    axis.title = element_blank(), 
    axis.text.x = element_text(size=12, face = "bold", family="Arial", angle = 90, vjust = -0.1),
    axis.text.y = element_text(size=12, face = "bold", family="Arial"),
    plot.title =  element_text(face="bold", size=14, hjust = 0.5)) + 
  labs(title ="Major Cell \n\ type Proportions", x ="Sample") + 
  theme(legend.position ="bottom",  
        axis.title = element_blank(), 
        legend.text = element_text(size=12, face="bold", family="Arial"), 
        legend.title = element_blank()) + 
  scale_y_continuous(position = "left", breaks = c(0,0.5,1)) + 
  guides(fill=guide_legend(label.position ="left", ncol = 4))



### Number of spots under tissue per sample
df_spots_sample <- results_STLiver_Korean %>% dplyr::count(sample ) %>% 
  dplyr::mutate(sample = case_when(sample == "ST-colon1" ~ "ST-colon1_Unt",
                                   sample =="ST-colon2" ~ "ST-colon2_Unt",
                                   sample =="ST-colon3" ~ "ST-colon3_Tre",
                                   sample =="ST-colon4" ~ "ST-colon4_Tre",
                                   sample =="ST-liver1" ~ "ST-liver1_Unt", 
                                   sample =="ST-liver2" ~ "ST-liver2_Unt", 
                                   sample =="ST-liver3" ~ "ST-liver3_Tre", 
                                   sample =="ST-liver4" ~ "ST-liver4_Tre")) 
  
  
df_spots_sample$sample <- 
  factor(df_spots_sample$sample , 
         levels = sort(unique(df_spots_sample$sample), decreasing = TRUE))

p2 <- df_spots_sample %>% ggplot(aes(x=n, y=sample)) + geom_point(size=3, col="black") +
  scale_x_continuous(trans = "log2", position = "bottom", breaks = c(3600,4096)) + 
  theme(axis.title= element_blank(), axis.text.y = element_blank(), 
        axis.ticks = element_blank(), panel.border = element_blank(),
        panel.grid.minor = element_blank(), panel.grid.major.y =  element_blank(),
        panel.grid.major.x = element_line(size=1, linetype=2, colour="Darkred"), 
        plot.title = element_text(size=14, face = "bold", family="Arial", hjust = 0.5), 
        axis.text.x = element_text(size=12, face = "bold", family="Arial", angle = 90, hjust = 1, vjust = 1)) + 
  labs(title = "# Spots \n\ Under Tissue") 


legend_p1 <- get_legend(p1)

p1_nolegend <- p1 + theme(legend.position = "none")

p1_nolegend + plot_spacer() + p2 + ggpubr::as_ggplot(legend_p1)+ plot_layout(widths = c(5,-0.35,2), heights = c(5,1))
```

```{r, warning=FALSE, message=FALSE}
results_STLiver_Korean_mainCellLines %>% 
  dplyr::group_by(sample) %>% 
  dplyr::mutate(Proportions = 
    round(cell_density_q05  / sum(cell_density_q05 ) * 100, 
    digits = 2)) %>% 
  dplyr::ungroup() %>% 
  dplyr::arrange(desc(sample), desc(Proportions)) %>% 
  kbl() %>% kable_styling()
```


# CMS Tumors 

```{r, warning=FALSE, message=FALSE}
results_STLiver_Korean_CMS <- results_STLiver_Korean %>%
  tidyr::pivot_longer(!c(spot_id,sample), names_to = "Cell_subtype", values_to = "cell_density_q05") %>%
  dplyr::inner_join(tumor_epithelial_cells, by="Cell_subtype") %>%
  dplyr::select(sample, cell_density_q05, Cell_subtype) %>%
  dplyr::group_by(sample, Cell_subtype) %>%
  summarise_all(funs(sum))


 results_STLiver_Korean_CMS <- results_STLiver_Korean_CMS %>% 
  dplyr::mutate(sample = case_when(sample == "ST-colon1" ~ "ST-colon1_Unt",
                                   sample =="ST-colon2" ~ "ST-colon2_Unt",
                                   sample =="ST-colon3" ~ "ST-colon3_Tre",
                                   sample =="ST-colon4" ~ "ST-colon4_Tre",
                                   sample =="ST-liver1" ~ "ST-liver1_Unt", 
                                   sample =="ST-liver2" ~ "ST-liver2_Unt", 
                                   sample =="ST-liver3" ~ "ST-liver3_Tre", 
                                   sample =="ST-liver4" ~ "ST-liver4_Tre"))
 
 results_STLiver_Korean_CMS$sample <- 
  factor(results_STLiver_Korean_CMS$sample , 
         levels = sort(unique(results_STLiver_Korean_CMS$sample), decreasing = TRUE))
 
 p1_CMS <- ggplot(results_STLiver_Korean_CMS,aes(x = sample, y= cell_density_q05, fill = Cell_subtype)) + 
  geom_bar(stat="identity", position = "fill", width = 0.75, col ="Black") + 
  coord_flip() + 
  theme_minimal() + 
  scale_fill_brewer(palette = "Set1") + 
  theme( # remove the vertical grid lines
    panel.grid.major = element_blank(), 
    axis.title = element_blank(), 
    axis.text.x = element_text(size=12, face = "bold", family="Arial", angle = 90, vjust = -0.1),
    axis.text.y = element_text(size=12, face = "bold", family="Arial"),
    plot.title =  element_text(face="bold", size=14, hjust = 0.5)) + 
  labs(title ="CMSs \n\ Proportions", x ="Sample") + 
  theme(legend.position ="bottom",  
        axis.title = element_blank(), 
        legend.text = element_text(size=12, face="bold", family="Arial"), 
        legend.title = element_blank()) + 
  scale_y_continuous(position = "left", breaks = c(0,0.5,1)) + 
  guides(fill=guide_legend(label.position ="left", ncol = 2))

### Compute the total number of spots with CMS signature (more than 20% of cells
### in the spot are CMS)

Total_content_spot_df <- results_STLiver_Korean %>%
  tidyr::pivot_longer(!c(spot_id,sample), names_to = "Cell_subtype", values_to = "cell_density_q05") %>%
  dplyr::group_by(spot_id, sample) %>%
  dplyr::summarise(Total_Content_spot= sum(cell_density_q05))  
  # dplyr::mutate(sample = str_remove(sample, "^[^A]+"))

CMS_content_spot_df <- results_STLiver_Korean %>%
  tidyr::pivot_longer(!c(spot_id,sample), names_to = "Cell_subtype", values_to = "cell_density_q05") %>%
  dplyr::filter(Cell_subtype %in% tumor_epithelial_cells$Cell_subtype)  %>%
  dplyr::group_by(spot_id, sample) %>%
  dplyr::summarise(CMS_Content_spot= sum(cell_density_q05))  

CMS_percentage_perSpot_df <- Total_content_spot_df %>%
  dplyr::left_join(CMS_content_spot_df) %>%
  dplyr::mutate(ratio = CMS_Content_spot/Total_Content_spot ) %>%
  dplyr::filter(ratio > 0.20) %>% ungroup() %>%
  dplyr::count(sample) %>%  
  dplyr::mutate(sample = case_when(sample == "ST-colon1" ~ "ST-colon1_Unt",
                                   sample =="ST-colon2" ~ "ST-colon2_Unt",
                                   sample =="ST-colon3" ~ "ST-colon3_Tre",
                                   sample =="ST-colon4" ~ "ST-colon4_Tre",
                                   sample =="ST-liver1" ~ "ST-liver1_Unt", 
                                   sample =="ST-liver2" ~ "ST-liver2_Unt", 
                                   sample =="ST-liver3" ~ "ST-liver3_Tre", 
                                   sample =="ST-liver4" ~ "ST-liver4_Tre"))
 
  
CMS_percentage_perSpot_df$sample <- 
  factor(CMS_percentage_perSpot_df$sample , 
         levels = sort(unique(CMS_percentage_perSpot_df$sample), decreasing = TRUE))

p2_CMS <- CMS_percentage_perSpot_df %>% ggplot(aes(x=n, y=sample)) + geom_point(size=3, col="black") +
  scale_x_continuous(trans = "log2", position = "bottom") + 
  theme(axis.title= element_blank(), axis.text.y = element_blank(), 
        axis.ticks = element_blank(), panel.border = element_blank(),
        panel.grid.minor = element_blank(), panel.grid.major.y =  element_blank(),
        panel.grid.major.x = element_line(size=1, linetype=2, colour="Darkred"), 
        plot.title = element_text(size=14, face = "bold", family="Arial", hjust = 0.5), 
        axis.text.x = element_text(size=12, face = "bold", family="Arial", angle = 90, hjust = 1, vjust = 1)) + 
  labs(title="# CMS \n\ spots") 
  
legend_p1_CMS <- get_legend(p1_CMS)

p1_CMS_nolegend <- p1_CMS + theme(legend.position = "none")

CMS_plots <- p1_CMS_nolegend + p2_CMS + ggpubr::as_ggplot(legend_p1_CMS)+ plot_layout(widths = c(2.5, 1), heights = c(5,1)) + theme(axis.text.y = element_blank())
```


```{r, warning=FALSE, message=FALSE}
p1_CMS + plot_spacer() + p2_CMS + plot_layout(widths = c(5, -0.35 ,2),guides = "collect")& theme(legend.position = "bottom")
```


```{r, warning=FALSE, message=FALSE}
results_STLiver_Korean_CMS %>% 
  dplyr::group_by(sample) %>% 
  dplyr::mutate(Proportions = 
    round(cell_density_q05  / sum(cell_density_q05 ) * 100, 
    digits = 2)) %>% 
  dplyr::ungroup() %>% 
  dplyr::arrange(desc(sample), desc(Proportions)) %>% 
  kbl() %>% kable_styling()
```
```{r}
p1_nolegend + plot_spacer() + p2 + 
  p1_CMS_nolegend + theme(axis.text.y = element_blank()) + plot_spacer() + 
  p2_CMS + 
  ggpubr::as_ggplot(legend_p1)+ plot_spacer() + plot_spacer() +  
  ggpubr::as_ggplot(legend_p1_CMS) + 
  plot_layout(widths = c(5,-0.35,2,5,-0.35,2), heights = c(5,1))
```


# T cells

```{r, warning=FALSE, message=FALSE}
results_STLiver_Korean_T_cells  <- results_STLiver_Korean %>%
  tidyr::pivot_longer(!c(spot_id,sample), names_to = "Cell_subtype", values_to = "cell_density_q05") %>%
  dplyr::inner_join(T_cells, by="Cell_subtype") %>%
  dplyr::filter(Cell_subtype != "Unknown") %>%
  dplyr::select(sample, cell_density_q05, Cell_subtype) %>%
  dplyr::group_by(sample, Cell_subtype) %>%
  summarise_all(funs(sum))  

 results_STLiver_Korean_T_cells <- results_STLiver_Korean_T_cells %>% 
  dplyr::mutate(sample = case_when(sample == "ST-colon1" ~ "ST-colon1_Unt",
                                   sample =="ST-colon2" ~ "ST-colon2_Unt",
                                   sample =="ST-colon3" ~ "ST-colon3_Tre",
                                   sample =="ST-colon4" ~ "ST-colon4_Tre",
                                   sample =="ST-liver1" ~ "ST-liver1_Unt", 
                                   sample =="ST-liver2" ~ "ST-liver2_Unt", 
                                   sample =="ST-liver3" ~ "ST-liver3_Tre", 
                                   sample =="ST-liver4" ~ "ST-liver4_Tre"))
 
 results_STLiver_Korean_T_cells$sample <- 
  factor(results_STLiver_Korean_T_cells$sample , 
         levels = sort(unique(results_STLiver_Korean_T_cells$sample), decreasing = TRUE))
 
 p1_T_cells <- ggplot(results_STLiver_Korean_T_cells,aes(x = sample , y= cell_density_q05, fill = Cell_subtype)) + 
  geom_bar(stat="identity", position = "fill", width = 0.75, col ="Black") + 
  coord_flip() + 
  theme_minimal() + 
  scale_fill_brewer(palette = "Set1") + 
  theme( # remove the vertical grid lines
    panel.grid.major = element_blank(), 
    axis.title = element_blank(), 
    axis.text.x = element_text(size=12, face = "bold", family="Arial", angle = 90, vjust = -0.1),
    axis.text.y = element_text(size=12, face = "bold", family="Arial"),
    plot.title =  element_text(face="bold", size=14, hjust = 0.5)) + 
  labs(title ="T-Cells \n\ Proportions", x ="Sample") + 
  theme(legend.position ="bottom",  
        axis.title = element_blank(), 
        legend.text = element_text(size=12, face="bold", family="Arial"), 
        legend.title = element_blank()) + 
  scale_y_continuous(position = "left", breaks = c(0,0.5,1)) + 
  guides(fill=guide_legend(label.position ="left", ncol = 3))



### Compute the total number of spots with CMS signature (more than 25% of cells
### in the spot are CMS)

Total_content_spot_df <- results_STLiver_Korean %>%
  tidyr::pivot_longer(!c(spot_id,sample), names_to = "Cell_subtype", values_to = "cell_density_q05") %>%
  dplyr::group_by(spot_id, sample) %>%
  dplyr::summarise(Total_Content_spot= sum(cell_density_q05))  


Tcells_content_spot_df <- results_STLiver_Korean %>%
  tidyr::pivot_longer(!c(spot_id,sample), names_to = "Cell_subtype", values_to = "cell_density_q05") %>%
  dplyr::filter(Cell_subtype %in% T_cells$Cell_subtype)  %>%
  dplyr::group_by(spot_id, sample) %>%
  dplyr::summarise(Tcell_Content_spot= sum(cell_density_q05))  

Tcells_percentage_perSpot_df <- Total_content_spot_df %>%
  dplyr::left_join(Tcells_content_spot_df) %>%
  dplyr::mutate(ratio = Tcell_Content_spot/Total_Content_spot ) %>%
  dplyr::filter(ratio > 0.20) %>% ungroup() %>%
  dplyr::count(sample) %>%  
  dplyr::mutate(sample = case_when(sample == "ST-colon1" ~ "ST-colon1_Unt",
                                   sample =="ST-colon2" ~ "ST-colon2_Unt",
                                   sample =="ST-colon3" ~ "ST-colon3_Tre",
                                   sample =="ST-colon4" ~ "ST-colon4_Tre",
                                   sample =="ST-liver1" ~ "ST-liver1_Unt", 
                                   sample =="ST-liver2" ~ "ST-liver2_Unt", 
                                   sample =="ST-liver3" ~ "ST-liver3_Tre", 
                                   sample =="ST-liver4" ~ "ST-liver4_Tre"))

Tcells_percentage_perSpot_df$sample <- 
  factor(Tcells_percentage_perSpot_df$sample , 
         levels = sort(unique(Tcells_percentage_perSpot_df$sample), decreasing = TRUE))

p2_Tcells <- Tcells_percentage_perSpot_df %>% ggplot(aes(x=n, y=sample)) + geom_point(size=3, col="black") +
  scale_x_continuous(trans = "log2", position = "bottom", breaks = c(64,256,1024)) + 
  theme(axis.title= element_blank(), axis.text.y = element_blank(), 
        axis.ticks = element_blank(), panel.border = element_blank(),
        panel.grid.minor = element_blank(), panel.grid.major.y =  element_blank(),
        panel.grid.major.x = element_line(size=1, linetype=2, colour="Darkred"), 
        plot.title = element_text(size=14, face = "bold", family="Arial", hjust = 0.5), 
        axis.text.x = element_text(size=12, face = "bold", family="Arial", angle = 90, hjust = 1, vjust = 1)) + 
  labs(title="# T-cells \n\ spots") 


legend_T_cells <- get_legend(p1_T_cells)

p1_T_cells_nolegend <- p1_T_cells + theme(legend.position = "none") 

T_cells_plots <- p1_T_cells_nolegend + p2_Tcells + ggpubr::as_ggplot(legend_T_cells)+ plot_layout(widths = c(2.5,1), heights = c(5,1))

p1_T_cells_nolegend + plot_spacer() + p2_Tcells + 
  plot_layout(widths = c(5, -0.35 ,2),guides = "collect")& theme(legend.position = "bottom")

```

# B_cells 

```{r, warning=FALSE, message=FALSE}
results_STLiver_Korean_B_cells  <- results_STLiver_Korean %>%
  tidyr::pivot_longer(!c(spot_id,sample), names_to = "Cell_subtype", values_to = "cell_density_q05") %>%
  dplyr::inner_join(B_cells, by="Cell_subtype") %>%
  dplyr::filter(Cell_subtype != "Unknown") %>%
  dplyr::select(sample, cell_density_q05, Cell_subtype) %>%
  dplyr::group_by(sample, Cell_subtype) %>%
  summarise_all(funs(sum))  
  

results_STLiver_Korean_B_cells <- 
  results_STLiver_Korean_B_cells %>% 
  dplyr::mutate(sample = case_when(sample == "ST-colon1" ~ "ST-colon1_Unt",
                                   sample =="ST-colon2" ~ "ST-colon2_Unt",
                                   sample =="ST-colon3" ~ "ST-colon3_Tre",
                                   sample =="ST-colon4" ~ "ST-colon4_Tre",
                                   sample =="ST-liver1" ~ "ST-liver1_Unt", 
                                   sample =="ST-liver2" ~ "ST-liver2_Unt", 
                                   sample =="ST-liver3" ~ "ST-liver3_Tre", 
                                   sample =="ST-liver4" ~ "ST-liver4_Tre"))

  

results_STLiver_Korean_B_cells$sample <- 
  factor(results_STLiver_Korean_B_cells$sample , 
         levels = sort(unique(results_STLiver_Korean_B_cells$sample), decreasing = TRUE))


p1_B_cells <- ggplot(results_STLiver_Korean_B_cells,aes(x = sample , y= cell_density_q05, fill = Cell_subtype)) + 
  geom_bar(stat="identity", position = "fill", width = 0.75, col ="Black") + 
  coord_flip() + 
  theme_minimal() + 
  scale_fill_brewer(palette = "Set1") + 
  theme( # remove the vertical grid lines
    panel.grid.major = element_blank(), 
    axis.title = element_blank(), 
    axis.text.x = element_text(size=12, face = "bold", family="Arial", angle = 90, vjust = -0.1),
    axis.text.y = element_text(size=12, face = "bold", family="Arial"),
    plot.title =  element_text(face="bold", size=14, hjust = 0.5)) + 
  labs(title ="B-Cells \n\ Proportions", x ="Sample") + 
  theme(legend.position ="bottom",  
        axis.title = element_blank(), 
        legend.text = element_text(size=12, face="bold", family="Arial"), 
        legend.title = element_blank()) + 
  scale_y_continuous(position = "left", breaks = c(0,0.5,1)) + 
  guides(fill=guide_legend(label.position ="left", ncol = 2))



### Compute the total number of spots with CMS signature (more than 25% of cells
### in the spot are CMS)

Total_content_spot_df <- results_STLiver_Korean %>%
  tidyr::pivot_longer(!c(spot_id,sample), names_to = "Cell_subtype", values_to = "cell_density_q05") %>%
  dplyr::group_by(spot_id, sample) %>%
  dplyr::summarise(Total_Content_spot= sum(cell_density_q05))  
# dplyr::mutate(sample = str_remove(sample, "^[^A]+"))

Bcells_content_spot_df <- results_STLiver_Korean %>%
  tidyr::pivot_longer(!c(spot_id,sample), names_to = "Cell_subtype", values_to = "cell_density_q05") %>%
  dplyr::filter(Cell_subtype %in% B_cells$Cell_subtype)  %>%
  dplyr::group_by(spot_id, sample) %>%
  dplyr::summarise(Bcell_Content_spot= sum(cell_density_q05))  

Bcells_percentage_perSpot_df <- Total_content_spot_df %>%
  dplyr::left_join(Bcells_content_spot_df) %>%
  dplyr::mutate(ratio = Bcell_Content_spot/Total_Content_spot ) %>%
  dplyr::filter(ratio > 0.20) %>% ungroup() %>%
  dplyr::count(sample) %>%  
  dplyr::mutate(sample = case_when(sample == "ST-colon1" ~ "ST-colon1_Unt",
                                   sample =="ST-colon2" ~ "ST-colon2_Unt",
                                   sample =="ST-colon3" ~ "ST-colon3_Tre",
                                   sample =="ST-colon4" ~ "ST-colon4_Tre",
                                   sample =="ST-liver1" ~ "ST-liver1_Unt", 
                                   sample =="ST-liver2" ~ "ST-liver2_Unt", 
                                   sample =="ST-liver3" ~ "ST-liver3_Tre", 
                                   sample =="ST-liver4" ~ "ST-liver4_Tre"))

  

Bcells_percentage_perSpot_df$sample <- 
  factor(Bcells_percentage_perSpot_df$sample , 
         levels = sort(unique(Bcells_percentage_perSpot_df$sample), decreasing = TRUE))

p2_Bcells <- Bcells_percentage_perSpot_df %>% ggplot(aes(x=n, y=sample)) + geom_point(size=3, col="black") +
  scale_x_continuous(trans = "log2", position = "bottom") + 
  theme(axis.title= element_blank(), axis.text.y = element_blank(), 
        axis.ticks = element_blank(), panel.border = element_blank(),
        panel.grid.minor = element_blank(), panel.grid.major.y =  element_blank(),
        panel.grid.major.x = element_line(size=1, linetype=2, colour="Darkred"), 
        plot.title = element_text(size=14, face = "bold", family="Arial", hjust = 0.5), 
        axis.text.x = element_text(size=12, face = "bold", family="Arial", angle = 90, hjust = 1, vjust = 1)) + 
  labs(title="# B-cells \n\ spots") 



legend_B_cells <- get_legend(p1_B_cells)

p1_B_cells_nolegend <- p1_B_cells + theme(legend.position = "none") 

p1_B_cells_nolegend + p2_Bcells + ggpubr::as_ggplot(legend_B_cells)+ plot_layout(widths = c(2.5, 1), heights = c(5,1))
```



# Myeloid cells

```{r, warning=FALSE, message=FALSE}
results_STLiver_Korean_Myeloid_cells  <- results_STLiver_Korean %>%
  tidyr::pivot_longer(!c(spot_id,sample), names_to = "Cell_subtype", values_to = "cell_density_q05") %>%
  dplyr::inner_join(myeloid_cells, by="Cell_subtype") %>%
  dplyr::filter(Cell_subtype != "Unknown") %>%
  dplyr::select(sample, cell_density_q05, Cell_subtype) %>%
  dplyr::group_by(sample, Cell_subtype) %>%
  summarise_all(funs(sum))  
  


results_STLiver_Korean_Myeloid_cells <- 
  results_STLiver_Korean_Myeloid_cells %>% 
  dplyr::mutate(sample = case_when(sample == "ST-colon1" ~ "ST-colon1_Unt",
                                   sample =="ST-colon2" ~ "ST-colon2_Unt",
                                   sample =="ST-colon3" ~ "ST-colon3_Tre",
                                   sample =="ST-colon4" ~ "ST-colon4_Tre",
                                   sample =="ST-liver1" ~ "ST-liver1_Unt", 
                                   sample =="ST-liver2" ~ "ST-liver2_Unt", 
                                   sample =="ST-liver3" ~ "ST-liver3_Tre", 
                                   sample =="ST-liver4" ~ "ST-liver4_Tre"))

  
  
results_STLiver_Korean_Myeloid_cells$sample <- 
  factor(results_STLiver_Korean_Myeloid_cells$sample , 
         levels = sort(unique(results_STLiver_Korean_Myeloid_cells$sample), decreasing = TRUE))

p1_Myeloid_cells <- ggplot(results_STLiver_Korean_Myeloid_cells,aes(x = sample , y= cell_density_q05, fill = Cell_subtype)) + 
  geom_bar(stat="identity", position = "fill", width = 0.75, col ="Black") + 
  coord_flip() + 
  theme_minimal() + 
  scale_fill_brewer(palette = "Set1") + 
  theme( # remove the vertical grid lines
    panel.grid.major = element_blank(), 
    axis.title = element_blank(), 
    axis.text.x = element_text(size=12, face = "bold", family="Arial", angle = 90, vjust = -0.1),
    axis.text.y = element_text(size=12, face = "bold", family="Arial"),
    plot.title =  element_text(face="bold", size=14, hjust = 0.5)) + 
  labs(title ="Myeloid Cells \n\ Proportions", x ="Sample") + 
  theme(legend.position ="bottom",  
        axis.title = element_blank(), 
        legend.text = element_text(size=12, face="bold", family="Arial"), 
        legend.title = element_blank()) + 
  scale_y_continuous(position = "left", breaks = c(0,0.5,1)) + 
  guides(fill=guide_legend(label.position ="left", ncol = 2))


### Compute the total number of spots with CMS signature (more than 25% of cells
### in the spot are CMS)

Total_content_spot_df <- results_STLiver_Korean %>%
  tidyr::pivot_longer(!c(spot_id,sample), names_to = "Cell_subtype", values_to = "cell_density_q05") %>%
  dplyr::group_by(spot_id, sample) %>%
  dplyr::summarise(Total_Content_spot= sum(cell_density_q05))  
# dplyr::mutate(sample = str_remove(sample, "^[^A]+"))

MyeloidCells_content_spot_df <- results_STLiver_Korean %>%
  tidyr::pivot_longer(!c(spot_id,sample), names_to = "Cell_subtype", values_to = "cell_density_q05") %>%
  dplyr::filter(Cell_subtype %in% myeloid_cells$Cell_subtype)  %>%
  dplyr::group_by(spot_id, sample) %>%
  dplyr::summarise(Bcell_Content_spot= sum(cell_density_q05))  

MyeloidCells_percentage_perSpot_df <- Total_content_spot_df %>%
  dplyr::left_join(MyeloidCells_content_spot_df) %>%
  dplyr::mutate(ratio = Bcell_Content_spot/Total_Content_spot ) %>%
  dplyr::filter(ratio > 0.20) %>% ungroup() %>%
  dplyr::count(sample) %>%  
  dplyr::mutate(sample = case_when(sample == "ST-colon1" ~ "ST-colon1_Unt",
                                   sample =="ST-colon2" ~ "ST-colon2_Unt",
                                   sample =="ST-colon3" ~ "ST-colon3_Tre",
                                   sample =="ST-colon4" ~ "ST-colon4_Tre",
                                   sample =="ST-liver1" ~ "ST-liver1_Unt", 
                                   sample =="ST-liver2" ~ "ST-liver2_Unt", 
                                   sample =="ST-liver3" ~ "ST-liver3_Tre", 
                                   sample =="ST-liver4" ~ "ST-liver4_Tre"))

  
  
MyeloidCells_percentage_perSpot_df$sample <- 
  factor(MyeloidCells_percentage_perSpot_df$sample , 
         levels = sort(unique(MyeloidCells_percentage_perSpot_df$sample), decreasing = TRUE))

p2_MyeloidCells <- MyeloidCells_percentage_perSpot_df %>% ggplot(aes(x=n, y=sample)) + geom_point(size=3, col="black") +
  scale_x_continuous(trans = "log2", position = "bottom") + 
  theme(axis.title= element_blank(), axis.text.y = element_blank(), 
        axis.ticks = element_blank(), panel.border = element_blank(),
        panel.grid.minor = element_blank(), panel.grid.major.y =  element_blank(),
        panel.grid.major.x = element_line(size=1, linetype=2, colour="Darkred"), 
        plot.title = element_text(size=14, face = "bold", family="Arial", hjust = 0.5), 
        axis.text.x = element_text(size=12, face = "bold", family="Arial", angle = 90, hjust = 1, vjust = 1)) + 
  labs(title="# Myeloid \n\ cells spots") 


legend_Myeloid_cells <- get_legend(p1_Myeloid_cells)

p1_Myeloid_cells_nolegend <- p1_Myeloid_cells + theme(legend.position = "none") 

p1_Myeloid_cells_nolegend + p2_MyeloidCells + ggpubr::as_ggplot(legend_Myeloid_cells)+ plot_layout(widths = c(2.5, 1), heights = c(5,1))

```


# Main Stromal cells. 

```{r, warning=FALSE, message=FALSE}
results_STLiver_Korean_stromal_cells  <- results_STLiver_Korean %>%
  tidyr::pivot_longer(!c(spot_id,sample), names_to = "Cell_subtype", values_to = "cell_density_q05") %>%
  dplyr::filter(Cell_subtype %in% c("Myofibroblasts","Stromal 1","Stromal 2","Stromal 3")) %>%
  dplyr::filter(Cell_subtype != "Unknown") %>%
  dplyr::select(sample, cell_density_q05, Cell_subtype) %>%
  dplyr::group_by(sample, Cell_subtype) %>%
  summarise_all(funs(sum))  
  

results_STLiver_Korean_stromal_cells <- 
  results_STLiver_Korean_stromal_cells %>% 
  dplyr::mutate(sample = case_when(sample == "ST-colon1" ~ "ST-colon1_Unt",
                                   sample =="ST-colon2" ~ "ST-colon2_Unt",
                                   sample =="ST-colon3" ~ "ST-colon3_Tre",
                                   sample =="ST-colon4" ~ "ST-colon4_Tre",
                                   sample =="ST-liver1" ~ "ST-liver1_Unt", 
                                   sample =="ST-liver2" ~ "ST-liver2_Unt", 
                                   sample =="ST-liver3" ~ "ST-liver3_Tre", 
                                   sample =="ST-liver4" ~ "ST-liver4_Tre"))



results_STLiver_Korean_stromal_cells$sample <- 
  factor(results_STLiver_Korean_stromal_cells$sample , 
         levels = sort(unique(results_STLiver_Korean_stromal_cells$sample), decreasing = TRUE))


p1_stromal_cells <- ggplot(results_STLiver_Korean_stromal_cells,aes(x = sample , y= cell_density_q05, fill = Cell_subtype)) + 
  geom_bar(stat="identity", position = "fill", width = 0.75, col ="Black") + 
  coord_flip() + 
  theme_minimal() + 
  scale_fill_brewer(palette = "Set1") + 
  theme( # remove the vertical grid lines
    panel.grid.major = element_blank(), 
    axis.title = element_blank(), 
    axis.text.x = element_text(size=12, face = "bold", family="Arial", angle = 90, vjust = -0.1),
    axis.text.y = element_text(size=12, face = "bold", family="Arial"),
    plot.title =  element_text(face="bold", size=14, hjust = 0.5)) + 
  labs(title ="Stromal Cells \n\ Proportions", x ="Sample") + 
  theme(legend.position ="bottom",  
        axis.title = element_blank(), 
        legend.text = element_text(size=12, face="bold", family="Arial"), 
        legend.title = element_blank()) + 
  scale_y_continuous(position = "left", breaks = c(0,0.5,1)) + 
  guides(fill=guide_legend(label.position ="left", ncol = 2))

### Compute the total number of spots with CMS signature (more than 25% of cells
### in the spot are CMS)

Total_content_spot_df <- results_STLiver_Korean %>%
  tidyr::pivot_longer(!c(spot_id,sample), names_to = "Cell_subtype", values_to = "cell_density_q05") %>%
  dplyr::group_by(spot_id, sample) %>%
  dplyr::summarise(Total_Content_spot= sum(cell_density_q05))  
# dplyr::mutate(sample = str_remove(sample, "^[^A]+"))

StromalCells_content_spot_df <- results_STLiver_Korean %>%
  tidyr::pivot_longer(!c(spot_id,sample), names_to = "Cell_subtype", values_to = "cell_density_q05") %>%
  dplyr::filter(Cell_subtype %in% c("Myofibroblasts","Stromal 1","Stromal 2","Stromal 3"))  %>%
  dplyr::group_by(spot_id, sample) %>%
  dplyr::summarise(StromalCell_Content_spot= sum(cell_density_q05))  

StromalCells_percentage_perSpot_df <- Total_content_spot_df %>%
  dplyr::left_join(StromalCells_content_spot_df) %>%
  dplyr::mutate(ratio = StromalCell_Content_spot/Total_Content_spot ) %>%
  dplyr::filter(ratio > 0.20) %>% ungroup() %>%
  dplyr::count(sample) %>%  
  dplyr::mutate(sample = case_when(sample == "ST-colon1" ~ "ST-colon1_Unt",
                                   sample =="ST-colon2" ~ "ST-colon2_Unt",
                                   sample =="ST-colon3" ~ "ST-colon3_Tre",
                                   sample =="ST-colon4" ~ "ST-colon4_Tre",
                                   sample =="ST-liver1" ~ "ST-liver1_Unt", 
                                   sample =="ST-liver2" ~ "ST-liver2_Unt", 
                                   sample =="ST-liver3" ~ "ST-liver3_Tre", 
                                   sample =="ST-liver4" ~ "ST-liver4_Tre"))

  
  
StromalCells_percentage_perSpot_df$sample <- 
  factor(StromalCells_percentage_perSpot_df$sample , 
         levels = sort(unique(StromalCells_percentage_perSpot_df$sample), decreasing = TRUE))


p2_StromalCells <- StromalCells_percentage_perSpot_df %>% ggplot(aes(x=n, y=sample)) + geom_point(size=3, col="black") +
  scale_x_continuous(trans = "log2", position = "bottom", breaks = c(8,128,1024)) + 
  theme(axis.title= element_blank(), axis.text.y = element_blank(), 
        axis.ticks = element_blank(), panel.border = element_blank(),
        panel.grid.minor = element_blank(), panel.grid.major.y =  element_blank(),
        panel.grid.major.x = element_line(size=1, linetype=2, colour="Darkred"), 
        plot.title = element_text(size=14, face = "bold", family="Arial", hjust = 0.5), 
        axis.text.x = element_text(size=12, face = "bold", family="Arial", angle = 90, hjust = 1, vjust = 1)) + 
  labs(title="# Stromal \n\ cells spots") 

legend_Stromal_cells <- get_legend(p1_stromal_cells)

p1_stromal_cells_nolegend <- p1_stromal_cells + theme(legend.position = "none") 

p1_stromal_cells_nolegend + p2_StromalCells + ggpubr::as_ggplot(legend_Stromal_cells)+ plot_layout(widths = c(2.5, 1), heights = c(5,1))
```





# Conclusion 


# Session Info Details

```{r, echo=FALSE, eval=TRUE}
sessionInfo()
```