---
title: "Analysis of CRC Spatial Transcriptomics data from 10X: CMS2 Inter-patient heterogeneity"
author: 'Alberto Valdeolivas Urbelz (PS-Biomics&Path)'
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    code_folding: hide
    self_contained: true
    number_sections: yes
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: spacelab
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
params:
  data_directory : "/projects/site/pred/SpatialOmics/Experiment_CRC_AllSamples/"
  analysis_name : "analysis_alberto/"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

The present script takes Spatial Omics data generated with 10X technology
of a set of colorectal cancer samples and explores the inter-patient 
hetereogeneity in those samples that have been classified as CMS2. To do so, 
we will perform pseudo-bulk of the CMS2 regions. 

```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(vctrs)
library(dplyr)
library(patchwork, lib.loc = "/apps/rocs/2020.08/cascadelake/software/R/4.1.2-foss-2020a/lib64/R/library")
library(cancersea)
library(Seurat)
library(readr)
library(stringr)
library(cowplot)
library(msigdbr)
library(clusterProfiler)
library(RColorBrewer)

source(file = "WrapperFunction/SeuratWrappers.R")
source(file = "WrapperFunction/PseudoBulkUtils.R")

data_directory <-  params$data_directory
analysis_name <- params$analysis_name

input_files <- "IntermediaryFiles/"
input_names <- "SeuratList_Clusters_Res05.rds"


files_to_read <- paste0(data_directory, analysis_name,input_files,input_names)
results_directory <- paste0(data_directory, analysis_name, input_files)
```

we read the results of the deconvolution containing the total
number of cells per spot. From that, we compute the cell type proportions per 
spot. 

```{r, warning=FALSE, message=FALSE}
data('available_pathways')
available_pathways
data('angiogenesis')
```

```{r warning=FALSE}
seurat_objects_all <- readRDS(files_to_read)

samples_to_consider <- c("SN048_A121573_Rep1", "SN048_A121573_Rep2",
  "SN123_A595688_Rep1","SN124_A595688_Rep2", "SN123_A938797_Rep1_X", 
  "SN124_A938797_Rep2", "SN84_A120838_Rep1", "SN84_A120838_Rep2") 

seurat_objects <- seurat_objects_all[samples_to_consider]

list_cancerCSA <- list(angiogenesis = angiogenesis$symbol, 
                       apoptosis = apoptosis$symbol, 
                       cell_cycle = cell_cycle$symbol, 
                       differentiation = differentiation$symbol,
                       dna_damage = dna_damage$symbol, 
                       dna_repair = dna_repair$symbol, 
                       emt = emt$symbol, 
                       hypoxia = hypoxia$symbol, 
                       inflammation = inflammation$symbol, 
                       invasion = invasion$symbol,
                       metastasis = metastasis$symbol,
                       proliferation = proliferation$symbol, 
                       quiescence = quiescence$symbol, 
                       stemness = stemness$symbol)


seurat_objects <- lapply(seurat_objects, function(x){
  a <- AddModuleScore(
    object = x,
    features = list_cancerCSA,
    ctrl = 20,
    name = names(list_cancerCSA))
  
  b <- SeuratObject::CreateAssayObject(data = t(a@meta.data[,9:ncol(a@meta.data)])) 
  
  Seurat::Key(b) <- "modules_"
  a[["modules_assay"]] <- b
  return(a)
})
```

```{r, warning=FALSE, message=FALSE}
current_object <- seurat_objects[["SN123_A595688_Rep1"]]
DefaultAssay(current_object) <- "modules_assay"

current_gene <- c("metastasis11")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)[current_gene,]), 
              max(GetAssayData(current_object)[current_gene,]), length.out= 4),1)
A595688_Rep1_Metastasis <- 
  Seurat::SpatialFeaturePlot(
      object = current_object,
      features = current_gene, alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 2.25, stroke = 1) + 
      ggtitle("Score Metastasis") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL ) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(NULL),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5))#  +
        # theme(plot.margin = unit(c(0, 0, 0, 0), "cm"),
        #      axis.title.y = element_text(face ="bold", size=14)) # + 
      # ylab("HBEGF Expression")
```


```{r, warning=FALSE, message=FALSE, dpi=100}
A595688_Rep1_Metastasis
```
```{r, warning=FALSE, message=FALSE}
current_object <- seurat_objects[["SN124_A938797_Rep2"]]
DefaultAssay(current_object) <- "modules_assay"

current_gene <- c("metastasis11")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)[current_gene,]), 
              max(GetAssayData(current_object)[current_gene,]), length.out= 4),1)
A938797_Rep2_Metastasis <- 
  Seurat::SpatialFeaturePlot(
      object = current_object,
      features = current_gene, alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 2.25, stroke = 1) + 
      ggtitle("Score Metastasis") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL ) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(NULL),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5))#  +
        # theme(plot.margin = unit(c(0, 0, 0, 0), "cm"),
        #      axis.title.y = element_text(face ="bold", size=14)) # + 
      # ylab("HBEGF Expression")
```


```{r, warning=FALSE, message=FALSE, dpi=100}
A938797_Rep2_Metastasis
```



```{r, warning=FALSE, message=FALSE}
current_object <- seurat_objects[["SN84_A120838_Rep2"]]
DefaultAssay(current_object) <- "modules_assay"

current_gene <- c("angiogenesis1")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)[current_gene,]), 
              max(GetAssayData(current_object)[current_gene,]), length.out= 4),1)
A120838_Rep2_angiogenesis <- 
  Seurat::SpatialFeaturePlot(
      object = current_object,
      features = current_gene, alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 2.25, stroke = 1) + 
      ggtitle("Score Angiogenesis") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL ) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(NULL),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5))#  +
        # theme(plot.margin = unit(c(0, 0, 0, 0), "cm"),
        #      axis.title.y = element_text(face ="bold", size=14)) # + 
      # ylab("HBEGF Expression")
```

```{r, warning=FALSE, message=FALSE, dpi=100}
A120838_Rep2_angiogenesis
```
```{r, warning=FALSE, message=FALSE}
current_object <- seurat_objects[["SN84_A120838_Rep2"]]
DefaultAssay(current_object) <- "modules_assay"

current_gene <- c("emt7")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)[current_gene,]), 
              max(GetAssayData(current_object)[current_gene,]), length.out= 4),1)
A120838_Rep2_Angiogenesis <- 
  Seurat::SpatialFeaturePlot(
      object = current_object,
      features = current_gene, alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 2.25, stroke = 1) + 
      ggtitle("Score EMT") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL ) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(NULL),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5))#  +
        # theme(plot.margin = unit(c(0, 0, 0, 0), "cm"),
        #      axis.title.y = element_text(face ="bold", size=14)) # + 
      # ylab("HBEGF Expression")
```

```{r, warning=FALSE, message=FALSE, dpi=100}
A120838_Rep2_Angiogenesis
```

```{r, warning=FALSE, message=FALSE}
current_object <- seurat_objects[["SN124_A938797_Rep2"]]
DefaultAssay(current_object) <- "modules_assay"

current_gene <- c("emt7")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)[current_gene,]), 
              max(GetAssayData(current_object)[current_gene,]), length.out= 4),1)
A938797_Rep2_EMT <- 
  Seurat::SpatialFeaturePlot(
      object = current_object,
      features = current_gene, alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 2.25, stroke = 1) + 
      ggtitle("Score EMT") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL ) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(NULL),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5))#  +
        # theme(plot.margin = unit(c(0, 0, 0, 0), "cm"),
        #      axis.title.y = element_text(face ="bold", size=14)) # + 
      # ylab("HBEGF Expression")
```
```{r, warning=FALSE, message=FALSE}
current_gene <- c("metastasis11")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)[current_gene,]), 
              max(GetAssayData(current_object)[current_gene,]), length.out= 4),1)
A938797_Rep2_metastasis <- 
  Seurat::SpatialFeaturePlot(
      object = current_object,
      features = current_gene, alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 2.25, stroke = 1) + 
      ggtitle("Score Metastasis") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL ) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(NULL),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5))#  +
        # theme(plot.margin = unit(c(0, 0, 0, 0), "cm"),
        #      axis.title.y = element_text(face ="bold", size=14)) # + 
      # ylab("HBEGF Expression")
```


```{r, warning=FALSE, message=FALSE}
current_gene <- c("invasion10")
breaks_vec <- 
    round(seq(min(GetAssayData(current_object)[current_gene,]), 
              max(GetAssayData(current_object)[current_gene,]), length.out= 4),1)
A938797_Rep2_invasion <- 
  Seurat::SpatialFeaturePlot(
      object = current_object,
      features = current_gene, alpha = c(0,1), 
      image.alpha = 0.6, pt.size.factor = 2.25, stroke = 1) + 
      ggtitle("Score Invasion") + 
      scale_fill_gradient(high = "#67000D", low = "white", 
                      breaks=breaks_vec,name=NULL ) + 
      theme(legend.position="right", legend.key.height = unit(0.75, 'cm'),
        legend.text = element_text(size=10, face="bold"),
        legend.title = element_text(NULL),
        title = element_text(face = "bold", size= 10), 
        plot.title = element_text(hjust = 0.5))#  +
        # theme(plot.margin = unit(c(0, 0, 0, 0), "cm"),
        #      axis.title.y = element_text(face ="bold", size=14)) # + 
      # ylab("HBEGF Expression")
```

# Session Info Details

```{r, echo=FALSE, eval=TRUE}
sessionInfo()
```