---
title: "Analysis of CRC Spatial Transcriptomics data from 10X: Proportion Plots"
author: 'Alberto Valdeolivas Urbelz (PS-Biomics&Path)'
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    code_folding: hide
    self_contained: true
    number_sections: yes
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: spacelab
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
params:
  data_directory : "/projects/site/pred/SpatialOmics/Experiment_CRC_AllSamples/"
  analysis_name : "analysis_alberto/"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(vctrs)
library(patchwork, lib.loc = "/apps/rocs/2020.08/cascadelake/software/R/4.1.2-foss-2020a/lib64/R/library")
library(Seurat)
library(readr)
library(stringr)
library(purrr)
library(cowplot)
library(ggpubr)


data_directory <-  params$data_directory
analysis_name <- params$analysis_name

input_folder <- "IntermediaryFiles/"
input_names_1 <- "SeuratList_Clusters_Res05.rds"
input_names_2 <- "Patho_Annotations/"

files_to_read_1 <- paste0(data_directory, analysis_name,input_folder,input_names_1)
files_to_read_2 <- paste0(data_directory, analysis_name,input_folder,input_names_2)

df_patient_ID_transform <- data.frame(
  patient = c("A120838","A121573","A416371","A551763","A595688","A798015",
              "A938797"), 
  patient_ID = c("S4_Col_Sig","S5_Rec","S3_Col_R","S1_Cec","S2_Col_R",
                 "S7_Rec/Sig","S6_Rec"))

## We consider all the samples but the one with low QC results
samples_to_consider <- c("SN048_A121573_Rep1","SN048_A121573_Rep2",
  "SN048_A416371_Rep1", "SN048_A416371_Rep2", "SN123_A551763_Rep1",
  "SN123_A595688_Rep1", "SN123_A798015_Rep1", "SN123_A938797_Rep1_X",
  "SN124_A595688_Rep2", "SN124_A798015_Rep2",
  "SN124_A938797_Rep2", "SN84_A120838_Rep1", "SN84_A120838_Rep2" )
```

We include pathologist annotations into Seurat objects

```{r, warning=FALSE, message=FALSE}
seurat_objects <- readRDS(files_to_read_1)
seurat_objects <- seurat_objects[samples_to_consider]

seurat_objects_patho <- 
  lapply(seurat_objects, function(x){
    current_sample <- unique(x@meta.data$orig.ident)
    
    patho_anno_current_sample <- 
      read_csv(file = paste0(files_to_read_2, "Pathologist_Annotations_", 
        current_sample, ".csv"))
    
    colnames(patho_anno_current_sample) <- c("Barcode", "Pathologist_Annotations")
    
    x@meta.data <- x@meta.data %>%
      tibble::rownames_to_column("spot_id") %>%
      dplyr::left_join(patho_anno_current_sample, by = c("spot_id" = "Barcode")) %>%
      dplyr::mutate(Pathologist_Annotations = str_replace(Pathologist_Annotations, 
       "tumor&stroma IC med to high", "tumor&stroma_IC med to high")) %>%
      dplyr::mutate(Pathologist_Annotations = str_replace(Pathologist_Annotations, 
       "epitehlium&submucosa", "epithelium&submucosa")) %>% 
      dplyr::mutate(Pathologist_Annotations = str_replace(Pathologist_Annotations, 
       "IC aggregate submucosa", "IC aggregate_submucosa")) %>%
      dplyr::mutate(Pathologist_Annotations = str_replace(Pathologist_Annotations, 
       "IC aggregregate_submucosa", "IC aggregate_submucosa")) %>%
      dplyr::mutate(Pathologist_Annotations = str_replace(Pathologist_Annotations, 
       "IC aggreates_stroma or muscularis", "IC aggregate_stroma or muscularis")) %>%
      dplyr::mutate(Pathologist_Annotations = str_replace(Pathologist_Annotations, 
       "IC aggragate_stroma or muscularis", "IC aggregate_stroma or muscularis")) %>%
      dplyr::mutate(Pathologist_Annotations = str_replace(Pathologist_Annotations, 
       "IC aggreates_stroma or muscularis", "IC aggregate_stroma or muscularis")) %>%
      dplyr::mutate(Pathologist_Annotations = str_replace(Pathologist_Annotations, 
       "IC aggregate_muscularis or stroma", "IC aggregate_stroma or muscularis")) %>%
      dplyr::mutate(Pathologist_Annotations = str_replace(Pathologist_Annotations, 
       "stroma desmoplastic_IC low", "stroma_desmoplastic_IC low")) %>%
      dplyr::mutate(Pathologist_Annotations = str_replace(Pathologist_Annotations, 
       "stroma desmoplastic_IC med to high", "stroma_desmoplastic_IC med to high")) %>%
      dplyr::mutate(Pathologist_Annotations = str_replace(Pathologist_Annotations, 
       "stroma_fibroblastic_IC high", "stroma_fibroblastic_IC_high")) %>%
      dplyr::mutate(Pathologist_Annotations = str_replace(Pathologist_Annotations, 
       "stroma_fibroblastic_IC med", "stroma_fibroblastic_IC_med")) %>%
      dplyr::mutate(Pathologist_Annotations = str_replace(Pathologist_Annotations, 
       "stroma_fibroblastic_IC_med", "stroma_fibroblastic_IC med to high")) %>%
      dplyr::mutate(Pathologist_Annotations = str_replace(Pathologist_Annotations, 
       "stroma_fibroblastic_IC_high", "stroma_fibroblastic_IC med to high")) %>%
      dplyr::mutate(Pathologist_Annotations = str_replace(Pathologist_Annotations, 
       "submucosa", "lamina propria")) %>% 
      dplyr::mutate(Pathologist_Annotations = str_replace(Pathologist_Annotations, 
       "epithelium&submucosa", "epithelium&lam propria")) %>% 
      dplyr::mutate(Pathologist_Annotations = str_replace(Pathologist_Annotations, 
       "IC aggregate_submucosa", "IC aggregate_lam propria")) %>% 
      tibble::column_to_rownames("spot_id") 
      
    
    saveRDS(object = x@meta.data %>%
      tibble::rownames_to_column("spot_id"), 
            file =  paste0(files_to_read_2, "Pathologist_Annotations_", 
        current_sample, ".rds"))
              
    Idents(x) <- 'Pathologist_Annotations'
    
    return(x)
})

```

We visualize the global number of spots with the different annotations

```{r, warning=FALSE, message=FALSE}
all_annotations <- 
  unlist(lapply(seurat_objects_patho, function(x){x@meta.data$Pathologist_Annotations}))
table(all_annotations)

list_all_annotations <- 
  lapply(seurat_objects_patho, function(x){unique(x@meta.data$Pathologist_Annotations)})
  
table(unlist(list_all_annotations))

selected_annotations <- names(which(table(unlist(list_all_annotations)) > 2))
selected_annotations <- selected_annotations[selected_annotations != "exclude"]

seurat_objects_patho_merge <- 
  merge(x= seurat_objects_patho[[1]],y=seurat_objects_patho[-1])
```

Then, we read cell2location results, to explore the cell type content/abundance 
of this tumor annotated spots

```{r, warning=FALSE, message=FALSE}
filename_results_C2L_Korean <- paste0(data_directory, analysis_name, "Cell2Location/results/LocationModelLinearDependentWMultiExperiment_14experiments_36clusters_20654locations_4188genes/W_cell_density_q05.csv")
  
results_C2L_Korean <- read_csv(filename_results_C2L_Korean) %>%
  dplyr::mutate(spot_id = str_remove(.$spot_id, pattern = "Count_")) 
  
colnames(results_C2L_Korean) <-
  str_replace(colnames(results_C2L_Korean), 
              pattern = "q05_spot_factors", replacement = "")

results_C2L_Korean <- results_C2L_Korean %>% 
  dplyr::mutate(sample = str_extract(.$spot_id, pattern =".*Rep[1-2]")) %>%
  dplyr::mutate(spot_id = str_remove(.$spot_id, pattern =".*Rep[1-2]_")) %>%
  dplyr::mutate(spot_id = str_remove(spot_id, pattern = "^X_")) %>%
  tidyr::pivot_longer(!c(spot_id,sample), names_to = "Cell_subtype", values_to = "cell_density_q05") 
```

We then match the spots annotated as tumors with the results of cell2location

```{r, warning=FALSE, message=FALSE}
spots_annotations <- seurat_objects_patho_merge@meta.data %>% 
  dplyr::select(orig.ident, Pathologist_Annotations) %>% 
  tibble::rownames_to_column(var = "spot_id") %>%
  dplyr::mutate(spot_id = str_remove(spot_id, "_[0-9]+")) %>% 
  dplyr::mutate(orig.ident = str_replace(orig.ident, "SN123_A938797_Rep1_X", "SN123_A938797_Rep1")) %>% 
  # dplyr::filter(!is.na(Pathologist_Annotations)) %>% 
  dplyr::filter(Pathologist_Annotations %in%  selected_annotations)

df_C2L_Proportions_annotatedspots <- results_C2L_Korean %>% 
  dplyr::mutate(Cell_subtype = str_replace(Cell_subtype, "CMS[1-4]", "Tumor cells")) %>%
  dplyr::inner_join(spots_annotations, by = c("spot_id" = "spot_id", "sample" = "orig.ident")) %>% 
  dplyr::select(spot_id, sample, cell_density_q05, Cell_subtype, Pathologist_Annotations) %>%
  dplyr::mutate(spot_sample_id = paste0(sample, "_", spot_id)) %>% 
  dplyr::group_by(spot_sample_id) %>% 
  dplyr::mutate(Proportions_perSpot = cell_density_q05/sum(cell_density_q05)) %>%
  dplyr::ungroup()

avg_proportion_annotations <- results_C2L_Korean %>%
  dplyr::mutate(Cell_subtype = str_replace(Cell_subtype, "CMS[1-4]", "Tumor cells")) %>%
  dplyr::inner_join(spots_annotations, by = c("spot_id" = "spot_id", "sample" = "orig.ident")) %>% 
  dplyr::select(spot_id, sample, cell_density_q05, Cell_subtype, Pathologist_Annotations) %>%
  dplyr::mutate(spot_sample_id = paste0(sample, "_", spot_id)) %>% 
  dplyr::group_by(spot_sample_id) %>% 
  dplyr::mutate(Proportions_perSpot = cell_density_q05/sum(cell_density_q05)) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(Pathologist_Annotations, Cell_subtype) %>% 
  dplyr::mutate(avg_proportion_cell_annotations = mean(Proportions_perSpot)) %>%
  dplyr::distinct(Cell_subtype, avg_proportion_cell_annotations)
```

```{r, warning=FALSE, message=FALSE, eval=FALSE}
n_random <- 10000
all_patho_annotations <- unique(spots_annotations$Pathologist_Annotations)
df_final <- data.frame()
  
for (i in seq(n_random)){
    
  df_random <- df_C2L_Proportions_annotatedspots %>%
    dplyr::mutate(rand_proportions = base::sample(Proportions_perSpot)) %>%
    dplyr::group_by(Pathologist_Annotations, Cell_subtype) %>% 
    dplyr::mutate(avg_proportion_random = mean(rand_proportions)) %>%
    dplyr::distinct(Cell_subtype, avg_proportion_random)
  
  df_random_real <- df_random %>% 
    dplyr::inner_join(avg_proportion_annotations, 
      by = c("Cell_subtype", "Pathologist_Annotations")) %>%
    dplyr::mutate(difference_real_random = avg_proportion_cell_annotations - avg_proportion_random) %>%
    dplyr::select(Pathologist_Annotations, Cell_subtype, difference_real_random)
     
  df_final <- rbind(df_final, df_random_real)
}  
  
df_all <-df_final %>% dplyr::group_by(Pathologist_Annotations, Cell_subtype) %>% 
  dplyr::mutate(NES = mean(difference_real_random)/sd(difference_real_random)) %>% 
  dplyr::distinct(Pathologist_Annotations, Cell_subtype, NES)

saveRDS(df_all, 
  paste0(data_directory, analysis_name, input_folder, "df_enrichment_depletion_analysis.rds"))
```

## Overall plot with all the samples and annotations 

```{r, warning=FALSE, message=FALSE, fig.width=12, fig.height=10}
df_all <- 
  readRDS(paste0(data_directory, analysis_name, input_folder, "df_enrichment_depletion_analysis.rds"))

cells_to_non_visualize <- c("Unknown")

final_df_avg_toplot<- df_all %>% 
  dplyr::filter(!(Cell_subtype %in% cells_to_non_visualize)) %>% 
  dplyr::mutate(Direction = ifelse(NES >= 0, "Enrichment", "Depletion"))

final_df_avg_toplot$Pathologist_Annotations <- 
  factor(final_df_avg_toplot$Pathologist_Annotations , 
         levels = rev(c("non neo epithelium", "lamina propria", 
          "epithelium&lamina propria", 
          "muscularis_IC med to high", "tumor", "tumor&stroma_IC med to high", 
          "stroma_fibroblastic_IC med to high", "IC aggregate_lamina propria", 
          "IC aggregate_stroma or muscularis")))

final_df_avg_toplot$Cell_subtype <- 
  factor(final_df_avg_toplot$Cell_subtype , 
         levels = c("Mature Enterocytes type 1", "Mature Enterocytes type 2", 
        "Goblet cells", "Stem-like/TA", "Intermediate", 
        "Smooth muscle cells", "Tumor cells", "Myofibroblasts", 
        "Stromal 1", "Stromal 2", "Stromal 3", "Tip-like ECs", "Stalk-like ECs",
        "Proliferative ECs", "Lymphatic ECs", "Pericytes", "Enteric glial cells",
        "SPP1+", "Pro-inflammatory", "Proliferating", "cDC", "CD19+CD20+ B", 
        "IgA+ Plasma", "IgG+ Plasma", "CD4+ T cells", "CD8+ T cells", "Regulatory T cells", 
        "T follicular helper cells", "T helper 17 cells", "gamma delta T cells", 
        "NK cells", "Mast cells"))

enrichment_depletion_plot <- 
  ggplot(final_df_avg_toplot, aes(x = Pathologist_Annotations, 
    y = Cell_subtype, colour = Direction, size = abs(NES))) + 
    geom_point() + 
    scale_size_continuous(range = c(1.5, 9.5), breaks = c(5,35), labels = c("Small","Large")) + theme_bw() + 
    theme(axis.title = element_blank(), 
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          axis.text.x = element_text(size=12, face = "bold", family="Arial", angle = 90, hjust = 1),
          axis.text.y = element_text(size=12, face = "bold", family="Arial"),
          plot.title =  element_text(face="bold", size=14, hjust = 0.5) , 
          legend.position ="right", legend.text = element_text(size=10, face="bold", family="Arial"),
          legend.title =  element_text(size=12, face="bold", family="Arial")) +
          scale_color_manual(values = c( "#377EB8", "#E41A1C")) + 
          coord_flip() + 
          guides(col = guide_legend(order = 1, override.aes = list(size = 5)),  
          size =guide_legend(title = "Size Effect", order =2, override.aes = list(size = c(3.5,7.5)))) +
          labs(title ="Enrichment \n\ Depletion") 
enrichment_depletion_plot
# ggsave(filename = "/pstore/home/valdeola/Figure1_EnrichmentDepletionPlot.tiff", device='tiff', dpi=300)
```


```{r, warning=FALSE, message=FALSE, fig.width=12, fig.height=10}
enrichment_depletion_plot_V2 <- 
  ggplot(final_df_avg_toplot, aes(x = Pathologist_Annotations, 
    y = Cell_subtype, colour = Direction, size = abs(NES))) + 
    geom_point() + 
    #scale_size_continuous(range = c(1.5, 9.5), breaks = c(5,35), labels = c("Small","Large")) 
    theme_bw() + 
    theme(axis.title = element_blank(), 
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          axis.text.x = element_text(size=12, face = "bold", family="Arial", angle = 90, hjust = 1),
          axis.text.y = element_text(size=12, face = "bold", family="Arial"),
          plot.title =  element_text(face="bold", size=14, hjust = 0.5) , 
          legend.position ="right", legend.text = element_text(size=10, face="bold", family="Arial"),
          legend.title =  element_text(size=12, face="bold", family="Arial")) +
          scale_color_manual(values = c( "#377EB8", "#E41A1C")) + 
          coord_flip() + 
          guides(col = guide_legend(title = "Effect", order = 1, override.aes = list(size = 5)),  
          size =guide_legend(title = "Enrichment Score", order =2)) 
          # labs(title ="Enrichment \n\ Depletion") 
enrichment_depletion_plot_V2
ggsave(filename = "home/valdeola/Desktop/Figure1_EnrichmentDepletionPlot_V2.tiff", device='tiff', dpi=300)
```


```{r, warning=FALSE, message=FALSE}
ggballoonplot(final_df_avg_toplot, fill = "NES") +
  scale_fill_gradient2(low = "#377EB8",mid = "white",high = "#E41A1C") + 
  theme(axis.title = element_blank(), 
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          axis.text.x = element_text(size=12, face = "bold", family="Arial", angle = 90, hjust = 1),
          axis.text.y = element_text(size=12, face = "bold", family="Arial"),
          plot.title =  element_text(face="bold", size=14, hjust = 0.5) , 
          legend.position ="right", legend.text = element_text(size=10, face="bold", family="Arial"),
          legend.title =  element_text(size=12, face="bold", family="Arial")) + 
          
  
  coord_flip()
```


## Plot considering only CMS2 

```{r, warning=FALSE, message=FALSE}
samples_to_consider_CMS2 <- c("SN048_A121573_Rep1", "SN048_A121573_Rep2", 
  "SN123_A595688_Rep1", "SN123_A938797_Rep1", "SN124_A595688_Rep2", 
  "SN124_A938797_Rep2", "SN84_A120838_Rep1", "SN84_A120838_Rep2")

spots_annotations_CMS2 <- seurat_objects_patho_merge@meta.data %>% 
  dplyr::select(orig.ident, Pathologist_Annotations) %>% 
  tibble::rownames_to_column(var = "spot_id") %>%
  dplyr::mutate(spot_id = str_remove(spot_id, "_[0-9]+")) %>% 
  dplyr::mutate(orig.ident = str_replace(orig.ident, "SN123_A938797_Rep1_X", "SN123_A938797_Rep1")) %>% 
  dplyr::filter(!is.na(Pathologist_Annotations)) %>% 
  dplyr::filter(Pathologist_Annotations !=  "exclude") %>% 
  # dplyr::filter(Pathologist_Annotations %in%  selected_annotations) %>% 
  dplyr::filter(orig.ident %in% samples_to_consider_CMS2)

df_C2L_Proportions_annotatedspots_CMS2 <- results_C2L_Korean %>%
  dplyr::inner_join(spots_annotations_CMS2, by = c("spot_id" = "spot_id", "sample" = "orig.ident")) %>% 
  dplyr::select(spot_id, sample, cell_density_q05, Cell_subtype, Pathologist_Annotations) %>%
  dplyr::mutate(spot_sample_id = paste0(sample, "_", spot_id)) %>% 
  dplyr::group_by(spot_sample_id) %>% 
  dplyr::mutate(Proportions_perSpot = cell_density_q05/sum(cell_density_q05)) %>%
  dplyr::ungroup()

avg_proportion_annotations_CMS2 <- results_C2L_Korean %>%
  dplyr::inner_join(spots_annotations_CMS2, by = c("spot_id" = "spot_id", "sample" = "orig.ident")) %>% 
  dplyr::select(spot_id, sample, cell_density_q05, Cell_subtype, Pathologist_Annotations) %>%
  dplyr::mutate(spot_sample_id = paste0(sample, "_", spot_id)) %>% 
  dplyr::group_by(spot_sample_id) %>% 
  dplyr::mutate(Proportions_perSpot = cell_density_q05/sum(cell_density_q05)) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(Pathologist_Annotations, Cell_subtype) %>% 
  dplyr::mutate(avg_proportion_cell_annotations = mean(Proportions_perSpot)) %>%
  dplyr::distinct(Cell_subtype, avg_proportion_cell_annotations)
```

```{r, warning=FALSE, message=FALSE, eval=FALSE}
n_random <- 10000
df_final_CMS2 <- data.frame()
  
for (i in seq(n_random)){
    
  df_random_CMS2 <- df_C2L_Proportions_annotatedspots_CMS2 %>%
    dplyr::mutate(rand_proportions = base::sample(Proportions_perSpot)) %>%
    dplyr::group_by(Pathologist_Annotations, Cell_subtype) %>% 
    dplyr::mutate(avg_proportion_random = mean(rand_proportions)) %>%
    dplyr::distinct(Cell_subtype, avg_proportion_random)
  
  df_random_real_CMS2 <- df_random_CMS2 %>% 
    dplyr::inner_join(avg_proportion_annotations_CMS2) %>%
    dplyr::mutate(difference_real_random = avg_proportion_cell_annotations - avg_proportion_random) %>%
    dplyr::select(Pathologist_Annotations, Cell_subtype, difference_real_random)
     
  df_final_CMS2 <- rbind(df_final_CMS2, df_random_real_CMS2)
}  
  
df_all_CMS2 <-df_final_CMS2 %>% 
  dplyr::group_by(Pathologist_Annotations, Cell_subtype) %>% 
  dplyr::mutate(NES = mean(difference_real_random)/sd(difference_real_random)) %>% 
  dplyr::distinct(Pathologist_Annotations, Cell_subtype, NES)

saveRDS(df_all_CMS2, 
  paste0(data_directory, analysis_name, input_folder, "df_enrichment_depletion_analysis_CMS2.rds"))
```

```{r, warning=FALSE, message=FALSE, fig.width=12, fig.height=10}
df_all_CMS2 <- 
  readRDS(paste0(data_directory, analysis_name, input_folder, "df_enrichment_depletion_analysis_CMS2.rds"))

cells_to_non_visualize <- c("Unknown", "Mature Enterocytes type 1", 
  "Mature Enterocytes type 2", "Goblet cells", "Stem-like/TA", "Intermediate", 
  "Smooth muscle cells", "Tip-like ECs", "cDC", "gamma delta T cells", 
  "Lymphatic ECs", "Pericytes", "Pro-inflammatory", "Proliferating", "Proliferative ECs", 
  "T helper 17 cells", "NK cells", "Enteric glial cells","Stalk-like ECs" )

final_df_avg_toplot_CMS2<- df_all_CMS2 %>% 
  dplyr::filter(!(Cell_subtype %in% cells_to_non_visualize)) %>% 
  dplyr::mutate(Direction = ifelse(NES >= 0, "Enrichment", "Depletion"))

final_df_avg_toplot_CMS2$Pathologist_Annotations <- 
  factor(final_df_avg_toplot_CMS2$Pathologist_Annotations , 
         levels = rev(c("lamina propria", "non neo epithelium", 
                    "epithelium&lamina propria", 
                    "muscularis_IC med to high", "tumor", 
                    "tumor&stroma_IC low", 
                    "tumor&stroma_IC med to high", 
                    "stroma_fibroblastic_IC low", 
                    "stroma_fibroblastic_IC med to high", 
                    "stroma_desmoplastic_IC low", 
                    "stroma_desmoplastic_IC med to high", 
                    "IC aggregate_lamina propria", 
                    "IC aggregate_stroma or muscularis")))

final_df_avg_toplot_CMS2$Cell_subtype <- 
  factor(final_df_avg_toplot_CMS2$Cell_subtype , 
         levels = c("CMS1", "CMS2", "CMS3", "CMS4", 
          "Myofibroblasts", "Stromal 1", "Stromal 2", "Stromal 3", 
          "SPP1+", 
          "CD19+CD20+ B", "IgA+ Plasma", "IgG+ Plasma", 
          "CD4+ T cells", "CD8+ T cells", "Regulatory T cells", 
          "T follicular helper cells", "Mast cells")) 
                    
        
enrichment_depletion_plot_CMS2 <- 
  ggplot(final_df_avg_toplot_CMS2, aes(x = Pathologist_Annotations, 
    y = Cell_subtype, colour = Direction, size = abs(NES))) + 
    geom_point() + 
    # scale_size_continuous(range = c(1.5, 9.5), breaks = c(5,35), labels = c("Small","Large")) + 
    theme_bw() + 
    theme(axis.title = element_blank(), 
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          axis.text.x = element_text(size=12, face = "bold", family="Arial", angle = 90, hjust = 1),
          axis.text.y = element_text(size=12, face = "bold", family="Arial"),
          plot.title =  element_text(face="bold", size=14, hjust = 0.5) , 
          legend.position ="right", legend.text = element_text(size=10, face="bold", family="Arial"),
          legend.title =  element_text(size=12, face="bold", family="Arial")) +
          scale_color_manual(values = c( "#377EB8", "#E41A1C")) + 
          coord_flip() +  
          guides(col = guide_legend(title = "Effect", order = 1, override.aes = list(size = 5)),  
          size =guide_legend(title = "Enrichment Score", order =2)) 

enrichment_depletion_plot_CMS2
```

## Plot considering only mixed phenotype samples 


```{r, warning=FALSE, message=FALSE}
samples_to_consider_CMS2_CMS1 <- c("SN048_A416371_Rep1", "SN048_A416371_Rep2", 
  "SN123_A551763_Rep1")

spots_annotations_CMS2_CMS1 <- seurat_objects_patho_merge@meta.data %>% 
  dplyr::select(orig.ident, Pathologist_Annotations) %>% 
  tibble::rownames_to_column(var = "spot_id") %>%
  dplyr::mutate(spot_id = str_remove(spot_id, "_[0-9]+")) %>% 
  dplyr::mutate(orig.ident = str_replace(orig.ident, "SN123_A938797_Rep1_X", "SN123_A938797_Rep1")) %>% 
  dplyr::filter(!is.na(Pathologist_Annotations)) %>% 
  dplyr::filter(Pathologist_Annotations !=  "exclude") %>% 
  dplyr::filter(orig.ident %in% samples_to_consider_CMS2_CMS1)

df_C2L_Proportions_annotatedspots_CMS2_CMS1 <- results_C2L_Korean %>%
  dplyr::inner_join(spots_annotations_CMS2_CMS1, by = c("spot_id" = "spot_id", "sample" = "orig.ident")) %>% 
  dplyr::select(spot_id, sample, cell_density_q05, Cell_subtype, Pathologist_Annotations) %>%
  dplyr::mutate(spot_sample_id = paste0(sample, "_", spot_id)) %>% 
  dplyr::group_by(spot_sample_id) %>% 
  dplyr::mutate(Proportions_perSpot = cell_density_q05/sum(cell_density_q05)) %>%
  dplyr::ungroup()

avg_proportion_annotations_CMS2_CMS1 <- results_C2L_Korean %>%
  dplyr::inner_join(spots_annotations_CMS2_CMS1, by = c("spot_id" = "spot_id", "sample" = "orig.ident")) %>% 
  dplyr::select(spot_id, sample, cell_density_q05, Cell_subtype, Pathologist_Annotations) %>%
  dplyr::mutate(spot_sample_id = paste0(sample, "_", spot_id)) %>% 
  dplyr::group_by(spot_sample_id) %>% 
  dplyr::mutate(Proportions_perSpot = cell_density_q05/sum(cell_density_q05)) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(Pathologist_Annotations, Cell_subtype) %>% 
  dplyr::mutate(avg_proportion_cell_annotations = mean(Proportions_perSpot)) %>%
  dplyr::distinct(Cell_subtype, avg_proportion_cell_annotations)
```

```{r, warning=FALSE, message=FALSE, eval=FALSE}
n_random <- 10000
df_final_CMS2_CMS1 <- data.frame()
  
for (i in seq(n_random)){
    
  df_random_CMS2_CMS1 <- df_C2L_Proportions_annotatedspots_CMS2_CMS1 %>%
    dplyr::mutate(rand_proportions = base::sample(Proportions_perSpot)) %>%
    dplyr::group_by(Pathologist_Annotations, Cell_subtype) %>% 
    dplyr::mutate(avg_proportion_random = mean(rand_proportions)) %>%
    dplyr::distinct(Cell_subtype, avg_proportion_random)
  
  df_random_real_CMS2_CMS1 <- df_random_CMS2_CMS1 %>% 
    dplyr::inner_join(avg_proportion_annotations_CMS2_CMS1) %>%
    dplyr::mutate(difference_real_random = avg_proportion_cell_annotations - avg_proportion_random) %>%
    dplyr::select(Pathologist_Annotations, Cell_subtype, difference_real_random)
     
  df_final_CMS2_CMS1 <- rbind(df_final_CMS2_CMS1, df_random_real_CMS2_CMS1)
}  
  
df_all_CMS2_CMS1 <-df_final_CMS2_CMS1 %>% 
  dplyr::group_by(Pathologist_Annotations, Cell_subtype) %>% 
  dplyr::mutate(NES = mean(difference_real_random)/sd(difference_real_random)) %>% 
  dplyr::distinct(Pathologist_Annotations, Cell_subtype, NES)

saveRDS(df_all_CMS2_CMS1, 
  paste0(data_directory, analysis_name, input_folder, "df_enrichment_depletion_analysis_CMS2_CMS1.rds"))
```

```{r, warning=FALSE, message=FALSE, fig.width=12, fig.height=10}
df_all_CMS2_CMS1 <- 
  readRDS(paste0(data_directory, analysis_name, input_folder, "df_enrichment_depletion_analysis_CMS2_CMS1.rds"))

cells_to_non_visualize <- c("Unknown", "Mature Enterocytes type 1", 
  "Mature Enterocytes type 2", "Goblet cells", "Stem-like/TA", "Intermediate", 
  "Smooth muscle cells", "Tip-like ECs", "cDC", "gamma delta T cells", 
  "Lymphatic ECs", "Pericytes", "Pro-inflammatory", "Proliferating", "Proliferative ECs", 
  "T helper 17 cells", "NK cells", "Enteric glial cells","Stalk-like ECs")

final_df_avg_toplot_CMS2_CMS1<- df_all_CMS2_CMS1 %>% 
  dplyr::filter(!(Cell_subtype %in% cells_to_non_visualize)) %>% 
  dplyr::mutate(Direction = ifelse(NES >= 0, "Enrichment", "Depletion"))

final_df_avg_toplot_CMS2_CMS1$Pathologist_Annotations <- 
  factor(final_df_avg_toplot_CMS2_CMS1$Pathologist_Annotations , 
         levels = rev(c("tumor", "tumor&stroma", 
                    "tumor&stroma_IC med to high", 
                    "stroma_fibroblastic_IC med to high")))

final_df_avg_toplot_CMS2_CMS1$Cell_subtype <- 
  factor(final_df_avg_toplot_CMS2_CMS1$Cell_subtype , 
         levels = c("CMS1", "CMS2", "CMS3", "CMS4", 
          "Myofibroblasts", "Stromal 1", "Stromal 2", "Stromal 3", 
          "SPP1+", 
          "CD19+CD20+ B", "IgA+ Plasma", "IgG+ Plasma", 
          "CD4+ T cells", "CD8+ T cells", "Regulatory T cells", 
          "T follicular helper cells", "Mast cells")) 

enrichment_depletion_plot_CMS2_CMS1 <- 
  ggplot(final_df_avg_toplot_CMS2_CMS1, aes(x = Pathologist_Annotations, 
    y = Cell_subtype, colour = Direction, size = abs(NES))) + 
    geom_point() + 
    # scale_size_continuous(range = c(1.5, 9.5), breaks = c(5,35), labels = c("Small","Large")) + 
    theme_bw() + 
    theme(axis.title = element_blank(), 
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          axis.text.x = element_text(size=12, face = "bold", family="Arial", angle = 90, hjust = 1),
          axis.text.y = element_text(size=12, face = "bold", family="Arial"),
          plot.title =  element_text(face="bold", size=14, hjust = 0.5) , 
          legend.position ="right", legend.text = element_text(size=10, face="bold", family="Arial"),
          legend.title =  element_text(size=12, face="bold", family="Arial")) +
          scale_color_manual(values = c( "#377EB8", "#E41A1C")) + 
          coord_flip() +  
          guides(col = guide_legend(title = "Effect", order = 1, override.aes = list(size = 5)),  
          size =guide_legend(title = "Enrichment Score", order =2)) 

enrichment_depletion_plot_CMS2_CMS1
```

```{r, warning=FALSE, message=FALSE, fig.width=12, fig.height=10}
p1 <- enrichment_depletion_plot_CMS2 + theme(legend.position = 'bottom',  
  axis.text.x = element_blank(), axis.ticks.x = element_blank(),
  legend.title =  element_text(size=16, face="bold", family="Arial"), 
  legend.text = element_text(size=14, face="bold", family="Arial")) + 
  ggtitle("Patients with CMS2 phenotype")
p2 <- enrichment_depletion_plot_CMS2_CMS1 + theme(legend.position='bottom', 
  legend.title =  element_text(size=16, face="bold", family="Arial"), 
  legend.text = element_text(size=14, face="bold", family="Arial")) + 
  ggtitle("Patients with CMS1-CMS2 mixed phenotype" )


ggarrange(p1, p2, nrow=2, common.legend = TRUE, legend="bottom", align = 'v') 
ggsave(filename = "home/valdeola/Desktop/Figure2_EnrichmentDepletionPlot_V2.tiff", device='tiff', dpi=300)
```



# Session Info Details

```{r, echo=FALSE, eval=TRUE}
sessionInfo()
```