---
title: "Analysis of CRC Spatial Transcriptomics data from 10X:Correlation Analysis of Pathway Activities and Cell Type abundances for CMS"
author: 'Alberto Valdeolivas Urbelz (PS-Biomics&Path)'
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    code_folding: hide
    self_contained: true
    number_sections: yes
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: spacelab
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
params:
  data_directory : "/pstore/data/biomics/_pre_portfolio/_platform_evaluation/SpatialTranscriptomics/Experiment_CRC_AllSamples/"
  analysis_name : "analysis_alberto/"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

The present script takes the the Spatial Omics data generated with 10X 
technology on a set of CRC samples. In particular, it explores the correlation 
between pathway activities and CMS cell type proportion per spot.  


```{r, warning=FALSE, message=FALSE}
# library(dplyr, lib.loc = "/pstore/apps/bioinfo/R/4.0.1-foss")
library(CCA)
library(ggplot2)
library(patchwork, lib.loc = "/apps/rocs/2020.08/cascadelake/software/R/4.1.2-foss-2020a/lib64/R/library")
library(Seurat)
library(readr)
library(stringr)
library(tibble)
library(ggrepel)
library(RColorBrewer)
library(pheatmap)
source(file = "WrapperFunction/SeuratWrappers.R")

data_directory <-  params$data_directory
analysis_name <- params$analysis_name

input_files <- "IntermediaryFiles/"
input_names <- "SeuratList_Clusters_Res05_Progeny.rds"

files_to_read <- paste0(data_directory, analysis_name,input_files,input_names)
results_directory <- paste0(data_directory, analysis_name, input_files)
# setwd(paste0(data_directory, analysis_name))
```

We read the Seurat objects containing the pathway activity. 

```{r, warning=FALSE, message=FALSE}
seurat_objects <- readRDS(files_to_read)
```

We read the results of the deconvolution process containing cell type abundances
per spot. 

```{r, warning=FALSE, message=FALSE}
filename <- 
  paste0(data_directory,analysis_name,  "Cell2Location/results/LocationModelLinearDependentWMultiExperiment_14experiments_36clusters_20654locations_4188genes/W_mRNA_count_q05.csv")

results_C2L <- read_csv(filename) %>%
  dplyr::mutate(spot_id = str_remove(.$spot_id, pattern = "Count_")) %>%
  dplyr::mutate(spot_id = str_replace(.$spot_id, pattern = "SN123_A938797_Rep1_" ,
  replacement = "SN123_A938797_Rep1_X_")) 


sum_UMI_spot <- rowSums(results_C2L[,-1])
results_C2L[,-1] <- results_C2L[,-1]/sum_UMI_spot

colnames(results_C2L) <- str_replace(colnames(results_C2L), 
    pattern = "q05_nUMI_factors", replacement = "")

## We are only going to consider the CMS signatures

results_C2L <- 
  results_C2L[,colnames(results_C2L) %in% 
                c("spot_id","CMS1", "CMS2", "CMS3", "CMS4")]

### We remove the low quality sample. 

seurat_objects <- 
  seurat_objects[names(seurat_objects) %in% 
                   c("SN124_A551763_Rep2") == FALSE]

```


# Results taking into account correlation between the spots of all the samples together

We display the different correlation plots when this is globally computed per 
all the samples. 

```{r, results = 'asis', message=FALSE, dpi=300, warning=FALSE}
## We merge the datasets and do so formatting. 
merge_CL2_pathways <- merge_assay(seurat_objects) %>%
  dplyr::inner_join(results_C2L, by ="spot_id") %>% 
  dplyr::arrange(spot_id) 


## We separate the merged dataset in our two matrices on which we will compute 
## the canonical correlation.
  
X <- merge_CL2_pathways[,2:15]
Y <- merge_CL2_pathways[,16:19]

correl <- matcor(X, Y)
canonical_correlation2 <- cc(X, Y)
  
cat( "\n\n")
  
cat( "## Pathway activities Correlation: ", "\n\n")
  
MatrixCorrX <-   correl$Xcor %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column(var ="pathway") %>%
  tidyr::pivot_longer(!pathway, values_drop_na = TRUE) # %>% 
  # dplyr::mutate(value = round(value,2))


p_cor_pathways <- ggplot(data = MatrixCorrX, aes(pathway  , name     , fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Pearson\nCorrelation") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 9, hjust = 1))+
  xlab("Pathways") + ylab("Pathways") + 
  coord_fixed()

print(p_cor_pathways)
  
cat( "\n\n")
  
cat( "## Cell type abundances Correlation: ", "\n\n")
  
MatrixCorrY <-   correl$Ycor %>% 
    as.data.frame() %>% 
    tibble::rownames_to_column(var ="CellType") %>%
    tidyr::pivot_longer(!CellType, values_drop_na = TRUE) 
  

p_cor_celltype <- ggplot(MatrixCorrY, aes(name , CellType , fill = value))+
    geom_tile(color = "white")+
    scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Pearson\nCorrelation") +
    theme_minimal()+ 
    theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 10, hjust = 1))+
    xlab("Cell Type") + ylab("Cell Type") +
    coord_fixed()

print(p_cor_celltype)
  
cat( "\n\n")
  
cat( "## Cross Pathway-Cell type abundances Correlation: ", "\n\n")
  
CrossCorrelation <- correl$XYcor[1:14,15:18]


MatrixCorrXY <-   CrossCorrelation %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column(var ="var1") %>%
  tidyr::pivot_longer(!var1, values_drop_na = TRUE) %>% 
  dplyr::mutate(value = round(value,2))


p_CrossCorrelation <- ggplot(data = MatrixCorrXY, aes(var1,name  , fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Pearson\nCorrelation") +
    theme_minimal()+ 
    theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 10, hjust = 1))+
    xlab("Pathways") + ylab("Cell Type") + 
    coord_fixed()

print(p_CrossCorrelation)

mypalette <- rev(brewer.pal(n = 9, name ="RdBu"))
  
par(mar=c(10,10,10,10))
    
pheatmap(t(CrossCorrelation), treeheight_row=0,treeheight_col=0, 
  color =colorRampPalette(mypalette)(100),cluster_rows = FALSE,
  fontsize = 16, cluster_cols = TRUE, breaks=seq(-1,1,length.out = 
  length(colorRampPalette(mypalette)(100)) + 1))
  
cat( "\n\n")
  
cat( "## Variables representation on the canonical variates: ", "\n\n")
  
  
d1 = 2 
d2 = 1
df2 = data.frame(comp1 = c(canonical_correlation2$scores$corr.X.xscores[, d1],
                           canonical_correlation2$scores$corr.Y.xscores[, d1]),
                 comp2 = c(canonical_correlation2$scores$corr.X.xscores[, d2],
                           canonical_correlation2$scores$corr.Y.xscores[, d2]),
                color = c(rep("Pathways", ncol(X)), 
                        (rep("Cell Types", ncol(Y)))))
  
p_Canonival_variates <-ggplot(df2, aes(comp1, comp2), colour = color) +
  geom_label_repel(aes(colour = color, label= rownames(df2)),
          vjust=0,size = 3, family = "sans", direction= "both") +
  theme_bw() + 
  theme(legend.position="none") +
  theme(legend.title=element_blank())
  
print(p_Canonival_variates)
  
cat( "\n\n") 
```


# Conclusion 


# Session Info Details

```{r, echo=FALSE, eval=TRUE}
sessionInfo()
```