---
title: "Analysis of CRC Spatial Transcriptomics data from 10X: Normalization and Clustering"
author: 'Alberto Valdeolivas Urbelz (PS-Biomics&Path)'
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    code_folding: hide
    self_contained: true
    number_sections: yes
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: spacelab
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
params:
  data_directory : "/projects/site/pred/SpatialOmics/Experiment_CRC_AllSamples/"
  analysis_name : "analysis_alberto/"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

The present script evaluates the correlation between the deconvolution results
in small clusters that can be considered equivalent between replicates. 

```{r, warning=FALSE,message=FALSE}
library(ggplot2)
library(vctrs)
library(dplyr)
library(patchwork, lib.loc = "/apps/rocs/2020.08/cascadelake/software/R/4.1.2-foss-2020a/lib64/R/library")
library(Seurat)
library(readr)
library(stringr)
library(tidyr)
library(ggpubr)
library(purrr)
library(harmony)


# library(EnhancedVolcano)
source(file = "WrapperFunction/SeuratWrappers.R")
source(file = "WrapperFunction/ComputeCorrelationReplicates.R")


data_directory <-  params$data_directory
analysis_name <- params$analysis_name

input_files <- "IntermediaryFiles/"
input_names <- "SeuratList_Clusters_Res05.rds"

files_to_read <- paste0(data_directory, analysis_name,input_files,input_names)

df_patient_ID_transform <- data.frame(
  patient = c("A120838","A121573","A416371","A551763","A595688","A798015",
              "A938797"), 
  patient_ID = c("S4_Col_Sig","S5_Rec","S3_Col_R","S1_Cec","S2_Col_R",
                 "S7_Rec/Sig","S6_Rec"))
```


We read the Cell2location results

```{r, message=FALSE, warning=FALSE}
filename_results_C2L <- 
  "Cell2Location/results/LocationModelLinearDependentWMultiExperiment_14experiments_36clusters_20654locations_4188genes/W_mRNA_count_q05.csv"

results_C2L <- read_csv(filename_results_C2L) %>%
  dplyr::mutate(spot_id = str_remove(.$spot_id, pattern = "Count_")) %>%
  dplyr::mutate(spot_id = str_replace(.$spot_id, pattern = "SN123_A938797_Rep1_" ,
  replacement = "SN123_A938797_Rep1_X_")) 

colnames(results_C2L) <- str_replace(colnames(results_C2L), 
    pattern = "q05_nUMI_factors", replacement = "")

cell_types <- colnames(results_C2L[,-1])
```

```{r, message=FALSE, warning=FALSE}
## Generating Seurat Objects 
seurat_objects <- readRDS(files_to_read)
```

# Clustering together technical replicates

We cluster together the technical replicates to have equivalent regions where
we can evaluate the proportions of different cell types between different 
samples

## SN048_A121573_Rep1 & SN048_A121573_Rep2

```{r, message=FALSE, warning=FALSE}
rep1_name <- "SN048_A121573_Rep1"
rep2_name <- "SN048_A121573_Rep2"

rep1_seurat_obj <- seurat_objects[[rep1_name]]
rep2_seurat_obj <- seurat_objects[[rep2_name]]

merge_seurat_object <- 
  merge(rep1_seurat_obj,rep2_seurat_obj)

rep1_rep2_merge <-  
  CreateSeuratObject(counts = GetAssayData(merge_seurat_object, slot = "counts"), 
    min.cells = 5) %>%
    SCTransform(verbose = FALSE) %>%
    RunPCA(pc.genes = .@var.genes, npcs = 30, verbose = FALSE) %>% 
    RunUMAP(dims = 1:30)

rep1_rep2_merge@images <- merge_seurat_object@images
rep1_rep2_merge@meta.data$orig.ident <- 
  merge_seurat_object@meta.data$orig.ident  
```

```{r, warning=FALSE, message=FALSE}
rep1_rep2_merge <- rep1_rep2_merge %>% 
    RunHarmony(c("orig.ident"), plot_convergence = TRUE, assay.use = "SCT",
               max.iter.harmony = 20)
```

Correlation for different resolution parameters

```{r, warning=FALSE, message=FALSE}
resolution_pars <- seq(from = 1, to = 2, by = 0.1)

correlations <- resolution_pars %>% 
  map_dbl(function(x){
    rep1_rep2_merge <- ComputeClustering(rep1_rep2_merge, x)
    ComputeCorrelation(rep1_rep2_merge, results_C2L)
    
  } )
df_A121573 <- data.frame(resolution_pars = numeric(length(resolution_pars)))
df_A121573$resolution_pars <- resolution_pars
df_A121573$correlations <- correlations
df_A121573$patient_ID <- "A121573"
```

## SN048_A416371_Rep1 & SN048_A416371_Rep2

```{r, message=FALSE, warning=FALSE}
rep1_name <- "SN048_A416371_Rep1"
rep2_name <- "SN048_A416371_Rep2"

rep1_seurat_obj <- seurat_objects[[rep1_name]]
rep2_seurat_obj <- seurat_objects[[rep2_name]]

merge_seurat_object <- 
  merge(rep1_seurat_obj,rep2_seurat_obj)

rep1_rep2_merge <-  
  CreateSeuratObject(counts = GetAssayData(merge_seurat_object, slot = "counts"), 
    min.cells = 5) %>%
    SCTransform(verbose = FALSE) %>%
    RunPCA(pc.genes = .@var.genes, npcs = 30, verbose = FALSE) %>% 
    RunUMAP(dims = 1:30)

rep1_rep2_merge@images <- merge_seurat_object@images
rep1_rep2_merge@meta.data$orig.ident <- 
  merge_seurat_object@meta.data$orig.ident  
```

```{r, warning=FALSE, message=FALSE}
rep1_rep2_merge <- rep1_rep2_merge %>% 
    RunHarmony(c("orig.ident"), plot_convergence = TRUE, assay.use = "SCT",
               max.iter.harmony = 20)
```

Correlation for different resolution parameters

```{r, warning=FALSE, message=FALSE}
resolution_pars <- seq(from = 1, to = 2, by = 0.1)

correlations <- resolution_pars %>% 
  map_dbl(function(x){
    rep1_rep2_merge <- ComputeClustering(rep1_rep2_merge, x)
    ComputeCorrelation(rep1_rep2_merge, results_C2L)
    
  } )
df_A416371 <- data.frame(resolution_pars = numeric(length(resolution_pars)))
df_A416371$resolution_pars <- resolution_pars
df_A416371$correlations <- correlations
df_A416371$patient_ID <- "A416371"
```

## SN123_A551763_Rep1 & SN124_A551763_Rep2

To consider these results with care. The SN124_A551763_Rep2 sample does not 
pass the QC checks. 

```{r, message=FALSE, warning=FALSE}
rep1_name <- "SN123_A551763_Rep1"
rep2_name <- "SN124_A551763_Rep2"

rep1_seurat_obj <- seurat_objects[[rep1_name]]
rep2_seurat_obj <- seurat_objects[[rep2_name]]

merge_seurat_object <- 
  merge(rep1_seurat_obj,rep2_seurat_obj)

rep1_rep2_merge <-  
  CreateSeuratObject(counts = GetAssayData(merge_seurat_object, slot = "counts"), 
    min.cells = 5) %>%
    SCTransform(verbose = FALSE) %>%
    RunPCA(pc.genes = .@var.genes, npcs = 30, verbose = FALSE) %>% 
    RunUMAP(dims = 1:30)

rep1_rep2_merge@images <- merge_seurat_object@images
rep1_rep2_merge@meta.data$orig.ident <- 
  merge_seurat_object@meta.data$orig.ident  
```


```{r, warning=FALSE, message=FALSE}
rep1_rep2_merge <- rep1_rep2_merge %>% 
    RunHarmony(c("orig.ident"), plot_convergence = TRUE, assay.use = "SCT",
               max.iter.harmony = 20)
```

Correlation for different resolution parameters

```{r, warning=FALSE, message=FALSE}
resolution_pars <- seq(from = 1, to = 2, by = 0.1)

correlations <- resolution_pars %>% 
  map_dbl(function(x){
    rep1_rep2_merge <- ComputeClustering(rep1_rep2_merge, x)
    ComputeCorrelation(rep1_rep2_merge, results_C2L)
    
  } )
df_A551763 <- data.frame(resolution_pars = numeric(length(resolution_pars)))
df_A551763$resolution_pars <- resolution_pars
df_A551763$correlations <- correlations
df_A551763$patient_ID <- "A551763"
```

## SN123_A595688_Rep1 & SN124_A595688_Rep2
 

```{r, message=FALSE, warning=FALSE}
rep1_name <- "SN123_A595688_Rep1"
rep2_name <- "SN124_A595688_Rep2"

rep1_seurat_obj <- seurat_objects[[rep1_name]]
rep2_seurat_obj <- seurat_objects[[rep2_name]]

merge_seurat_object <- 
  merge(rep1_seurat_obj,rep2_seurat_obj)

rep1_rep2_merge <-  
  CreateSeuratObject(counts = GetAssayData(merge_seurat_object, slot = "counts"), 
    min.cells = 5) %>%
    SCTransform(verbose = FALSE) %>%
    RunPCA(pc.genes = .@var.genes, npcs = 30, verbose = FALSE) %>% 
    RunUMAP(dims = 1:30)

rep1_rep2_merge@images <- merge_seurat_object@images
rep1_rep2_merge@meta.data$orig.ident <- 
  merge_seurat_object@meta.data$orig.ident  
```

```{r, warning=FALSE, message=FALSE}
rep1_rep2_merge <- rep1_rep2_merge %>% 
    RunHarmony(c("orig.ident"), plot_convergence = TRUE, assay.use = "SCT",
               max.iter.harmony = 20)
```

Correlation for different resolution parameters

```{r, warning=FALSE, message=FALSE}
resolution_pars <- seq(from = 1, to = 2, by = 0.1)

correlations <- resolution_pars %>% 
  map_dbl(function(x){
    rep1_rep2_merge <- ComputeClustering(rep1_rep2_merge, x)
    ComputeCorrelation(rep1_rep2_merge, results_C2L)
    
  } )
df_A595688 <- data.frame(resolution_pars = numeric(length(resolution_pars)))
df_A595688$resolution_pars <- resolution_pars
df_A595688$correlations <- correlations
df_A595688$patient_ID <- "A595688"
```

## SN123_A798015_Rep1 & SN124_A798015_Rep2
 

```{r, message=FALSE, warning=FALSE}
rep1_name <- "SN123_A798015_Rep1"
rep2_name <- "SN124_A798015_Rep2"

rep1_seurat_obj <- seurat_objects[[rep1_name]]
rep2_seurat_obj <- seurat_objects[[rep2_name]]

merge_seurat_object <- 
  merge(rep1_seurat_obj,rep2_seurat_obj)

rep1_rep2_merge <-  
  CreateSeuratObject(counts = GetAssayData(merge_seurat_object, slot = "counts"), 
    min.cells = 5) %>%
    SCTransform(verbose = FALSE) %>%
    RunPCA(pc.genes = .@var.genes, npcs = 30, verbose = FALSE) %>% 
    RunUMAP(dims = 1:30)

rep1_rep2_merge@images <- merge_seurat_object@images
rep1_rep2_merge@meta.data$orig.ident <- 
  merge_seurat_object@meta.data$orig.ident  
```

```{r, warning=FALSE, message=FALSE}
rep1_rep2_merge <- rep1_rep2_merge %>% 
    RunHarmony(c("orig.ident"), plot_convergence = TRUE, assay.use = "SCT",
               max.iter.harmony = 20)
```

Correlation for different resolution parameters

```{r, warning=FALSE, message=FALSE}
resolution_pars <- seq(from = 1, to = 2, by = 0.1)

correlations <- resolution_pars %>% 
  map_dbl(function(x){
    rep1_rep2_merge <- ComputeClustering(rep1_rep2_merge, x)
    ComputeCorrelation(rep1_rep2_merge, results_C2L)
    
  } )
df_A798015 <- data.frame(resolution_pars = numeric(length(resolution_pars)))
df_A798015$resolution_pars <- resolution_pars
df_A798015$correlations <- correlations
df_A798015$patient_ID <- "A798015"
```

## SN123_A938797_Rep1_X & SN124_A938797_Rep2
 

```{r, message=FALSE, warning=FALSE}
rep1_name <- "SN123_A938797_Rep1_X"
rep2_name <- "SN124_A938797_Rep2"

rep1_seurat_obj <- seurat_objects[[rep1_name]]
rep2_seurat_obj <- seurat_objects[[rep2_name]]

merge_seurat_object <- 
  merge(rep1_seurat_obj,rep2_seurat_obj)

rep1_rep2_merge <-  
  CreateSeuratObject(counts = GetAssayData(merge_seurat_object, slot = "counts"), 
    min.cells = 5) %>%
    SCTransform(verbose = FALSE) %>%
    RunPCA(pc.genes = .@var.genes, npcs = 30, verbose = FALSE) %>% 
    RunUMAP(dims = 1:30)

rep1_rep2_merge@images <- merge_seurat_object@images
rep1_rep2_merge@meta.data$orig.ident <- 
  merge_seurat_object@meta.data$orig.ident  
```

```{r, warning=FALSE, message=FALSE}
rep1_rep2_merge <- rep1_rep2_merge %>% 
    RunHarmony(c("orig.ident"), plot_convergence = TRUE, assay.use = "SCT",
               max.iter.harmony = 20)
```

Correlation for different resolution parameters

```{r, warning=FALSE, message=FALSE}
resolution_pars <- seq(from = 1, to = 2, by = 0.1)

correlations <- resolution_pars %>% 
  map_dbl(function(x){
    rep1_rep2_merge <- ComputeClustering(rep1_rep2_merge, x)
    ComputeCorrelation(rep1_rep2_merge, results_C2L)
    
  } )
df_A938797 <- data.frame(resolution_pars = numeric(length(resolution_pars)))
df_A938797$resolution_pars <- resolution_pars
df_A938797$correlations <- correlations
df_A938797$patient_ID <- "A938797"
```

## SN84_A120838_Rep1 & SN84_A120838_Rep2
 

```{r, message=FALSE, warning=FALSE}
rep1_name <- "SN84_A120838_Rep1"
rep2_name <- "SN84_A120838_Rep2"

rep1_seurat_obj <- seurat_objects[[rep1_name]]
rep2_seurat_obj <- seurat_objects[[rep2_name]]

merge_seurat_object <- 
  merge(rep1_seurat_obj,rep2_seurat_obj)

rep1_rep2_merge <-  
  CreateSeuratObject(counts = GetAssayData(merge_seurat_object, slot = "counts"), 
    min.cells = 5) %>%
    SCTransform(verbose = FALSE) %>%
    RunPCA(pc.genes = .@var.genes, npcs = 30, verbose = FALSE) %>% 
    RunUMAP(dims = 1:30)

rep1_rep2_merge@images <- merge_seurat_object@images
rep1_rep2_merge@meta.data$orig.ident <- 
  merge_seurat_object@meta.data$orig.ident  
```

```{r, warning=FALSE, message=FALSE}
rep1_rep2_merge <- rep1_rep2_merge %>% 
    RunHarmony(c("orig.ident"), plot_convergence = TRUE, assay.use = "SCT",
               max.iter.harmony = 20)
```

Correlation for different resolution parameters

```{r, warning=FALSE, message=FALSE}
resolution_pars <- seq(from = 1, to = 2, by = 0.1)

correlations <- resolution_pars %>% 
  map_dbl(function(x){
    rep1_rep2_merge <- ComputeClustering(rep1_rep2_merge, x)
    ComputeCorrelation(rep1_rep2_merge, results_C2L)
    
  } )
df_A120838 <- data.frame(resolution_pars = numeric(length(resolution_pars)))
df_A120838$resolution_pars <- resolution_pars
df_A120838$correlations <- correlations
df_A120838$patient_ID <- "A120838"
```

## Merge Data frame for boxplot

```{r, warning=FALSE, message=FALSE, dpi=300}
df_all <- rbind(df_A121573,
  df_A416371,
  df_A551763,
  df_A595688,
  df_A798015,
  df_A938797,
  df_A120838)

df_all <- df_all %>%
  dplyr::left_join(df_patient_ID_transform, c("patient_ID" = "patient"))


df_all %>%
  ggplot( aes(x=patient_ID.y, y=correlations, fill=patient_ID.y)) +
    geom_boxplot(outlier.size = 0.75, alpha= 0.75) +
    # scale_fill_brewer(discrete = TRUE, alpha=0.6) +
    geom_jitter(color="black", size=0.75, alpha=1) +
    theme_linedraw() +  
    theme(legend.position="none", axis.title.x = element_blank(), 
          axis.text = element_text(size=12, face ="bold", family="Arial"),
          axis.text.x = element_text(angle = 90, hjust=0.95,vjust=0.2), 
          axis.title.y = element_text(size=16,family="Arial"),
          axis.ticks = element_blank(),  panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank()) +
  ylab("Pearson Correlation Coefficients")
```



# Conclusion 


# Session Info Details

```{r, echo=FALSE, eval=TRUE}
sessionInfo()
```